---
editor_options: 
  markdown: 
    wrap: none
---

# Results {#results}

\ChapFrame[Results][bhtteal]

This chapter presents the main results for the two primary research questions and the supplementary research question addressed in this thesis:

```{r, include=knitr::is_html_output(), results='asis', echo=FALSE}
cat("<ol class='rs-questions' style='--counter: 1;'><li>How can we use LLMs effectively to locate specific information in a financial report?</li><li>How can we use LLMs effectively to extract this information from the document?</li><li>Can we use additional information from the extraction process to guide the user on which values need to be checked and which can be trusted as they are?</li></ol>")
```

\begin{enumerate}[label={\textbf{Q\theenumi}}]
  \item How can we use LLMs effectively to locate specific information in a financial report?
  \item How can we use LLMs effectively to extract this information from the document?
  \item Can we use additional information from the extraction process to guide the user on which values need to be checked and which can be trusted as they are?
\end{enumerate}

Section \@ref(page-identification-introduction) presents the results for the first research question. Section \@ref(table-extraction-introduction) presents the results for the second question. Section \@ref(error-rate-guidance-results) shows the results for the side research question. Finally, we will summarize all results in Section \@ref(summary-results).

Section \@ref(page-identification-introduction) presents the results for the first research question. Section \@ref(table-extraction-introduction) presents the results for the second question. Section \@ref(error-rate-guidance-results) shows the results for the side research question. Finally, all results are summarized in Section \@ref(summary-results).

Each section begins with an overview of the specific sub-tasks, as well as the models, methods and data used to investigate the research question.

This chapter focuses on presenting results that demonstrate whether an approach can effectively solve a given task. Additional findings are discussed in Chapter \@ref(discussion). Further results are provided in the appendix, with references to relevant chapters documenting our investigation and the results of sub-tasks at the beginning of each following section.

## Page Identification {#page-identification-introduction}

```{r page-identification-llm-data-loading, echo=echo_flag, warning=warning_flag, message=message_flag}
temp_list <- readRDS("data_storage/page_identification_llm.rds")
df_binary <- temp_list$df_binary %>% mutate(
        # model_family = sub("_.*", "", model),
        model_family = if_else(str_detect(model, "Qwen2"), "Qwen 2.5", model_family),
        model_family = if_else(str_detect(model, "Qwen3"), "Qwen 3", model_family),
        model_family = if_else(str_detect(model, "Llama-3"), "Llama-3", model_family),
        model_family = if_else(str_detect(model, "Llama-4"), "Llama-4", model_family),
        model = str_remove(model, "-0-9-1")
      )
df_multi <- temp_list$df_multi %>% mutate(
        # model_family = sub("_.*", "", model),
        model_family = if_else(str_detect(model, "Qwen2"), "Qwen 2.5", model_family),
        model_family = if_else(str_detect(model, "Qwen3"), "Qwen 3", model_family),
        model_family = if_else(str_detect(model, "Llama-3"), "Llama-3", model_family),
        model_family = if_else(str_detect(model, "Llama-4"), "Llama-4", model_family),
        model = str_remove(model, "-0-9-1")
      )

method_families <- c("zero_shot", "law_context", "top_n_rag_examples", "n_random_examples", 'n_rag_examples')

method_familiy_colors <- c(
  "zero_shot" = "#e41a1c", 
  "law_context" = "#377eb8", 
  "top_n_rag_examples" = "#4daf4a", 
  "n_random_examples" = "#984ea3", 
  'n_rag_examples' = "#ff7f00"
  )

model_by_size_classification <- c('google_gemma-3-4b-it', 'google_gemma-3n-E4B-it', "google_gemma-3-12b-it",
  "google_gemma-3-27b-it", "meta-llama_Llama-3.1-8B-Instruct", 
  "meta-llama_Llama-3.1-70B-Instruct", "meta-llama_Llama-3.3-70B-Instruct",
  "meta-llama_Llama-4-Scout-17B-16E-Instruct", "meta-llama_Llama-4-Maverick-17B-128E-Instruct-FP8",
  "mistralai_Ministral-8B-Instruct-2410", "mistralai_Mistral-Small-3.1-24B-Instruct-2503",
  "mistralai_Mistral-Large-Instruct-2411", "Qwen_Qwen2.5-0.5B-Instruct",
  "Qwen_Qwen2.5-1.5B-Instruct", "Qwen_Qwen2.5-3B-Instruct", "Qwen_Qwen2.5-7B-Instruct",
  "Qwen_Qwen2.5-14B-Instruct", "Qwen_Qwen2.5-32B-Instruct", "Qwen_Qwen2.5-72B-Instruct",
  "Qwen_Qwen3-8B", "Qwen_Qwen3-30B-A3B-Instruct-2507", "Qwen_Qwen3-32B", "Qwen_Qwen3-235B-A22B-Instruct-2507",
  "tiiuae_Falcon3-10B-Instruct", "microsoft_phi-4"
  ) %>% gsub("^[^_]+_", "", .)

df_binary <- df_binary %>% 
  mutate(
  n_examples = as.numeric(n_examples),
  n_examples = if_else(method_family == "zero_shot", 0, n_examples),
  n_examples = if_else(method_family == "law_context", 1, n_examples),
  method_family = factor(method_family, levels = method_families)
) %>% mutate(model = gsub("^[^_]+_", "", model)) %>% 
  filter(model %in% model_by_size_classification) %>% 
  ungroup() # %>% 
  # mutate(
  #   model = str_replace(model, "_vllm", ""),
  #   model_family = sub("_.*", "", model),
  #   model_family = if_else(str_detect(model, "Qwen2"), "Qwen 2.5", model_family),
  #   model_family = if_else(str_detect(model, "Qwen3"), "Qwen 3", model_family),
  #   model_family = if_else(str_detect(model, "Llama-3"), "Llama-3", model_family),
  #   model_family = if_else(str_detect(model, "Llama-4"), "Llama-4", model_family)
  # )

binary_task <- list()
binary_task$n_models <- df_binary$model %>% unique() %>% length()
binary_task$n_model_families <- df_binary$model_family %>% unique() %>% length()
binary_task$n_method_families <- df_binary$method_family %>% unique() %>% length()

top_performer_binary <- df_binary %>% 
  filter(is.finite(f1_score), loop == 0) %>% 
  filter(n_examples <= 3 | is.na(n_examples)) %>%
  group_by(model_family, classification_type) %>% 
  slice_max(n = 1, f1_score) %>% 
  arrange(desc(f1_score)) %>% 
  select(model_family, model, classification_type, method_family, n_examples, f1_score, norm_runtime) %>%
  mutate(
    f1_score = round(f1_score, 2),
    norm_runtime = round(norm_runtime, 0),
  ) %>% rename(
    "runtime in s" = norm_runtime,
  )

top_performer_binary_median <- top_performer_binary %>% group_by(classification_type) %>% summarise(median = median(f1_score))
top_performer_binary_median_guv <- top_performer_binary_median %>% filter(classification_type == "GuV") %>% pull(median)
top_performer_binary_median_aktiva <- top_performer_binary_median %>% filter(classification_type == "Aktiva") %>% pull(median)
top_performer_binary_median_passiva <- top_performer_binary_median %>% filter(classification_type == "Passiva") %>% pull(median)

df_multi <- df_multi %>% 
  mutate(
  n_examples = as.numeric(n_examples),
  n_examples = if_else(method_family == "zero_shot", 0, n_examples),
  n_examples = if_else(method_family == "law_context", 1, n_examples),
  method_family = factor(method_family, levels = method_families)
) %>% mutate(model = gsub("^[^_]+_", "", model)) %>% 
  filter(model %in% model_by_size_classification) %>% 
  ungroup() #%>%
  # mutate(
  #   model = str_replace(model, "_vllm", ""),
  #   model_family = sub("_.*", "", model),
  #   model_family = if_else(str_detect(model, "Qwen2"), "Qwen 2.5", model_family),
  #   model_family = if_else(str_detect(model, "Qwen3"), "Qwen 3", model_family),
  #   model_family = if_else(str_detect(model, "Llama-3"), "Llama-3", model_family),
  #   model_family = if_else(str_detect(model, "Llama-4"), "Llama-4", model_family)
  # )

top_performer_multu <- df_multi %>% 
  unnest(metrics) %>% 
  filter(metric_type %in% c("Aktiva", "Passiva", "GuV")) %>% 
  filter(is.finite(f1_score), loop == 0) %>% 
  filter(n_examples <= 3 | is.na(n_examples)) %>%
  group_by(model_family, metric_type) %>% 
  slice_max(n = 1, f1_score) %>% 
  arrange(desc(f1_score)) %>% # head(10) %>% 
  select(model_family, model, metric_type, method_family, n_examples, f1_score, norm_runtime) %>%
  mutate(
    f1_score = round(f1_score, 2),
    norm_runtime = round(norm_runtime, 0),
  ) %>% rename(
    "runtime in s" = norm_runtime,
  )
```

(ref:page-identification-intro-text) The first research question examines how \acrfull{LLM}s can be used to effectively locate specific information in a financial report. The task for this thesis is to identify the pages containing the balance sheet (*Bilanz*) and the profit-and-loss statement (*Gewinn- und Verlustrechnung, GuV*). The balance sheet consists of two tables: assets (*Aktiva*) and liabilities (*Passiva*), which are often presented on separate pages. Hereafter, the German terms **Aktiva**, **Passiva** and **GuV** will be used.

(ref:page-identification-intro-text)

@liExtractingFinancialData2023 outlines two methods for identifying relevant pages (see Figure \@ref(fig:extraction-framework-flow-chart)). For longer documents, they recommend using the \acrfull{TOC} to determine a page range that encompasses the information of interest. Additionally, they develop targeted regular expressions and rules to filter out irrelevant pages[^05_results-1]. The outcome of this "Page Range Refinement" is subsequently provided to the \acr{LLM} for information extraction.

[^05_results-1]: In our experience, developing effective regular expressions can be very time-consuming, and setting appropriate rules requires domain expertise. While this effort is worthwhile for large collections of similar documents, it took several months for this thesis. Fortunately, the resulting pipeline can be reused by updating the rules and keyword lists, enabling faster solutions for future tasks.

This section presents the results of four approaches to identify the page[^05_results-2] of interest. A detailed report on all experiments and their results is available in the appendix:

[^05_results-2]: In some cases, the information of interest spans two pages. These rare cases are not yet addressed by the approaches presented here.

-   Section \@ref(regex-page-identification) presents the findings of page range refinement using a list of keywords and regular expressions.
-   Section \@ref(toc-understanding) presents the findings of a \acr{TOC} understanding approach.
-   Section \@ref(llm-page-identification) presents the findings of a text classification approach using \acr{LLM}s.
-   Section \@ref(tf-classifier) presents the findings of a term-frequency approach.

We compare and summarize the results in Section \@ref(comparison-page-identification). Section \@ref(efficency-discussion) proposes an efficient combination of approaches to solve the task and discusses its limitations.

###  {.unlisted .unnumbered}

::: paragraph-start
##### Dataset Description

For the page identification task, we selected companies primarily from the first row of Figure \@ref(fig:beteiligungsunternehmen) to construct the ground truth. The main motivation is that documents from companies within a category are more similar to each other than to those from other categories. For the selected companies, all available annual reports are included. Since one company mainly published documents requiring \acrfull{OCR} preprocessing, we added documents from a second company in this category.
:::

### Approaches

For all approaches presented in this section except the \acr{TOC} understanding approach, the page identification task is formulated as a classification problem. This subsection briefly describes our approaches. A detailed report is provided in the Appendix \@ref(page-identification-report).

::: paragraph-start
##### Regular Expressions

We developed multiple sets of regular expressions to filter out pages that do not meet all criteria in a given set. Different sets are used for each target type: **Aktiva**, **Passiva** and **GuV**. The sets vary in their ability to handle additional whitespace introduced by imperfect text extraction and in the range of accepted synonyms for each term.
:::

::: paragraph-start
##### Table of Contents Understanding

We use a \acr{LLM} to extract the \acr{TOC} from the initial pages of a document or use the embedded \acr{TOC}, prompting the \acr{LLM} to identify the pages where **Aktiva**, **Passiva** and **GuV** are located.
:::

::: paragraph-start
##### Large Language Model Classification

We use \acr{LLM}s to classify whether the text extract of a given page contains an **Aktiva**, **Passiva** or **GuV** table, or something else. Both binary and multi-class classification approaches are tested. The reported confidence scores are used to rank which text extracts are most similar to the target type.
:::

We evaluate a wide range of open-weight models and compare different prompting techniques. In addition to zero-shot approaches, we test few-shot in-context learning with examples selected either randomly or based on vector similarity. We also test passing the legal text instead of examples from annual reports.

::: paragraph-start
##### Term Frequency Ranking

We use normalized term frequencies and normalized float frequency as features for classification with a \acrfull{RF}. The predicted scores are used to rank pages by their likelihood of containing the target information. Undersampling is applied during training to address data imbalance.
:::

### Comparison {#comparison-page-identification}

```{r load-page-identification-comapison-data, echo=echo_flag, warning=warning_flag, message=message_flag}
df_page_identification_comparing_performance <- readODS::read_ods("data_storage/page_identification_summary.ods", sheet = 1)
df_page_identification_comparing_efficiency <- readODS::read_ods("data_storage/page_identification_summary.ods", sheet = 2)

df_df_page_identification_comparing_performance_formatted <- df_page_identification_comparing_performance %>% 
  group_by(type) %>% 
  mutate(
    across(
      c(precision, recall, F1, `top 1 recall`),
      ~ifelse(      
        . == max(., na.rm = TRUE),
        paste0("**", round(., 3), "**"),
        round(., 3)
        )
      )) %>% 
  mutate(`k for full recall` = ifelse(      
        `k for full recall` == min(`k for full recall`, na.rm = TRUE),
        paste0("**", round(`k for full recall`, 3), "**"),
        round(`k for full recall`, 3)
        ))
```

This subsection presents the performance and efficiency for all four presented approaches and compares it with the the results a human achieves manually

::: paragraph-start
##### Prediction Performance

Table \@ref(tab:page-identification-performance-overview-all) presents the best performance achieved by the four approaches in terms of precision, recall and F1 score.

```{r page-identification-performance-overview-all, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
df_df_page_identification_comparing_performance_formatted %>% 
  select(-c(`top 1 recall`, `k for full recall`)) %>% 
  arrange(type) %>% 
  render_table(
    alignment = "lllrrr",
    caption = "Comparing page identification perfromance among all four approaches.",
    ref = opts_current$get("label"),
    row_group_col = 3
    )
```

Llama 4 Scout achieves the highest F1 score for **Aktiva** and **Passiva** in the multi-class classification approach. For **GuV**, the best F1 score (0.985) is obtained with Ministral-8B-Instruct in the binary classification approach. Llama 4 Scout achieves an F1 score of 0.971 for **GuV** in multi-class classification.
:::

In preparing the dataset for the table extraction task (see Section \@ref(ground-truth-creation)), 106 **Aktiva** pages were selected. During this manual process, two mistakes were made, accidentally selecting one **Passiva** and one **GuV** page. Thus, the human baseline to compete with is 0.981. Llama 4 Scout is therefore more precise than manual selection.

Furthermore, Llama 4 Scout achieves a recall of 1.0 for all target types, meaning the results can be used downstream even if precision is not always perfect. Pages classified as targets can be double-checked by a human without missing any relevant page.

The performance of approaches that do not utilize \acr{LLM}s is substantially lower. Only the term-frequency approach yields results suitable for downstream use, as it achieves a recall of 1.0. Table \@ref(tab:page-identification-performance-overview-top-k-recall) shows the top k recall results for the term-frequency and \acr{LLM} approaches. For the term-frequency approach, the ranking is based on the \acr{RF} classifier scores; for \acr{LLM}s, it is based on calculated confidence scores.

The \acr{LLM}s consistently rank the correct **GuV** page highest. With Llama Scout 4, all target pages are found within the top two ranked pages. For the term-frequency approach, a human may need to check up to five pages.

```{r page-identification-performance-overview-top-k-recall, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
df_df_page_identification_comparing_performance_formatted %>% 
  select(-c(precision, recall, F1)) %>% 
  filter(!is.na(`top 1 recall`)) %>% 
  arrange(type) %>% 
  render_table(
    alignment = "lllrr",
    caption = "Comparing the top k recall for the term-frequency and LLM approaches.",
    ref = opts_current$get("label"),
    row_group_col = 3
    )
```

::: paragraph-start
##### Energy Usage and Runtime

Table \@ref(tab:page-identification-efficiency-overview-all) presents the runtime per document (in seconds), estimated energy consumption (in Joules per document), and costs (in CENTS per 1,000 documents). The runtime for \acr{LLM}s is normalized to seconds on a NVIDIA B200, with a \acrfull{TDP} of 700 W used to calculate energy consumption. For other approaches running on a laptop (see Section \@ref(local-machine)), a \acr{TDP} of 28 W is used. For manual work, an additional 60 W is added for the screen. It is assumed that the \acr{LLM} is hosted locally.
:::

```{r page-identification-efficiency-overview-all, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
table_energy <- df_page_identification_comparing_efficiency %>% filter(!drop == 1) %>% 
  select(
    approach, strategy, 
    `runtime per document in s`, `energy in J`, 
    `costs in CENTS per 1000 documents`
    ) %>%
  mutate(
    across(
      where(is.numeric),
      ~ifelse(      
        . == min(., na.rm = TRUE),
        paste0("**", format(round(., 3), nsmall=3), "**"),
        format(round(., 3), nsmall=3)
        )
      )) %>% 
  render_table(
    alignment = "llrrr",
    caption = "Comparing page identification efficiency among all four approaches.",
    ref = opts_current$get("label")
    )
if(knitr::is_latex_output()) {
  table_energy <- table_energy %>% 
    column_spec(2, width = "3.7cm") %>% 
    column_spec(3, width = "2.3cm") %>% 
    column_spec(5, width = "3cm")
}
table_energy
```

Table \@ref(tab:page-identification-efficiency-overview-all) shows that the regular expression approach is fastest and consumes the least energy. However, since its results are insufficient, another approach must be chosen to reduce manual labor for the human in the loop.

Second place in all criteria is the term-frequency approach, which guarantees perfect recall while reducing the number of pages to investigate to five per target type. This is similar to the number of pages a human must check to find the \acr{TOC} of the document, and represents a reduction to 7.4% of the average 67 pages per document in this dataset. The costs remain negligible.

The \acr{LLM} approaches have the highest runtime and energy consumption, as they process every page with computationally intensive algorithms. For the \acr{TOC} approach, \acr{LLM}s are also used, but they process far fewer pages, resulting in lower energy consumption.

Since all approaches except manual identification require text extraction, this runtime and energy consumption are not listed and negligible.

## Information Extraction {#table-extraction-introduction}

```{r loading-table-extraction-datasets, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum(c("data_storage/real_table_extraction_extended_llm.rds", "data_storage/real_table_extraction_extended_synth.rds", "data_storage/real_table_extraction_extended_azure.rds"))}
df_real_table_extraction <- readRDS("data_storage/real_table_extraction_extended_llm.rds") %>% 
  filter(!model %in% c("deepseek-ai_DeepSeek-R1-Distill-Qwen-32B", 'google_gemma-3n-E4B-it')) %>% 
  mutate(
    model = gsub("^[^_]+_", "", model),
    company = map_chr(filepath, ~str_split(str_split(., "/")[[1]][5], "__")[[1]][1])
    ) %>% filter(company != "MEAB GmbH")
df_real_table_extraction_synth <- readRDS("data_storage/real_table_extraction_extended_synth.rds") %>% 
  mutate(
    model = gsub("^[^_]+_", "", model),
    company = map_chr(filepath, ~str_split(str_split(., "/")[[1]][5], "__")[[1]][1])
    ) %>% filter(company != "MEAB GmbH")
df_real_table_extraction_azure <- readRDS("data_storage/real_table_extraction_extended_azure.rds") %>% 
  mutate(
    model = gsub("^[^_]+_", "", model),
    company = map_chr(filepath, ~str_split(str_split(., "/")[[1]][5], "__")[[1]][1])
    ) %>% filter(company != "MEAB GmbH")

model_by_size <- c(
  'gemma-3-4b-it', #'gemma-3n-E4B-it', 
  "gemma-3-12b-it", "gemma-3-27b-it",
  "gpt-oss-20b", "gpt-oss-120b",
  "Llama-3.1-8B-Instruct", "Llama-3.1-70B-Instruct", "Llama-3.3-70B-Instruct",
  "Llama-4-Scout-17B-16E-Instruct", "Llama-4-Maverick-17B-128E-Instruct-FP8",
  "Ministral-8B-Instruct-2410", "Mistral-Small-3.1-24B-Instruct-2503",
  "Mistral-Large-Instruct-2411", "Qwen2.5-0.5B-Instruct",
  "Qwen2.5-1.5B-Instruct", "Qwen2.5-3B-Instruct", "Qwen2.5-7B-Instruct",
  "Qwen2.5-14B-Instruct", "Qwen2.5-32B-Instruct", "Qwen2.5-72B-Instruct",
  "Qwen3-0.6B", "Qwen3-1.7B", "Qwen3-4B",
  "Qwen3-8B", "Qwen3-14B", "Qwen3-30B-A3B-Instruct-2507", "Qwen3-32B",
  "Qwen3-235B-A22B-Instruct-2507-FP8", "Qwen3-235B-A22B-Instruct-2507",
  # "gpt-4.1-nano", "gpt-4.1-mini", "gpt-4.1",
  "Falcon3-10B-Instruct", "phi-4"
)

method_order <- c("top_n_rag_examples", "n_random_examples", "top_n_rag_examples_out_of_sample", "static_example", "zero_shot" )

norm_factors <- read_csv("../benchmark_jobs/page_identification/gpu_benchmark/runtime_factors_real_table_extraction.csv") %>% 
  mutate(
    model_name = model_name %>% str_replace("/", "_")
  )
norm_factors_few_examples <- norm_factors %>% filter((str_ends(filename, "binary.yaml") | str_ends(filename, "multi.yaml") | str_ends(filename, "vllm_batched.yaml")))

df_real_table_extraction <- df_real_table_extraction %>% left_join(
  norm_factors_few_examples %>% mutate(model_name = gsub("^[^_]+_", "", model_name)) %>% select(model_name, parameter_count, normalization_factor),
  by = c("model" = "model_name")
  ) %>% mutate(normalized_runtime = round(runtime*normalization_factor, 0))

df_overview <- bind_rows(df_real_table_extraction, df_real_table_extraction_azure) %>% 
  filter(out_of_company != TRUE | is.na(out_of_company), n_examples <= 5) %>% 
  filter(model %in% model_by_size) %>%
  filter(n_examples != 2) %>% 
  mutate(
    model = factor(model, levels = model_by_size),
    method_family = factor(method_family, levels = method_order),
    n_examples = fct_rev(ordered(paste("n =", n_examples)))
    )

units_real_tables <- read_csv("../benchmark_truth/real_tables_extended/table_characteristics_more_examples.csv") %>% mutate(
  filepath = paste0('/pvc/benchmark_truth/real_tables_extended/', company, '__', filename),
  T_EUR = (T_in_year + T_in_previous_year)>0,
  T_EUR_both = (T_in_year + T_in_previous_year)>1,
  n_T_EUR = T_in_year + T_in_previous_year
) %>% select(filepath, T_EUR, T_EUR_both, n_T_EUR)

df_real_table_extraction <- df_real_table_extraction %>% left_join(units_real_tables) %>% 
  mutate(
    .before = 1, 
    company = map_chr(filepath, ~str_split(str_split(., "/")[[1]][5], "__")[[1]][1]),
    same_company = !out_of_company
    )
df_real_table_extraction_synth <- df_real_table_extraction_synth %>% 
  left_join(units_real_tables) %>% 
  mutate(
    .before = 1, 
    company = map_chr(filepath, ~str_split(str_split(., "/")[[1]][5], "__")[[1]][1]),
    same_company = !out_of_company
    )
```

(ref:information-extraction-intro-text) The second research question investigates how \acr{LLM}s can be used to effectively extract specific information from a financial report. The task for this thesis is to extract the numeric values from the assets (*Aktiva*) table, which is part of the balance sheet (*Bilanz*). Hereafter, the German term **Aktiva** will be used. The scope is further limited compared to Section \@ref(page-identification-introduction), as manual creation of the reference dataset is time-consuming.

(ref:information-extraction-intro-text)

###  {.unlisted .unnumbered}

::: paragraph-start
##### Structured Output

We use a strict schema for the extraction process, derived from the legal text [@hgbHandelsgesetzbuchImBundesgesetzblatt2025, §266]. Three levels of verbosity are defined by law, with smaller companies permitted to produce less detailed balance sheets. Our schema is based on the most detailed level, which is most commonly found in the document base[^05_results-3].
:::

[^05_results-3]: Unfortunately, well-known companies such as BVG and BSR publish less detailed forms, so their documents are not included in the document base for this task.

Using a strict schema offers advantages for downstream processing, such as adding results to a relational database. It also simplifies comparison with ground truth, as row names and order are fixed. The schema is defined as \acr{ebnf} grammar and passed as an argument to \acr{vllm}.

::: paragraph-start
##### Ground Truth Dataset

For information extraction, we use two datasets. First, a collection of 106 real **Aktiva** tables is created through two sampling iterations. In the first iteration, a single report is selected for each company, and all available reports from the first listed company are chosen to test an in-company learning approach. In the second iteration, more reports from other companies are added to increase the ground truth size and enable in-company testing for all companies.
:::

Second, a dataset of 16_504 synthetic **Aktiva** tables is generated. These tables are created based on the extraction schema and filled with random numeric values. Different table characteristics are systematically combined to investigate their effects on extraction performance. The tables are produced as \acrfull{PDF}, \acrfull{HTML} and Markdown files. This dataset allows estimation of extraction performance when no unknown row identifiers are present.

### Approaches

::: paragraph-start
##### Regular Expressions

We use regular expressions to extract numeric values for matching row identifiers. The regular expressions handle line breaks between words in row identifiers, but not within words. They accommodate multiple whitespace characters and attempt to fully match labels from the legal text with the text extract, ignoring case. Numbers with "." as thousands separator are extracted.
:::

::: paragraph-start
##### Real Tables

We use \acr{LLM}s to extract numeric values from real **Aktiva** tables using restricted generation. The \acr{LLM} must group row identifiers and corresponding numeric values, matching row identifiers to schema labels. Unknown row identifiers are discarded and missing labels result in *null* predictions. All values are extracted in one pass. No instructions are provided regarding currency units for specific columns.
:::

We evaluate a wide range of open-weight models and compare different prompting techniques. In addition to zero-shot approaches, we test few-shot in-context learning with examples selected randomly or by vector similarity. We also test using a synthetic **Aktiva** table as an example. Models from OpenAI's GPT family are included alongside open-weight models.

::: paragraph-start
##### Synthetic Tables

We use \acr{LLM}s to extract numeric values from synthetic **Aktiva** tables using restricted generation. The procedure is identical to that for real **Aktiva** tables. Values are extracted with and without explicit instructions regarding currency units. Tests are limited to open-weight models.
:::

::: paragraph-start
##### Hybrid Approach

We use \acr{LLM}s to extract numeric values from real **Aktiva** tables using restricted generation, providing examples from synthetic **Aktiva** tables. The procedure is identical to that for real **Aktiva** tables. Values are extracted with and without explicit instructions regarding currency units. Tests are limited to open-weight models.
:::

### Comparison {#comparing-table-extraction-methods}

This subsection compares the results for the table extraction tasks. It will discuss the findings about performance and runtime and compare it with the results a human may achieve wit manual labor. The detailed report on all experiments conducted in their results can be found in the Appendix \@ref(information-extraction-report). The evaluation of feature importance can be found in Appendix \@ref(feature-effect-analysis).

::: paragraph-start
##### Performance

Table \@ref(tab:compare-extraction-performance) summarizes the mean percentage of correct predictions for all approaches and both types of **Aktiva** tables. The highest baseline for extraction tasks is set by our own manual performance, achieving 97.6% correct values on real **Aktiva** tables. The \acr{regex} approach performs nearly as well on synthetic tables but is much less accurate on real tables.
:::

The mean performance of Qwen3-235B does not match our baseline on real **Aktiva** tables, but its median performance is already 100%.

On synthetic tables, its mean performance is nearly perfect when currency units are respected. With \acr{HTML} documents, we find 100% correct predictions; with Markdown documents, 99.9%. Figure \@ref(fig:comparing-table-extraction-performance-among-real-and-synth-aktiva-data) shows that better performance on synthetic tables is observed for almost all models.

Qwen3-8B performed best among the smaller \acr{LLM}s but produced over 4% more incorrect predictions than Qwen3-235B.

Using synthetic examples results in lower performance, but it can demonstrate how to handle currency units.

```{r compare-extraction-performance, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
tribble(
  ~approach, ~strategy, ~table_type, ~percentage_correct_total,
  "human", "manual", "real", 97.6,
  "regex", "", "real", 68.6,
  "llm", "Qwen3-235B, top_5_rag_examples", "real", 97.0,
  "llm", "Qwen3-8B, top_5_rag_examples", "real", 92.7,
  "llm", "Qwen3-235B, top_5_rag_examples, synth examples", "real", 91.8,
  "regex", "", "synth", 96.9,
  "llm", "Qwen3-235B, top_5_rag_examples, respect_units", "synth", 99.9,
  "llm", "Qwen3-8B, top_5_rag_examples", "synth", 94.6,
) %>% # arrange(table_type) %>% 
  group_by(table_type) %>% 
  mutate(
    across(
      c(percentage_correct_total),
      ~ifelse(      
        . == max(., na.rm = TRUE),
        paste0("**", round(., 3), "**"),
        round(., 3)
        )
      )) %>% 
  rename(mean_percentage_correct_total = percentage_correct_total) %>% 
  render_table(
    alignment = "lllr",
    caption = "Comparing the mean percentage of correct predictions total among all approaches and table types.",
    ref = opts_current$get("label"),
    row_group_col = 3
    )
```

::: paragraph-start
##### Runtime

Extracting values from all 106 tables took Qwen3-235B approximately six minutes. Excluding \acr{LLM} setup time, Qwen3-235B-A22B-Instruct is about 100 times faster than manual extraction.
:::

::: paragraph-start
##### Feature Effects

Currency unit is the predictor with the strongest effect across all approaches. Incorporating this into table extraction is crucial for optimizing performance. For \acr{LLM}-based approaches, most model- and method-related variables have a strong effect. Using a versatile model and providing high-quality learning examples is essential.
:::

Findings from approaches using synthetic tables suggest that input format can significantly influence extraction outcomes. Preventing errors during text extraction and converting extracted text to \acr{HTML} may help resolve ambiguities. However, it remains unclear whether a perfect text extract would perform as well as \acr{HTML} or Markdown formats.

## Error Rate Guidance {#error-rate-guidance-results}

(ref:error-rate-guidance-intro-text) The supplementary research question investigates whether it is possible to guide user attention to predictions with higher empirical error rates. This thesis focuses on the confidence score calculated from reported log probabilities.

(ref:error-rate-guidance-intro-text)

Section \@ref(page-identification-error-rate-results) presents results for the page identification task. Section \@ref(information-extraction-error-rate-results) presents results for the information extraction task. Detailed reports are available in the Appendix \@ref(error-guidance-report).

### Page Identification {#page-identification-error-rate-results}

We find that the confidence score can be used in the page identification task to identify confidence intervals containing few or no errors. The number of predictions in these intervals varies by model, classification task and target class.

::: paragraph-start
##### Binary Classification

Distinguishing correct from incorrect classifications based on confidence score works well for Ministral-8B-Instruct responses, but less so for other models, such as those from the Qwen family. This is because Ministral-8B-Instruct reports confidence scores over a wide range, while Qwen 2.5 models consistently report high confidence.
:::

Thus, it is possible to define broad confidence intervals with zero empirical error rate for Ministral. For Qwen 2.5 32B, the highest confidence interval contains some mistakes, but the error rate remains below 1%. Most predictions for both models fall within the highest confidence interval.

::: paragraph-start
##### Multi-class Classification

Distinguishing correct from incorrect classifications based on confidence score continues to work well for Ministral-8B-Instruct when predicting **Aktiva** and **GuV**. For **Passiva**, a single incorrect prediction occurs in the highest confidence interval. For Llama 4 Scout, this approach works for **Aktiva** and **Passiva**. For Qwen 2.5 32B, this approach is the least effective among the three models, although all achieve high F1 scores.
:::

For Ministral, the empirical error rate is zero in the highest confidence interval, except for **Passiva**. For Llama, the error rate is less than 1% for all classes. For Qwen 2.5 32B, only **Passiva** has a high confidence interval with an error rate below 1%. Again, most predictions for all models are in the highest confidence interval.

### Information Extraction {#information-extraction-error-rate-results}

We find that the confidence score alone cannot be used in the information extraction task to identify confidence intervals with few or no errors. Only a few exceptions achieve an error rate below 1% across all annual reports. Results for the best-performing model are reported here.

::: paragraph-start
##### Real Tables

For the best-performing model, Qwen3-235B-A22B-Instruct, the empirical error rate is 1.3% for predicting missing values and 3.3% for predicting numeric values. Almost all predictions fall within the highest confidence interval. Reducing the interval width does not yield intervals with lower error rates.
:::

::: paragraph-start
##### Synthetic Tables

For the best-performing model, Qwen3-235B-A22B-Instruct, the empirical error rate is below 1% for predicting missing values and numeric values when explicit instructions regarding currency units are provided. If the input format is perfect \acr{HTML}, the error rate is 0%. With perfect Markdown, the error rate is above 0% but below 1%.
:::

::: paragraph-start
##### Hybrid Approach

For the best-performing model, Qwen3-235B-A22B-Instruct, the empirical error rate is below 4% for predicting missing values and around 20% to 26% for predicting numeric values. The error rate is lower when explicit instructions regarding currency units are provided.
:::

## Summary {#summary-results}

In summary, the results presented in this chapter demonstrate the effectiveness of \acr{LLM}s for both page identification and information extraction tasks in financial reports. \acr{LLM}-based approaches consistently outperform traditional methods, achieving high recall and precision and enabling efficient automation of information extraction. The experiments further highlight the importance of input format and model selection, as well as the potential for guiding user attention through confidence scores.

These findings provide a solid foundation for the subsequent discussion, where we interpret the results in the context of our research questions and hypotheses, analyze sources of error and explore practical implications for future applications.
