---
editor_options: 
  markdown: 
    wrap: 72
---

# Results

\ChapFrame

This chapter presents the results for the two research questions of this
thesis:

1.  How can we use \acr{LLM}s effectively to locate specific information
    in a financial report?
2.  How can we use \acr{LLM}s effectively to extract these information
    from the document?

Section \@ref(page-identification-introduction) presents the results for
the first research question. Section
\@ref(table-extraction-introduction) presents the results for the second
question.

Each section will start with an overview about the specific sub tasks as
well about the models, methods and data used to investigate the research
question. The subsections present the results of the sub tasks. At the
end of each section all results get compared and summarized.

## Page identification {#page-identification-introduction}

```{r count-benchmark-documents, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum(c("../benchmark_truth/aktiva_passiva_guv_table_pages_no_ocr.csv", "./scripts/page_identification_preparations.R"))}
source("./scripts/page_identification_preparations.R")

num_multiple_tables_per_document <- df_targets_no_ocr %>%
  anti_join(consecutive_pages, by = c("filepath", "page", "type")) %>% 
  arrange(filepath, page) %>%
  group_by(filepath, type) %>% 
  summarise(count = n()) %>% 
  filter(count > 1) %>%
  group_by(filepath) %>% 
  summarise(count = n()) %>% 
  nrow()

multiple_tables_per_type_and_document <- data_unnested %>%
  anti_join(consecutive_pages, by = c("filepath", "page", "type")) %>% 
  arrange(filepath, page) %>%
  group_by(filepath, type) %>% 
  summarise(count = n()) %>% 
  filter(count > 1) %>% 
  group_by(type) %>% 
  summarise("multiple targets in document" = n())

df_special_targets <- multiple_tables_per_type_and_document %>% left_join(consecutive_pages %>% group_by(type) %>% summarise("target two pages long" = n())) %>% mutate_if(is.numeric,coalesce,0)
```

The first research question asks, how \acr{LLM}s can be used, to
effectively locate specific information in a financial report. The task
for this thesis is identifying the pages where the balance sheet
(*Bilanz*) and the profit-and-loss-and-statement (*Gewinn- und
Verlustrechnung, GuV*) are located. The balance sheet is composed of two
tables showing the assets (*Aktiva*) and liabilities (*Passiva*) of a
company. Often these two tables are on separate pages. Hereafter, the
German terms **Aktiva**, **Passiva** and \acr{GuV} will be used.

@li_extracting_2023 describes two ways to identify the relevant pages
(see Figure \@ref(fig:extraction-framework-flow-chart)). For longer
documents they propose to use the \acr{TOC} to determine a page range
that includes the information of interest. In addition, they develop
target specific regular expressions and rules to filter out irrelevant
pages[^05_results-1]. The result of this "Page Range Refinement" is then
passed to the \acr{LLM} to extract information from.

[^05_results-1]: Personal opinion: Developing well performing regular
    expressions can be a very tedious and setting appropriate rules
    requires some domain knowledge. It can be worth the effort if there
    are a lot of documents with similar information to extract. For this
    thesis it took multiple months. At least, now there is kind of a
    pipeline one can reuse, exchanging the rules and key word lists.
    Thus the next similar task should be solved faster.

This section is presenting four approaches to identify the
page[^05_results-2] of interest.

[^05_results-2]: In some cases the information of interest is spanning
    two pages. These rare cases are not covered from the approaches
    presented here, yet.

-   Subsection \@ref(regex-page-identification) presents the performance
    of a page range refinement using a list of key words with a regular
    expression.
-   Subsection \@ref(toc-understanding) presents the performance of a
    \acr{TOC} understanding approach
-   Subsection \@ref(llm-page-identification) presents the performance
    of a text classification using \acr{LLM}s.
-   Subsection \@ref(tf-classifier) presents the performance of a
    term-frequency approach.

In subsection \@ref(comparison-page-identification) the results get
compared and summarized. Subsection \@ref() proposes an efficient
combination of approaches to solve the task of this thesis and discusses
its limitations.

##### Dataset describtion

Figure \@ref(fig:document-base-page-identification) shows how the
document base for the tasks in this section is composed[^05_results-3].
Overall `r num_documents` annual reports from `r num_companies`
companies are used. For this thesis the tables of interest are those
that show **Aktiva**, **Passiva** and **GuV**. Among the
`r total_pages_no_ocr` pages `r num_tables` tables have to be identified
on `r num_target_pages` pages. Figure
\@ref(fig:document-base-page-identification) also gives an impression on
how many pages the documents have. The documents of *IBB* tend to be
longer. The documents of *Amt für Statistik Berlin-Brandenburg* tend to
be shorter.

[^05_results-3]: I downloaded all publicly available annual reports for
    some of the companies shown in the first row of Figure
    \@ref(fig:beteiligungsunternehmen). I assumed that this will give a
    representative sample of document structures for the other companies
    of the same type. Realizing that the degewo AG reports would require
    ocr preprocessing I additionally downloaded reports for GESOBAU AG.
    This approach could have been more systematic. For the second task I
    downloaded reports for all companies available and tried to use a
    balanced amount of reports per company.

```{r, eval = knitr::is_html_output(), echo=FALSE, warning=warning_flag, message=message_flag, results='asis', class.chunk='pre-detail'}
cat("<p>You can have a look at the ground truth data unfolding the following details element.</p>")
```

```{r ground-truth-tables-of-interest, eval = knitr::is_html_output(), echo=FALSE, warning=warning_flag, message=message_flag, class.chunk='hideme', results='asis'}
df_targets_no_ocr %>% render_table(
  caption = "Showing all pages of interest with the target table type and filepath.",
  ref = opts_current$get("label")
  )
```

```{r document-base-page-identification, echo=echo_flag, warning=warning_flag, message=message_flag, fig.cap="Showing the number of pages (bar height) and number of documents (number above the bar) per company for the data used for the page identification task. Some documents would require ocr before being processed and were not used.", results='asis'}
df_pages %>% group_by(company) %>% mutate(
  docs_per_company = n(), pages_per_company = sum(pages)
  ) %>% 
  ggplot() +
  geom_col(aes(x = company, y = pages, fill = needs_ocr), color = "#00000033") +
  geom_text(
    data = . %>% group_by(company) %>% slice_head(n = 1),
    aes(x = company, y = pages_per_company, label = docs_per_company, group = company),
    stat = "identity",
    vjust = -0.5,
    size = 3
  ) +
  scale_x_discrete(guide = guide_axis(angle = 30))
```

Table \@ref(tab:display_multiple_tables_per_type_and_document) shows how
many documents have multiple target tables per type and how many target
tables span two pages. In total `r num_consecutive_pages` tables are
distributed on two pages. In `r num_multiple_tables_per_document`
documents there are multiple tables per type of interest. There are
`r num_two_tables_on_one_page` pages with two target tables (**Aktiva**
and **Passiva**) on it.

```{r display-multiple-tables-per-type-and-document, echo=FALSE, warning=warning_flag, message=message_flag, results='asis'}
df_special_targets %>% render_table(
  caption = "Showing the number of documents with multiple target tables per type and the number of target tables that span two pages.",
  ref = opts_current$get("label")
  )
```

This task is broken down to a classification task all of the approaches
presented in this section but the \acr{TOC} understanding approach.

Thus, we prompt the \acr{LLM} to classify if the text extract of a given
page

for implementation: As described in \@ref(text-extraction-benchmark)
open source libraries have been used to extract the text from the annual
reports.

### Baseline: Regex {#regex-page-identification}

```{r page-identification-regex-data-loading, echo=echo_flag, warning=warning_flag, message=message_flag, message=FALSE, cache=TRUE, cache.extra = tools::md5sum("data_storage/page_identification_regex.rds")}
data_page_identification_regex <- readRDS("data_storage/page_identification_regex.rds")

metrics <- data_page_identification_regex$metrics
metric_summaries <- data_page_identification_regex$metric_summaries
metrics_by_company_and_type <- data_page_identification_regex$metrics_by_company_and_type

metrics_plot_regex_page_identification <- metrics %>% bind_rows() %>%
  pivot_longer(
    cols = -c(package, method, classification_type),
    names_to = "metric",
    values_to = "value"
  ) %>%
  filter(metric %in% c(
    # "acc", 
    "precision", "recall", "F1")) %>%
  ggplot() +
  geom_jitter(aes(x = method, y = value, color = package), alpha = 0.5, width = 0.2, height = 0) +
  facet_grid(metric~classification_type) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  theme(
    legend.position = "bottom"
  ) +
  coord_cartesian(ylim = c(0, 1))

# creating combi table
df_list <- list()

for (type in c("Aktiva", "Passiva", "GuV")) {
  df_temp <- metric_summaries[type][[1]] %>%
    mutate_if(
        is.numeric, 
        ~ifelse(
            . == max(., na.rm = TRUE),
            paste0("**", format(round(., 3), nsmall=3), "**"),
            format(round(., 3), nsmall=3)
        )
    ) %>% pivot_wider(names_from = stat, values_from = c(precision, recall, F1)) %>% mutate(
        precision = paste(precision_mean, "±", precision_sd),
        recall = paste(recall_mean, "±", recall_sd),
        F1 = paste(F1_mean, "±", F1_sd),
    ) %>% 
    select(-ends_with("_mean")) %>% select(-ends_with("_sd")) %>% mutate_all(~str_remove(., " ± NA")) %>% 
  arrange(desc(method)) %>% mutate(.before = 2, type = type)
  # rownames(df_temp) <- df_temp$method
  
  df_list[[type]] <- df_temp
}

df_combined <- bind_rows(df_list) %>% bold_value_in_table()
```

The first approach presented in this section is, to use a key word list
and \acrfull{regex} to filter out irrelevant pages. It is setting the
performance baseline for the following approaches. Building a sound
regular expression often is an iterative process. In a first approach a
very *simple regex* was implemented. To increase the recall to 1.0 the
regular expression was extended[^05_results-4]. This second regex is
called *exhaustive regex*. In a third attempt minor changes have been
made to the *exhaustive regex* to increase the precision without
decreasing the recall. This regular expression is called *exhausitve
regex restricted*. The regular expressions can be found in the appendix
(see section \@ref(regex-page-identification-code)).

[^05_results-4]: The idea is that the regular expression approach is
    computationally cheap. If we can rely on the fact, that it keeps all
    relevant pages we can use additional, computationally more expensive
    approaches to further refine the page range.

Table \@ref(tab:display-metrics-regex-page-identification-all) shows the
mean performance for precision, recall and F1 for the three regular
expressions for the three types of pages to identify[^05_results-5]. It
was possible to create a regular expression that has a high recall for
all target types. The precision is low for all tested regular
expressions and target types. Figure
\@ref(fig:display-metrics-plot-regex-page-identification-details) gives
insight into performance differences between the companies. There is
only one document from *Berlin Energie und Netzholding* where the
**GuV** is not identified except with the *exhausitve regex
restricted*[^05_results-6].

[^05_results-5]: See Figure
    \@ref(fig:display-metrics-plot-regex-page-identification) for a
    graphical representation.

[^05_results-6]: I don't understand why the restricted version is
    finding the page but the non-restricted regex is not.

The regular expressions have been tested on the texts extracted with
multiple Python libraries. The reported standard deviations are very
small. This means that there are no substantial differences in the
extracted texts on a word level[^05_results-7]. But table
\@ref(tab:display-data-text-extraction) in section
\@ref(text-extraction-benchmark) shows that there are differences in the
extraction speed.

[^05_results-7]: Since the results are not depending on the text
    extraction library, the *exhaustive regex restricted* ran only with
    the text extracted by the fastest extraction library: *pdfium*. This
    library is used for the most tasks in this thesis. Later faced
    issues with the text extracted by *pdfium* are discussed in \@ref().

Code can be found at
"benchmark_jobs/page_identification/page_identification_benchmark_regex.ipynb"

Todo: \* look into details where they differ and if it is because of a
line break or whitespace?

```{r display-metrics-regex-page-identification-all, echo=echo_flag, warning=warning_flag, message=message_flag, results=ifelse(knitr::is_html_output(), "asis", 'markup')}
if (knitr::is_html_output()) {
  # Set caption
  cat("<table>",paste0("<caption>", "(#tab:", opts_current$get("label"), ")", "Comparing page identification metrics for different regular expressions for each classification task by type of the target table.", "</caption>"),"</table>", sep ="\n")

  df_combined %>% datatable(
    escape = FALSE, options = list(
      pageLength = 10, scrollX = TRUE, dom = "t", rowGroup = list(dataSrc=c(2)), ordering=FALSE, 
      columnDefs = list(list(visible=FALSE, targets=c(2)))
    ), extensions = 'RowGroup'
  )
} else if (knitr::is_latex_output()) {
  df_combined %>% kbl(caption = "Comparing page identification metrics for different regular expressions for each classification task by type of the target table.", escape = FALSE) %>%
    pack_rows("Aktiva", 1, 3, label_row_css = "background-color: #666; color: #fff;") %>%
    pack_rows("Passiva", 4, 6, label_row_css = "background-color: #666; color: #fff;") %>%
    pack_rows("GuV", 7, 9, label_row_css = "background-color: #666; color: #fff;")
}
```

```{r display-metrics-plot-regex-page-identification-details, echo=echo_flag, warning=warning_flag, message=message_flag, figwidth=8, fig.height=8, out.width="100%", fig.cap="Comparing the performance among different companies."}
metrics_by_company_and_type %>% 
    select(company, n_files, method, precision, recall, classification_type) %>%
    pivot_longer(cols=c(precision, recall), names_to = "metric") %>% 
    ggplot() +
    geom_boxplot(aes(x = company, y = value, fill = n_files), alpha = 0.5) +
    geom_jitter(aes(x = company, y = value, color = method), alpha = 1, height = 0) +
    facet_grid(classification_type~metric) +
    scale_x_discrete(guide = guide_axis(angle = 30)) +
  theme(
    legend.position = "bottom"
  )
```

### Table of Contents understanding {#toc-understanding}

```{r page-identification-toc-data-loading, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum(c("scripts/page_identification_toc_data_loading.R", "../benchmark_truth/toc_data.json", "../Python/pdf_texts.json"))}
source("scripts/page_identification_toc_data_loading.R")
```

The second approach presented in this section leverages the \acr{TOC}
understanding capabilities of \acr{LLM}s. @li_extracting_2023 use this
approach with long documents as a first step to determine a page range
of interest. If the predicted page range is correct and narrow, this
approach is more efficient than processing the whole document with a
\acr{LLM} directly. The \acr{TOC} in a \acr{PDF} document can be
embedded in a standardized, machine readable format or be presented in
varying, human readable forms of text on any page. Of course there are
documents without any \acr{TOC}.

Thus, the task is investigated based on two different input data formats
In one case the \acr{LLM} is provided with text extracted from the
beginning of the document. In the other case the \acr{LLM} is provided
with the Markdown formatted version of the machine readable \acr{TOC}
embedded in the document. Subsection \@ref(text-based-toc-understanding)
shows the results for the text based approach. Subsection
\@ref(code-based-toc-understanding) shows the results for the approach,
using the embedded \acr{TOC}.

Additionally, each approach is performed three times with minor changes
in the prompt. The prompts used for both approaches can be found at
\@ref(toc-understanding-promts). The prompt was adjusted two times to
tackle shortcomings in the results. The first change adds the
information, that assets and liabilities are part of the balance sheet.
It is the balance sheet, that is listed in the \acr{TOC} - not the
assets or liabilities itself. The second change specifies the
information, that assets and liabilities are often on separated pages,
into, liabilities often are found on the page after the assets.

The code can be found in:

-   "benchmark_jobs/page_identification/toc_extraction_mistral.ipynb"

-   "benchmark_jobs/page_identification/toc_extraction_qwen.ipynb"

Discussion:

-   @li_extracting_2023 did not report any issues with this approach.
    They use few-shot learning and Chain-of-Thought techniques to help
    the \acr{LLM} to understand the task. They ask just for one
    information at a time.
-   ChatGPT 4 vs Mistral 2410 8B (huge parameter difference)
-   For a lot of short annual reports one can find the tables of
    interest within the first eight pages as well.

#### Details for the approaches {#text-based-toc-understanding}

##### Text based

@li_extracting_2023 used the \acr{TOC} to identify the pages of
interest. In their approach the table of contents is extracted from the
text. Based on their observation, that the \acr{TOC} in \acr{ACFR}s is
found within the initial 165 lines of the converted document
[@li_extracting_2023, p. 20], they use the first 200 lines of text.

My initial expectation was to find the \acr{TOC} within the first five
pages. Often there are way less than 200 lines of text on the five first
pages (see Figure \@ref(fig:page-identification-toc-histogram)). In my
approach the first step is to prompt the \acr{LLM} to identify and
extract the \acr{TOC} in a given text extract\^[The prompt can be found
in section \@ref(toc-understanding-promts)]. For the same documents
Mistral 2410 8B finds\^[The strings extracted in this step have not been
checked in detail.]

-   `r n_found_toc_5_pages` strings that should represent a table of
    contents among the first five pages.

-   `r n_found_toc_200_lines` strings that should represent a table of
    contents among the first 200 lines.

```{r page-identification-toc-histogram, echo=echo_flag, warning=warning_flag, message=message_flag, fig.width=8, fig.height=3, out.width="80%", fig.cap="Histogram of the number of lines in the first 5 pages of the annual reports"}
df_num_lines %>% 
  filter(filepath %in% df_targets_no_ocr$filepath) %>% # filter documents that would need ocr preprocessing
  ggplot() +
  geom_histogram(aes(x = num_lines), bins = 20) +
  labs(x = "Number of lines in the first 5 pages", y = "Count") +
  coord_cartesian(xlim = c(0, max(df_num_lines$num_lines)))
```

##### Machine readable TOC based {#code-based-toc-understanding}

I also tested to use the \acr{TOC} representation embedded within the
PDF files. First, this limits the text amount to process. Second, this
hopefully increases the quality of the data passed to the \acr{LLM}.
`r n_toc` of the `r n_toc+n_no_toc` annual reports have a machine
readable embedded \acr{TOC}. The embedded \acr{TOC} is converted into
markdown format before it gets passed to the \acr{LLM}. Here is an
example:

```{r page-identification-markdown-code, echo=FALSE, warning=warning_flag, message=message_flag}
cat(toc_data$files_with_toc$markdown_toc[[1]])
```

#### Results

##### Comparison of the different approaches: base prompt

Table \@ref(tab:page-identification-toc-analysis-table) shows that the
machine readable \acr{TOC} approach has the highest rate of correct page
ranges for all types with the base prompt. It also predicts the most
correct page ranges in absolute numbers for **Aktiva** and **GuV**.
Thus, it also has the highest rate of correct page ranges based on the
total number of page ranges to identify over all documents - no matter,
if there was a \acr{TOC} of any type in the document or not - for
**Aktiva** and **GuV** of around `r perc_correct_total_base` %.

```{r page-identification-toc-analysis-table, echo=FALSE, warning=warning_flag, message=message_flag, results='asis'}
df_toc_benchmark %>% group_by(benchmark_type, type) %>% 
  reframe(n_correct = sum(in_range), n = n()) %>% 
  left_join(data_unnested %>% group_by(type) %>% summarise(n_total = n())) %>% 
  mutate(
    perc_correct = n_correct/n*100,
    perc_correct_total = n_correct/n_total*100
    ) %>% 
  group_by(type) %>% 
  mutate_at(
    vars(perc_correct, n_correct, perc_correct_total), 
    ~ifelse(
      . == max(., na.rm = TRUE),
      paste0("**", format_floats(.), "**"),
      format_floats(.)
    )
  ) %>% 
  render_table(alignment="llrrrrr", caption="Comparing the number and percentage of correct identified page ranges among the approaches.", ref = opts_current$get("label"), dom="t")
```

Figure \@ref(fig:page-identification-toc-analysis) shows that the amount
of correct predicted page ranges for **Passiva** is lowest for all
approaches but can be improved by simply extending the predicted end
page number by one the most. This improvement would be best for the
machine readable \acr{TOC} approach. This approach is the only one,
where the number of correct page ranges **Aktiva** would not increase if
we extend its range by one. Table
\@ref(tab:page-identification-toc-analysis-same-end-page) shows that
this is the case, because the machine readable \acr{TOC} approach
predicts the same end page for **Passiva** as for **Aktiva** in
`r perc_equal_end_page_mr` % of the cases, even though the prompt for
all approaches included the information, that **Aktiva** and **Passiva**
are on separate pages.

```{r page-identification-toc-analysis, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparing number of found TOC and amount of correct and incorrect predicted page ranges"}
df_toc_benchmark %>% ggplot() +
  geom_bar(aes(x = type, fill = in_range, colour = min_distance <= 1)) +
  geom_text(
    data = . %>% filter(in_range == TRUE),
    aes(x = type, label = paste0(round(perc_correct, 2), "")),
    stat = "count",
    vjust = 1.2,
    color = "white"
  ) +
  geom_text(
    aes(x = type, label = paste0(round(1-perc_correct, 2), "")),
    stat = "count",
    vjust = 1.5,
    color = "white"
  ) +
  facet_wrap(~benchmark_type, nrow = 1)
```

```{r page-identification-toc-analysis-same-end-page, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
df_toc_benchmark %>%
  filter(type != "GuV") %>%
  group_by(benchmark_type, filepath) %>%
  summarise(all_equal = n_distinct(end_page) == 1) %>% 
  group_by(benchmark_type) %>% 
  reframe(equal_end_page = sum(all_equal), n = n()) %>% 
  mutate(perc_equal_end_page = round(equal_end_page/n*100,1)) %>% 
  render_table(alignment="lrrr", caption="Comparing the number and percentage end pages prediction for Aktiva and Passiva that are equal.", ref = opts_current$get("label"), dom="t")
```

##### Comparison of the different approaches: advanved prompts

As a first attempt, to increase the correct page range rate for
**Passiva** I tried to specify, that assets and liabilities are part of
the balance sheet. This did work for the text based approaches, but not
for the machine readable approach (see Figure
\@ref(fig:page-identification-toc-analysis-balanced)). Figure
\@ref(fig:page-identification-toc-analysis-next-page) shows that it is
more successful, to explicit tell the \acr{LLM} that the liabilities
table is often on the page, after the assets table.

```{r page-identification-toc-analysis-next-page, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparing number of fount TOC and amount of correct and incorrect predicted page ranges"}
next_page_df_toc_benchmark %>% ggplot() +
  geom_bar(aes(x = type, fill = in_range, colour = min_distance <= 1)) +
  geom_text(
    data = . %>% filter(in_range == TRUE),
    aes(x = type, label = paste0(round(perc_correct, 2), "")),
    stat = "count",
    vjust = 1.2,
    color = "white"
  ) +
  geom_text(
    aes(x = type, label = paste0(round(1-perc_correct, 2), "")),
    stat = "count",
    vjust = 1.5,
    color = "white"
  ) +
  facet_wrap(~benchmark_type, nrow = 1) # +
  # theme(
  #   legend.position = "bottom"
  # )
```

Table \@ref(tab:page-identification-toc-analysis-table-next-page) shows
the results from the final zero shot prompt. The machine readable
\acr{TOC} approach is now predicting best for all types. Nevertheless, a
correct page range prediction rate below `r worst_finaLcorrect_rate_toc`
% is still unsufficient to build downstream task on withou human
checkups. Table \@ref(tab:page-identification-toc-gpu-time-comparison)
shows, that the machine readable \acr{TOC} approach is the fastest as
well.

```{r page-identification-toc-analysis-table-next-page, echo=FALSE, warning=warning_flag, message=message_flag, results='asis'}
next_page_df_toc_benchmark %>% group_by(benchmark_type, type) %>% 
  reframe(n_correct = sum(in_range), n = n()) %>% 
  left_join(data_unnested %>% group_by(type) %>% summarise(n_total = n())) %>% 
  mutate(
    perc_correct = n_correct/n*100,
    perc_correct_total = n_correct/n_total*100
    ) %>% 
  group_by(type) %>% 
  mutate_at(
    vars(perc_correct, n_correct, perc_correct_total), 
    ~ifelse(
      . == max(., na.rm = TRUE),
      paste0("**", format_floats(.), "**"),
      format_floats(.)
    )
  ) %>% 
  render_table(alignment="llrrrrr", caption="Comparing the number and percentage of correct identified page ranges among the approaches.", ref = opts_current$get("label"), dom="t")
```

```{r page-identification-toc-gpu-time-comparison, echo=echo_flag, warning=warning_flag, message=message_flag, results="asis"}
gpu_time_per_document_page_range %>% mutate_if(
  is.numeric, 
  ~ifelse(
    . == min(., na.rm = TRUE),
    paste0("**", ., "**"),
    .
  )
) %>% 
  render_table(alignment="lrr", caption="Comparing GPU time for page range prediction and table of contents extraction. Time in seconds per text processed.", ref = opts_current$get("label"))
```

Table \@ref(tab:page-range-prediction-performance-table) shows, that
this advantage of the machine readable \acr{TOC} approach is not coming
from wide predicted page ranges. It has the smallest median range size
among all approaches. Figure
\@ref(fig:page-identification-toc-range-plot) shows, that especially the
ranges for **GuV** are not normally distributed. Some far off lying
range sizes are shifting the mean off from the median.

```{r page-range-prediction-performance-table, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
mean_ranges %>%
  mutate_at(
    vars(median_range),
    ~ifelse(
      . == min(., na.rm = TRUE),
      paste0("**", ., "**"),
      .
    )
  ) %>%
  render_table(alignment="lrrrr", caption="Comparing the mean and median page range sizes.", ref = opts_current$get("label"))
```

```{r page-identification-toc-range-plot, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparint the predicted page range sizes. The red vertical line shows the mean and the green one shows the median of these sizes.", fig.height=4}
next_page_df_toc_benchmark %>% 
  # mutate(range = range-1) %>% 
  ggplot() +
  geom_histogram(aes(x = range), binwidth = 1) +
  geom_vline(data = mean_ranges, aes(xintercept = mean_range), color = "red") +
  geom_vline(data = mean_ranges, aes(xintercept = median_range), color = "green") +
  facet_grid(benchmark_type~type)
```

Figure \@ref(fig:page-identification-toc-range-logprobs) shows that the
confidence of the \acr{LLM}s responses is higher for the machine
readable \acr{TOC} approach as well. Besides a single group that was
predicted far off, the page ranges are closer to the correct pages too.
A linear regression of the correlation between minimal page distance and
logistic probability shows that is has a similar slope for all
approaches and target types.

```{r page-identification-toc-range-logprobs, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing the minimal distance of the predicted page range to the actual page number overthe logprobs of the models response confidence.", fig.height=4}
next_page_df_toc_benchmark %>% group_by(filepath, benchmark_type) %>%
  distance_confidence_plot() +
  facet_grid(type~benchmark_type) +
  theme(
    legend.position = "bottom"
  )
```

##### Machine readable TOC approach specific results

Figure \@ref(fig:page-identification-toc-mr-degration) shows, that
correct predictions for the page range are more probable when the
embedded \acr{TOC} has a medium number of entries. It is possible to
drop documents with less than `r min_n_entries_max_correct` without
loosing a single correct prediction. This means that the \acr{LLM} was
not able to make a correct prediction for documents with \acr{TOC}, that
have less then `r min_n_entries_max_correct` entries. This is not
surprising since neither **Bilanz** nor **GuV** are mentioned there
explicit.

It has no big influence on the predictions, if the \acr{TOC} is passed
formatted as markdown or json. With the json formatted \acr{TOC} it
found two more correct page ranges[^05_results-8]. This was tested
because the relation between heading and value for the column
*page_number* might have been clearer[^05_results-9] in json for a
one-dimensional working \acr{LLM}.

[^05_results-8]: This result is based on a single test run.

[^05_results-9]: With json the key *page_number* gets repeated every
    line, while it is just mentiones once in the beginning of the
    markdown formatted tables.

```{r page-identification-toc-mr-degration, echo=echo_flag, warning=warning_flag, message=message_flag, fig.cap="Showing the amount of correct and incorrect predicted page ranges (bars) and the percentage of correct predictions (black line)."}
df_toc_benchmark_mr_degration_next_page %>% 
  ggplot() +
  geom_col(aes(x = n_entries, y = value, fill = correct)) +
  geom_line(aes(x = as.numeric(n_entries), y = 100*perc_correct, group = 1)) +
  scale_x_discrete(guide = guide_axis(angle = 90)) +
  facet_wrap(~type, ncol = 1)
```

Delete or place somewhere else?:

-   Thus it is safer to go with the 200 lines approach. But it also
    takes longer. \@ref(tab:page-identification-toc-gpu-time-comparison)

-   Values can be higher than 80, the total number of PDF files, since
    there can be multiple tables of interested for the same type in a
    single document or a table of interest can span two pages.

### Classification with LLMs {#llm-page-identification}

```{r page-identification-llm-data-loading, echo=echo_flag, warning=warning_flag, message=message_flag}
temp_list <- readRDS("data_storage/page_identification_llm.rds")
df_binary <- temp_list$df_binary
df_multi <- temp_list$df_multi

method_families <- c("zero_shot", "law_context", "top_n_rag_examples", "n_random_examples", 'n_rag_examples')

method_familiy_colors <- c(
  "zero_shot" = "#e41a1c", 
  "law_context" = "#377eb8", 
  "top_n_rag_examples" = "#4daf4a", 
  "n_random_examples" = "#984ea3", 
  'n_rag_examples' = "#ff7f00"
  )

model_by_size_classification <- c('google_gemma-3-4b-it-0-9-1', 'google_gemma-3n-E4B-it-0-9-1', "google_gemma-3-12b-it-0-9-1",
  "google_gemma-3-27b-it-0-9-1", "meta-llama_Llama-3.1-8B-Instruct", 
  "meta-llama_Llama-3.1-70B-Instruct", "meta-llama_Llama-3.3-70B-Instruct",
  "meta-llama_Llama-4-Scout-17B-16E-Instruct", "meta-llama_Llama-4-Maverick-17B-128E-Instruct-FP8",
  "mistralai_Ministral-8B-Instruct-2410", "mistralai_Mistral-Small-3.1-24B-Instruct-2503",
  "mistralai_Mistral-Large-Instruct-2411", "Qwen_Qwen2.5-0.5B-Instruct",
  "Qwen_Qwen2.5-1.5B-Instruct", "Qwen_Qwen2.5-3B-Instruct", "Qwen_Qwen2.5-7B-Instruct",
  "Qwen_Qwen2.5-14B-Instruct", "Qwen_Qwen2.5-32B-Instruct", "Qwen_Qwen2.5-72B-Instruct",
  "Qwen_Qwen3-8B", "Qwen_Qwen3-30B-A3B-Instruct-2507", "Qwen_Qwen3-32B", "Qwen_Qwen3-235B-A22B-Instruct-2507",
  "tiiuae_Falcon3-10B-Instruct", "microsoft_phi-4"
  ) %>% gsub("^[^_]+_", "", .)

df_binary <- df_binary %>% 
  mutate(
  n_examples = as.numeric(n_examples),
  n_examples = if_else(method_family == "zero_shot", 0, n_examples),
  n_examples = if_else(method_family == "law_context", 1, n_examples),
  method_family = factor(method_family, levels = method_families)
) %>% mutate(model = gsub("^[^_]+_", "", model)) %>% 
  filter(model %in% model_by_size_classification) %>% 
  ungroup() # %>% 
  # mutate(
  #   model = str_replace(model, "_vllm", ""),
  #   model_family = sub("_.*", "", model),
  #   model_family = if_else(str_detect(model, "Qwen2"), "Qwen 2.5", model_family),
  #   model_family = if_else(str_detect(model, "Qwen3"), "Qwen 3", model_family),
  #   model_family = if_else(str_detect(model, "Llama-3"), "Llama-3", model_family),
  #   model_family = if_else(str_detect(model, "Llama-4"), "Llama-4", model_family)
  # )

binary_task <- list()
binary_task$n_models <- df_binary$model %>% unique() %>% length()
binary_task$n_model_families <- df_binary$model_family %>% unique() %>% length()
binary_task$n_method_families <- df_binary$method_family %>% unique() %>% length()

top_performer_binary <- df_binary %>% 
  filter(is.finite(f1_score), loop == 0) %>% 
  filter(n_examples <= 3 | is.na(n_examples)) %>%
  group_by(model_family, classification_type) %>% 
  slice_max(n = 1, f1_score) %>% 
  arrange(desc(f1_score)) %>% 
  select(model_family, model, classification_type, method_family, n_examples, f1_score, norm_runtime) %>%
  mutate(
    f1_score = round(f1_score, 2),
    norm_runtime = round(norm_runtime, 0),
  ) %>% rename(
    "runtime in s" = norm_runtime,
  )

top_performer_binary_median <- top_performer_binary %>% group_by(classification_type) %>% summarise(median = median(f1_score))
top_performer_binary_median_guv <- top_performer_binary_median %>% filter(classification_type == "GuV") %>% pull(median)
top_performer_binary_median_aktiva <- top_performer_binary_median %>% filter(classification_type == "Aktiva") %>% pull(median)
top_performer_binary_median_passiva <- top_performer_binary_median %>% filter(classification_type == "Passiva") %>% pull(median)

df_multi <- df_multi %>% 
  mutate(
  n_examples = as.numeric(n_examples),
  n_examples = if_else(method_family == "zero_shot", 0, n_examples),
  n_examples = if_else(method_family == "law_context", 1, n_examples),
  method_family = factor(method_family, levels = method_families)
) %>% mutate(model = gsub("^[^_]+_", "", model)) %>% 
  filter(model %in% model_by_size_classification) %>% 
  ungroup() #%>%
  # mutate(
  #   model = str_replace(model, "_vllm", ""),
  #   model_family = sub("_.*", "", model),
  #   model_family = if_else(str_detect(model, "Qwen2"), "Qwen 2.5", model_family),
  #   model_family = if_else(str_detect(model, "Qwen3"), "Qwen 3", model_family),
  #   model_family = if_else(str_detect(model, "Llama-3"), "Llama-3", model_family),
  #   model_family = if_else(str_detect(model, "Llama-4"), "Llama-4", model_family)
  # )

top_performer_multu <- df_multi %>% 
  unnest(metrics) %>% 
  filter(metric_type %in% c("Aktiva", "Passiva", "GuV")) %>% 
  filter(is.finite(f1_score), loop == 0) %>% 
  filter(n_examples <= 3 | is.na(n_examples)) %>%
  group_by(model_family, metric_type) %>% 
  slice_max(n = 1, f1_score) %>% 
  arrange(desc(f1_score)) %>% # head(10) %>% 
  select(model_family, model, metric_type, method_family, n_examples, f1_score, norm_runtime) %>%
  mutate(
    f1_score = round(f1_score, 2),
    norm_runtime = round(norm_runtime, 0),
  ) %>% rename(
    "runtime in s" = norm_runtime,
  )
```

The third approach we present in this section, uses pretrained
\acr{LLM}s to classify, if a given text extract is including any of the
target tables. Two classification approaches are presented.

On the one hand, a binary classification is used three times, to
predict, if the text extract is including an **Aktiva**, **Passiva** or
**GuV** table, once at a time. In this case the \acr{LLM} is forced to
answer with either *yes* or *no*. On the other hand, the \acr{LLM}
performed a \\acr{mcc]. For the \acr{mcc} the \acr{LLM} is forced to
answer *Aktiva*, *Passiva*, *GuV* or *other*. The prompts can be found
in appendix in section \@ref(classification-prompts).

The different classification tasks are combined with different prompting
strategies. A zero shot approach is setting the baseline. In a second
approach the excerpt of the relevant law is provided with the context.
Additionally, three few shot approaches are used.

In the few shot approaches text examples and a correct classification
for the text examples are provided. Figure
\@ref(fig:comparing-actual-number-of-in-context-examples-plot) shows how
many examples the \acr{LLM} gets provided, depending on the
classification type and chosen parameter *n_example*[^05_results-10].
For both approaches three example selection strategies are implemented.
First, random examples for each page type get sampled from the truth
dataset. Second, a vector database provides the entries that are closest
to the target text for each page type. Third, the vector database just
provides the texts that are closest to the target text without
considering the page type of the examples returned.

[^05_results-10]: See also Table
    \@ref(tab:comparing-actual-number-of-in-context-examples).

For the binary classification task the \acr{LLM} is provided with more
examples for the target type than for other types. Thus, the number of
examples and tokens is reduced. This should reduce the runtime as well.
On the same time the \acr{LLM} should get enough information about the
structure and contents of the target class and some information how it
differs from other big tables or general text pages.

For the \acr{mcc} the same amount of every possible class is provided.
Thus, the relation between the parameter *n_examples* and the number of
tokens to process is stronger for the *n_random_examples* and
*n_rag_examples* strategies.

```{r comparing-actual-number-of-in-context-examples-plot, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparing the actual number of provided examples depending on the classification type, example selection strategy and chosen parameter n-examples. The slope for the top-n-rag-examples strategy is the same for both approaches. The line for the strategies n-random-examples and n-rag-examples is equal within each approach."}
df_classification_n_examples <- read_csv("../benchmark_truth/real_example_count.csv")

df_classification_n_examples %>% 
  rename(strategy = approach) %>% 
  ggplot() +
  geom_point(aes(x = n_example, y = sum, color = strategy), alpha = .5) +
  geom_line(data = . %>% filter(strategy != "n_random_examples"), aes(x = n_example, y = sum, color = strategy)) +
  geom_line(data = . %>% filter(strategy == "n_random_examples"), aes(x = n_example, y = sum, color = strategy), linetype = "dashed") +
  facet_grid(~classification)
```

Table \@ref(tab:page-identification-llm-overview) shows, which models
have been used in the classification benchmarks. Overall
`r binary_task$n_models` models from `r binary_task$n_model_families`
model families have been tested. Prerequirement for a model to be tested
is, that it can be used with the \acr{vllm} library, accessed via
hugging face and fits into the combined VRAM of 8 nvidia B200 graphic
cards (1.536 TB). The models cover a wide range of (active) parameter
sizes. Especially for the Qwen family many models of different parameter
sizes are used in the benchmark, to investigate if there is a clear
minimum amount of parameters needed, to solve the classification task.

The results of the benchmarks have been logged as \acr{json} files
totaling in 2.1 GB of data for the final results.

```{r page-identification-llm-overview, echo=FALSE, warning=warning_flag, message=message_flag, results='asis'}
df_binary %>% 
  mutate(
   model = str_replace(model, "_vllm", ""),
   model_family = sub("_.*", "", model),
   model_family = if_else(str_detect(model, "Qwen2"), "Qwen 2.5", model_family),
   model_family = if_else(str_detect(model, "Qwen3"), "Qwen 3", model_family),
   model_family = if_else(str_detect(model, "Llama-3"), "Llama-3", model_family),
   model_family = if_else(str_detect(model, "Llama-4"), "Llama-4", model_family)
  ) %>% filter(model %in% model_by_size_classification) %>% select(model_family, model, parameter_count) %>% unique() %>% arrange(tolower(model_family), parameter_count) %>% 
  render_table(alignment="llr", caption="Overview of benchmarked LLMs for the classification tasks.", ref = opts_current$get("label"))
```

To do:

-   compare out of company vs in company rag

#### Binary classification

Table \@ref(tab:table-metrics-llm-page-identification-binary) shows the
best performing combination of model family and prompting method for
each classification target type. The classification of **GuV** tables
works best and is solved almost perfectly. The F1 score for **Aktiva**
and **Passiva** are 0.07 lower for the top performing model. The median
F1 score of **GuV** is `r top_performer_binary_median_guv` 0.1 higher
than the median F1 score for **Aktiva**
(`r top_performer_binary_median_aktiva`) and 0.2 higher than the median
F1 score for **Passiva**
(`r round(top_performer_binary_median_passiva, 2)`).

Mistral 8B Instruct 2410 is performing best for the binary
classification task for each target type. Llama-4-Scout-17B-16E-Instruct
is performing second best for **Aktiva** and **GuV** and is close to the
second best for **Passiva** as well. The runtime of Mistral 8B Instruct
2410 is four times lower than the runtime of
Llama-4-Scout-17B-16E-Instruct. In addition, the time to load
Llama-4-Scout-17B-16E-Instruct into the VRAM is much
longer[^05_results-11], because it has a total of 109B parameters. It
was surprising that Googles gemma models perform so bad[^05_results-12].

[^05_results-11]: It takes around 30 minutes to setup a vllm instance
    with Llama-4 Scout compared to 4:30 minutes setuptime for Mistral 8B
    2410.

[^05_results-12]: This is not due to a temporary technical problems
    caused by a bug in the transformers version shipped with the vllm
    0-9-2 image. Those problems have been overcome. The performance
    stays bad.

```{r table-metrics-llm-page-identification-binary, echo=echo_flag, warning=warning_flag, message=message_flag, results="asis", class.chunk="big-table"}
top_performer_binary %>% 
  render_table(alignment="llllrrr", caption="Overview of benchmarked LLMs for the binary classification tasks. Limiting the number of examples provided for the few shot approach to 3.", ref = opts_current$get("label"))
```

Figure \@ref(fig:binary-classification-result-mistral-plot) shows, the
classification performance for Mistral 8B 2410 in detail. It shows the
F1 score for each target type over the models runtime. It shows the
results for the different prompting strategies (*method_families*) with
differently colored shapes. The *zero_shot* strategy performs worst with
a F1 score below 0.6. Next come the *law_context* and
*top_n_rag_examples* strategy. Above those the *n_random_examples* and
finally the *n_rag_examples* strategy perform best.

The shape is giving information, if the example proveded to the
\acr{LLM} are selected from other companies than the target table comes
from only, or if they can also be selected from documents of the same
company. This is only relevant for strategies that get the examples
picked by the documents vector embedding distances. The \acr{LLM}
performs better[^05_results-13], if examples from documents of the same
company can be used. If this is not permitted, the *n_random_example*
approach performs better than the *n_rag_example* for the classification
of **GuV** and **Passiva** tables.

[^05_results-13]: It has a better F1 score, when examples from the same
    company are permitted. The recall is better with examples from same
    company. The precision is better without. The improvement in the
    recall is stronger.

The number inside of the shapes is referring to the *n_examples*
function parameter. Most models got benchmarked with an *n_examples*
value of up to three. The actual number of examples provided to the
models is depending on the method family / example selection strategy
and can be looked up in Table
\@ref(tab:comparing-actual-number-of-in-context-examples).

The best performing model, Mistral 8B 2410, was provided with mode
examples to investigate the effect of a richer context. The predictions
do not get better by providing more and more examples. Figure
\@ref(fig:binary-classification-result-mistral-plot) shows, that the
improvements get smaller naturally going from three to five examples
while approaching an F1 score of 1.0.

But for the *n_rag_example* strategy we find a significant drop in the
F1 score, if we set the *n_examples* to five[^05_results-14] and
examples pages come from annual reports of other companies. This is
caused by a sever recall drop. For the *n_random_example* strategy we
see a small drop with the F1 score for the class **Passiva** as well.
Taking into account that the runtime also almost is twice as high, this
is very inefficient.

[^05_results-14]: In this case five examples for the target table type
    and two examples for each other type are provided, totaling at
    twelve examples.

Figure \@ref(fig:binary-classification-result-mistral-plot) also shows,
that the results are stable[^05_results-15]. Running the benchmark three
times shows similar results in the F1 score for each strategy. This is
reflected by closely overlapping shapes of the same color with the same
number within.

[^05_results-15]: Earlier experiments on a subset of the pages have been
    run five times indicating stable results. Running the experiments up
    to tree times in this very task indicate this as well.

```{r binary-classification-result-mistral-plot, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing F1 score performance over normalized runtime for binary classification for Mistral-8B-Instruct-2410."}
df_binary %>% filter(model == "Ministral-8B-Instruct-2410", loop < 3) %>% 
  filter(n_examples <= 5 | is.na(n_examples)) %>% 
  ggplot(aes(x = norm_runtime/60, y = f1_score)) +
  geom_point(aes(color = method_family, shape = out_of_company), size = 7, alpha = .6) +
  scale_shape(na.value = 15, guide = "legend") +
  geom_text(aes(label = n_examples)) +
  scale_color_manual(values = method_familiy_colors) +
  facet_grid(model~classification_type) +
  # theme(legend.position = "bottom") +
  guides(
    color = guide_legend(ncol = 1, title.position = "top"),
    shape = guide_legend(ncol = 1, title.position = "top")
  ) +
  labs(x = "normalized runtime in minutes")
```

Figure
\@ref(fig:binary-classification-result-mistral-many-random-examples-plot)
shows the experiments for Ministral-8B-Instruct-2410 with *n_examples*
greater or equal three. This time the actual number of examples provided
to the \acr{LLM} are shown in the shapes to increase the comparability
among the different strategies. Additionally, it shows results for the
*top_n_rag_example* strategy with *n_examples* up to 13. The F1 score of
the *top_n_rag_example* strategy stays lower than the F1 score of the
*n_rag_examples* strategy, even though there are more examples used.
This is mainly caused by lower precision scores, probably because there
are no contrasting examples provided.

```{r binary-classification-result-mistral-many-random-examples-plot, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing F1 score performance over normalized runtime for binary classification for Mistral-8B-Instruct-2410. Comparing the performance based on the real number of provided examples."}
df_binary %>% left_join(df_classification_n_examples, by = c(
  "method_family" = "approach", 
  "n_examples" = "n_example", 
  "classifier_type" = "classification"
  )) %>% rename(sum_examples = sum) %>% 
  filter(model == "Ministral-8B-Instruct-2410", loop < 2) %>% 
  filter(n_examples > 1 | is.na(n_examples)) %>% 
  ggplot(aes(x = norm_runtime/60, y = f1_score)) +
  geom_point(aes(color = method_family, shape = out_of_company), size = 7, alpha = .6) +
  scale_shape(na.value = 15, guide = "legend") +
  scale_color_manual(values = method_familiy_colors) +
  geom_text(aes(label = sum_examples)) +
  facet_grid(model~classification_type) +
  # theme(legend.position = "bottom") +
  guides(
    color = guide_legend(ncol = 1, title.position = "top"),
    shape = guide_legend(ncol = 1, title.position = "top")
  ) +
  labs(x = "normalized runtime in minutes")
```

Figure \@ref(fig:binary-classification-performance-over-runtime) and
Figure
\@ref(fig:binary-classification-performance-over-runtime-full-time)
shows the F1 performance over normalized rundtime for all benchmarked
models. Comparing Mistral-8B-Instruct-2410 with
Mistral-124B-Instruct-2411 shows that one can spend over tenfold amount
of computation power without getting better results.

It also shows, that with Qwen 2.5 it needs at least the 3B parameter
model to achieve good results. Comparing the 32B and 72B parameter
models shows, that th performance does not increase anymore, but starts
to decrease. For Qwen 3 it shows, that only the newer mix of experts
models give reasonable results.

The mix of expert models show good performance for the Llama 4 family as
well and reduce the compute time compared with the 72B models of LLama
3. But for LLama 4 Maverick the performance drops using the
*n_rag_examples* strategy with three *n_examples*. The performance of
Llama 3.1 70B was higher than the performance of Llama 3.3 70B.

Summary:

Neither do newer generations always improve the performance for the
binary classification task, nor do more parameters always improve or at
least show stable performance.

##### Confidence

```{r confidence-match-plot-mistral-binary-data-prep, echo=echo_flag, warning=warning_flag, message=message_flag}
target_type <- "Aktiva"

df_filtered <- df_binary %>% filter(
  classification_type == target_type, n_examples<=3
  ) %>%
  arrange(desc(f1_score))

df_temp <- df_filtered[1,"predictions"][[1]][[1]] %>% as_tibble()

df_flipped_score <- df_temp %>% 
  mutate(
    confidence_score = if_else(predicted_type == "no", 1-confidence_score, confidence_score),
    is_aktiva = str_detect(type, target_type)
  )

model_name_best_f1_aktiva <- df_filtered[1, "model"]
method__best_f1_aktiva <- df_filtered[1, "method"]
```

I investigate the relation between the reported confidence for an answer
and its correctness, to check if it is possible to inform humans in the
loop about results they should double check and which results they can
trust. Figure \@ref(fig:confidence-match-plot-mistral-binary) shows the
distribution of reported confidence score for the binary classification
with target type **Aktiva** for all tables types grouped by their
correctness for Mistral-8B-Instruct-2410. The \acr{LLM} just returns one
prediction and its confidence[^05_results-16]. The abscissa shows
$confidence = exp(logprob)$, if the answer is *yes* and
$confidence = 1-exp(logprob)$ if the answer is *no*.

[^05_results-16]: The model could be forced to return multiple answers,
    but it was not. The confidence score is given as log probability.
    The exponential function was applied to show the results on the more
    common scale of 0 to 1.

One can see that the predictions are very accurate making just
`r sum(df_temp$match==FALSE)` mistakes for `r length(df_temp$match)`
predictions. The reported confidence for answer *yes* is showing a wide
spread from around 0.25 to 1.0. This is true for the answer *no* as
well. Most wrong decisions are made for responses that have a reported
confidence in the range from 0.25 to 0.75. But there are more correct
answers in this range as well. It never misclassifies **GuV** or
**Passiva**[^05_results-17] as **Aktiva**. But it with shows some not
recalled **Aktiva** tables and is predicting some of the pages of
majority class, with not further described content and structure, as
**Aktiva**.

[^05_results-17]: There was a single prediction where \acr{LLM} predicts
    **Aktiva** with high confidence, when the truth is **Passiva**
    instead. Because Qwen was showing the same wrong prediction for one
    **Passiva** table, I double checked the ground truth. I found, that
    the page shows **Aktiva** and **Passiva** simultaneously and was not
    correct codified. This was not the only time, where a mistake in the
    gold truth was found, by examining potential \acr{LLM} mistakes.

```{r confidence-match-plot-mistral-binary, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing the confidence score for the Aktiva classification task grouped by table type and correctness for Mistral-8B-Instruct-2410."}
df_flipped_score %>% 
  ggplot() +
  geom_boxplot(aes(x = predicted_type, y = confidence_score)) +
  geom_jitter(aes(x = predicted_type, y = confidence_score, color = match), alpha = .3, height = 0) +
  facet_wrap(~type) +
  labs(title = model_name_best_f1_aktiva,
       subtitle = method__best_f1_aktiva) +
  coord_cartesian(ylim=c(0,1))
```

```{r confidence-match-plot-qwen-binary-data-prep, echo=echo_flag, warning=warning_flag, message=message_flag}
df_filtered <- df_binary %>% filter(classification_type == "Aktiva", model_family=="Qwen") %>% 
  arrange(desc(f1_score))

df_temp <- df_filtered[1,"predictions"][[1]][[1]] %>% as_tibble()

df_flipped_score <- df_temp %>% 
  mutate(
    confidence_score = if_else(predicted_type == "no", 1-confidence_score, confidence_score),
    is_aktiva = str_detect(type, "Aktiva")
  )

model_name_best_f1_aktiva <- df_filtered[1, "model"]
method__best_f1_aktiva <- df_filtered[1, "method"]
```

This is different for models of most other model families. Figure
\@ref(fig:confidence-match-plot-qwen-binary) shows, that
`r model_name_best_f1_aktiva` returns always high confidence scores,
even when it is wrong. The model shows perfect recall but its precision
is worse than the precision of the Mistral model.

```{r confidence-match-plot-qwen-binary, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing the confidence score for the Aktiva classification task grouped by table type and correctness for Qwen-2.5-32B-Instruct."}
df_flipped_score %>% 
  ggplot() +
  geom_boxplot(aes(x = predicted_type, y = confidence_score)) +
  geom_jitter(aes(x = predicted_type, y = confidence_score, color = match), alpha = .3, height = 0) +
  facet_wrap(~type) +
  labs(title = model_name_best_f1_aktiva,
       subtitle = method__best_f1_aktiva)
```

```{r def-micro-pr-function-binary, echo=echo_flag, warning=warning_flag, message=message_flag}
library(yardstick)

calc_flipped_conf_all_types <- function(df, model_name, method_name) {
  l_temp <- list()
  # browser()
  
  for (target in c('Aktiva', 'GuV', 'Passiva')) {
    # t <- "Aktiva"
    # df_filtered <- df %>% filter(
    #   classification_type == t,
    #   n_examples <= 3,
    #   loop == 0) %>% 
    #   arrange(desc(f1_score))
    # model_name_best_f1_aktiva <- df_filtered[model_rank, "model"]
    # method_best_f1_aktiva <- df_filtered[model_rank, "method"]
    
    df_filtered <- df %>%
      filter(
        classification_type == target,
        model == model_name,
        method == method_name,
        loop == 0
      ) %>% 
      arrange(desc(f1_score))
    
    df_temp <- df_filtered[1,"predictions"][[1]][[1]] %>% as_tibble()
    
    df_flipped_score <- df_temp %>% 
      mutate(
        confidence_score = if_else(predicted_type == "no", 1-confidence_score, confidence_score),
        target = target
      )
    l_temp[target] <- list(df_flipped_score)
  }
  
  list(df = bind_rows(l_temp), model = model_name, method = method_name)
}

plot_pr_double_curve <- function(l, selected_type, x_stuff = FALSE) {
  # browser()
  df_temp2 <- l$df %>% filter(target == selected_type) %>% 
    mutate(type = if_else(type == selected_type, type, "no")) %>% 
    mutate(type = factor(type, levels = c(selected_type, "no")))
  # pr_obj <- pr.curve(scores.class0 = df_temp2$confidence_score[df_temp2$match == 1],
  #                    scores.class1 = df_temp2$confidence_score[df_temp2$match == 0],
  #                    curve = TRUE)
  
  # Precision-Recall Curve with ggplot2
  
  pr_df <- df_temp2 %>% pr_curve(type, confidence_score) %>%
    rename(threshold = .threshold) %>% filter(threshold<=1) %>% 
    mutate(f1 = 2 * precision * recall / (precision + recall))
  
  pr_auc <- round(df_temp2 %>% pr_auc(type, confidence_score) %>% .$.estimate, 3)
  best_F1_row <- pr_df %>% slice_max(n = 1, f1)  
  best_F1_row_high_recall <- pr_df %>% 
    filter(recall > 0.999, precision > 0.1) %>% slice_max(n = 1, f1)  
  best_F1 <- best_F1_row  %>% pull(f1)
  best_F1_high_recall <- best_F1_row_high_recall %>% pull(f1)
  best_threshold <- best_F1_row  %>% pull(threshold)
  best_threshold_high_recall <- best_F1_row_high_recall  %>% pull(threshold)
  best_precision_high_recall <- best_F1_row_high_recall  %>% pull(precision)
  
  g1 <- pr_df %>%
    ggplot(aes(x = recall, y = precision)) +
    geom_line(aes(color = threshold), size = 1.2) +
    scale_color_viridis_c(option = "plasma", limits = c(0, 1)) +
    labs(
      subtitle = paste0("Precision-Recall Curve for ", selected_type, " (AUC = ", pr_auc, ")"),
      x = "Recall",
      y = "Precision"
    ) +
    coord_cartesian(ylim = c(0,1)) +
    theme(
      legend.position = "bottom"
    )
  
  g2 <- pr_df %>%
    ggplot(aes(x = recall, y = precision, color = f1)) +
    geom_line(size = 1.2) +
    scale_color_viridis_c(option = "viridis", limits = c(0, 1)) +
    labs(
      # title = "Precision-Recall Curve colored by F1 score",
      x = "Recall",
      y = NULL,
      color = "F1 score"
    ) +
    coord_cartesian(ylim = c(0,1))+
    theme(
      legend.position = "bottom"
    ) 
  
  if (x_stuff == FALSE) {
    g1 <- g1 + guides(color = FALSE) +
      labs(x = element_blank())
    
    g2 <- g2 + guides(color = FALSE) +
      labs(x = element_blank())
  }
  
  combined_plot <- g1 + g2
  combined_plot + plot_annotation(caption = paste0(
    'Best F1 score of ', round(best_F1,3) , ' gets reached with threshold of value ', round(best_threshold,3), '\n',
    'Best F1 score with recall > 0.999 of ', round(best_F1_high_recall,3) , ' gets reached with threshold of value ', round(best_threshold_high_recall,3), ' (corresp. precision value: ', round(best_precision_high_recall,3) , ')'))
}
```

Figure \@ref(fig:micro-pr-curve-llm-binary-1) shows the
precision-recall-curve for the best performing model twice for each
target type. On the left plots the line color represents the threshold
score one could use to decide when to accept a response as it is. On the
right plots the line color is showing the F1 score that results with a
chosen threshold.

The \acr{AUC} value is lowest for **Aktiva**. Here the F1 score is
highest for a threshold value of 0.73. This prevents to classify the
pages of type *other* to get classified as **Aktiva**. If it is required
to have a very high recall value a threshold of 0.44 should be chosen.

The precision-recall-curve for **Passiva** is very similar but there is
a step close to the recall value of 1.0. This has the effect that for a
a guaranteed high recall a very low precision (0.24) and F1 (0.38) has
to be accepted[^05_results-18].

[^05_results-18]: Thus, a human has in average to check four pages and
    select the correct **Passiva** page among them.

The shape of the precision-recall-curve for **GuV** almost perfectly
reaches the top right corner. The highest F1 score is found with a
threshold value of .56. With a threshold value of 0.5 a very high recall
is guaranteed and the F1 score is just a little lower.

```{r micro-pr-curve-llm-binary-1, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.height=8, fig.cap="Showing the precsion-recall-curve for the best performing model."}
model <- "Ministral-8B-Instruct-2410"
method <- "3_rag_examples"

df_temp_mistral <- df_binary %>% calc_flipped_conf_all_types(model, method)
p_aktiva <- df_temp_mistral %>% plot_pr_double_curve("Aktiva", x_stuff = FALSE)
p_passiva <- df_temp_mistral %>% plot_pr_double_curve("Passiva", x_stuff = FALSE)
p_guv <- df_temp_mistral %>% plot_pr_double_curve("GuV", x_stuff = TRUE)

wrap_elements(p_aktiva) / wrap_elements(p_passiva) / wrap_elements(p_guv) +
  plot_layout(heights = c(2, 2, 3)) +
  plot_annotation(title = str_c(model, " with ", method))
```

Figure \@ref(fig:table-identification-llm-confidence) summarizes the
relation between reported confidence and correctness of the
classification for all target types and compares it among the best
performing model-strategy combinations for Ministral-8B-Instruct-2410
and Qwen3-8B. One can see, that the reported confidence for correct and
incorrect classifications are separable in most cases for Mistral-8B.
This separation is worse for Qwen3-8B and worst for target type
**Passiva**.

Figure \@ref(fig:table-identification-llm-confidence-lm) shows, that for
Ministral-8B values with a confidence of 0.7 and more, a human don't has
to double check the classification for target type **GuV**. This
interval is smallest for **Passiva** where only confidences above 0.9
can be fully trusted. These empirical intervals might shrink, once more
data is evaluated. If one is less strict and accepts misclassifiaction
rates of 1 % the found interval for **Passiva** starts at 0.8 and is
probably less depended on the sample evaluated. The percentage of
predictions that can be trusted without risk is greater than 93 % even
for target type **Passiva**.

For Qwen3-8B we find almost no range without any wrong classifications.
For **GuV** this range includes 35 % of all predictions. The ranges that
allow for 1 % of wrong classifications cover 57 % of all predictions at
least.

```{r table-identification-llm-confidence, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparing the reported confidence scores for the page identification task for the Mistral and Qwen 3 with 8B parameters.", cache=TRUE, cache.extra = tools::md5sum("data_storage/data_storage/page_identification_llm.rds"), fig.height=3}
confidence_vs_truth_binary <- df_binary %>% 
  filter(model %in% c("Ministral-8B-Instruct-2410", "Qwen3-8B")) %>% 
  group_by(method, model) %>% mutate(
    mean_f1 = mean(f1_score, na.rm=TRUE), .before = 1
  ) %>% group_by(model, classification_type) %>% 
  arrange(desc(mean_f1)) %>% 
  slice_max(mean_f1, n = 1, with_ties = FALSE) %>% 
  select(-filepath) %>% 
  unnest(predictions) %>% mutate(
    match = factor(match, levels = c(F, T)),
    # truth_NA = factor(truth_NA, levels = c(F, T))
  )

confidence_vs_truth_binary %>% ggplot() +
  geom_boxplot(
    aes(x = match, y = confidence_score, fill = classification_type), 
    position = position_dodge2(preserve = "single")) +
  scale_fill_discrete(drop = FALSE) +
  scale_x_discrete(drop = FALSE) +
  facet_nested(~ model + classification_type)
```

```{r table-identification-llm-confidence-lm, echo=echo_flag, warning=warning_flag, message=message_flag, fig.height=5, out.width="100%", fig.cap="Estimating the relative frequency to find a wrong classification over different confidence intervals"}
confidence_vs_truth_binary %>% rename(confidence = confidence_score) %>% 
  mutate(
    conf_interval = cut(confidence, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE),
    conf_center = as.numeric(sub("\\((.+),(.+)\\]", "\\1", levels(conf_interval))[conf_interval]) + 0.025
  ) %>%
  group_by(conf_center, classification_type, model) %>%
  summarise(
    n_true = sum(match == TRUE, na.rm = TRUE),
    n_false = sum(match == FALSE, na.rm = TRUE),
    total = n_true + n_false,
    chance_false = if_else(total > 0, n_false / total * 100, NA_real_),
    chance_zero = chance_false == 0,
    chance_below_1 = chance_false < 1,
    chance_low = if_else(chance_zero, 0, if_else(chance_below_1, 1, 2)),
    chance_low = factor(chance_low, levels = c(0,1,2), labels = c("equls 0 %", "below 1 %", "more"))
  ) %>% group_by(classification_type, model) %>% mutate(
    perc = total/sum(total)*100
  ) %>% ungroup() %>% 
  mutate(
    chance_false_interval = cut(
      chance_false,
      breaks = c(0, 1, 2, 4, 8, 16, 32, 64, Inf),
      labels = c("[0,1)", "[1,2)", "[2,4)", "[4,8)", 
                 "[8,16)", "[16,32)", "[32,64)", "[64,Inf)"),
      right = FALSE,
      ordered_result = TRUE
    ),
  ) %>%
  ggplot() +
  geom_col(aes(x = conf_center, y = perc, color = chance_low, fill = chance_false_interval), alpha = 1) +
  # geom_text(
  #   aes(x = conf_center, y = perc, label = round(perc, 0)), 
  #   position = position_stack(vjust = 1), vjust = -0.6, 
  #   size = 3, color = "black"
  # ) +
  scale_color_manual(values = c("#00CC00", "orange", "#555555")) +
  scale_fill_manual(values = rev(c("#d53e4f", "#f46d43", "#fdae61", "#fee08b", "#e6f598", "#abdda4", "#66c2a5", "#3288bd"))) +
  labs(
    x = "Confidence Interval Center", 
    y = "Percentage of predictions", 
    color = "mistake rate") +
  coord_cartesian(
    ylim = c(0, 100), 
    xlim = c(0,1)
  ) +
  facet_grid(classification_type ~ model)

```

Discussion:

-   Could be more efficient to predict "is any of interest" and then
    which type, because dataset is highly imbalanced.

-   Why takes n_rag_examples so much longer?

-   **Aktiva** and **Passiva** sometimes on the same page and more
    similar than **GuV**?

-   Recall = 1 for human in the loop (looking at selection of pages that
    could be target and none else, if the number of wrong pages are few
    =\> what says F1 with recall 1?)

-   Confidence range to error rate

#### Multi-class classification

Table \@ref(tab:table-metrics-llm-page-identification-multi) shows that
Llama-4-Scout solves the \acr{mcc} task almost perfect for all classes.
Mistral-Large-Instruct-2411 performs second best. In contrast to the
binary classification task no order is visible, what class was easiest
to predict. Googles gemma models perform much better in the \acr{mcc}
task with F1 scores of 0.89 instead of 0.58 for the binary
classification task.

Table
\@ref(tab:table-metrics-llm-page-identification-multi-small-models)
shows that the smaller models do perform good, too. Ministral-8B
performs good but is around tenfold faster than Mistral-Large and
Llama-4 Scout. For the larger models the *n_rag_examples* strategy is
performing best. For the smaller models the *top_n_rag_examples*
strategy is performing good as well and is faster because of shorter
contexts.

```{r table-metrics-llm-page-identification-multi, echo=echo_flag, warning=warning_flag, message=message_flag, results="asis", class.chunk="big-table"}
top_performer_multu %>% render_table(alignment="llllrrr", caption="Overview of benchmarked LLMs for the multi-class classification tasks. Limiting the number of examples provided for the few shot approach to 3.", ref = opts_current$get("label"))

```

```{r table-metrics-llm-page-identification-multi-small-models, echo=echo_flag, warning=warning_flag, message=message_flag, results="asis", class.chunk="big-table"}
df_multi %>% 
  unnest(metrics) %>% 
  filter(metric_type %in% c("Aktiva", "Passiva", "GuV")) %>% 
  filter(is.finite(f1_score), loop == 0) %>% 
  filter(parameter_count<17) %>% 
  filter(n_examples <= 3 | is.na(n_examples)) %>%
  group_by(model_family, metric_type) %>% 
  filter(f1_score == max(f1_score, na.rm = TRUE)) %>% 
  arrange(desc(f1_score)) %>% # head(10) %>% 
  select(model_family, model, metric_type, method_family, n_examples, f1_score, norm_runtime) %>%
  mutate(
    f1_score = round(f1_score, 2),
    norm_runtime = round(norm_runtime, 0),
  ) %>% rename(
    "runtime in s" = norm_runtime,
  ) %>% 
  render_table(alignment="llllrrr", caption="Overview of benchmarked LLMs for the multi-class classification tasks focussing on models with less than 17B parameters. Limiting the number of examples provided for the few shot approach to 3.", ref = opts_current$get("label"))
```

Figure
\@ref(fig:performance-over-runtime-multi-classification-mistral-vs-llama-scout)
shows, the micro averaged F1 score for the three minority classes over
the normalized runtime for two models. It shows the results for the
different prompting strategies (*method_families*) with differently
colored shapes.

The shape is giving information, if the examples provided to the
\acr{LLM}, are exclusively selected from other companies than the target
table comes from, or if they can also be selected from documents of the
same company. This is only relevant for strategies that get the examples
picked by the documents vector embedding distances (*top_n_rag_examples*
and *n_rag_examples*). The \acr{LLM} performs better, if examples from
documents of the same company can be used.

The number inside of the shapes is referring to the *n_examples*
function parameter. Most models got benchmarked with an *n_examples*
value of up to three. The actual number of examples provided to the
models is depending on the method family / example selection strategy
and can be looked up in Table
\@ref(tab:comparing-actual-number-of-in-context-examples).

*n_rag_examples* better for Llama 4 Scout than *n_random_examples* ; For
Ministral it is depending on the *out_of_company* setting

On can see, that Ministral–8B 2410 reaches a good performance already
with few examples, but only if *out_of_company* is false. It performs
moderate with the *law-context* strategy and *zero_shot*, too. Adding
more examples does not improve the performance. Best with
*top_n_rag_examples*

```{r performance-over-runtime-multi-classification-mistral-vs-llama-scout, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparing F1 score micro averaged for the minority classes for two models over their normalized runtime."}
df_selected <- df_multi %>% unnest(metrics) %>% filter(metric_type == "micro_minorities")

df_selected %>% 
  filter(model %in% c(
    "Ministral-8B-Instruct-2410",
    # "Mistral-Large-Instruct-2411",
    # "Mistral-Small-3.1-24B-Instruct-2503",
    "Llama-4-Scout-17B-16E-Instruct"
    # "Llama-4-Maverick-17B-128E-Instruct-FP8"
  ),
  loop == 0,
  n_examples <= 9
  ) %>% 
  ggplot(aes(x = norm_runtime/60, y = f1_score)) +
  geom_point(aes(color = method_family, shape = out_of_company), size = 7, alpha = .6) +
  scale_shape(na.value = 15, guide = "legend") +
  geom_text(aes(label = n_examples)) +
  facet_grid(model~metric_type) +
  scale_color_manual(values = method_familiy_colors) +
  labs(x = "normalized runtime in minutes") #+
  # theme(legend.position = "bottom") +
  # guides(
  #   color = guide_legend(ncol = 1, title.position = "top"),
  #   shape = guide_legend(ncol = 1, title.position = "top")
  # ) +
  # scale_x_discrete(guide = guide_axis(angle = 30))
```

##### Confidence

Figure \@ref(fig:confidence-match-plot-llama-multi) shows the reported
confidence scores for the predictions for the best performing
model-strategy combination, Llama 4 Scout with *3_rag_examples*. It is
confident for most correct predictions and only misclassifies some of
the pages with unknown characteristics. The target types are all
recognized correct. All confidences are greater than 0.5. Probably
because there is no case where the confidences for all possible classes
is below 0.5 and there always is a most probable class. It would have
been interesting to use the classification framework of \acr{vllm} to
get predictions for all competing classes. But this requires special
trained models with pooling capability[^05_results-19].

[^05_results-19]: It might be possible to request the n most probable
    answers to get confidence scores for all different predictions. But
    this was not investigated.

```{r confidence-match-plot-llama-multi, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing the reported confidence scores for all predictions of Llama 4 Scout grouped by the true target type. Errors have only been made within the majority class."}
df_filtered <- df_selected %>%
  arrange(desc(f1_score))
df_temp <- df_filtered[1,"predictions"][[1]][[1]] %>% as_tibble()
df_flipped_score <- df_temp %>% 
  mutate(
    confidence_score = if_else(predicted_type == "no", 1-confidence_score, confidence_score),
    is_aktiva = str_detect(type, "Aktiva")
  )
model_name_best_f1_aktiva <- df_filtered[1, "model"]
method__best_f1_aktiva <- df_filtered[1, "method"]

df_flipped_score %>% 
  ggplot() +
  geom_boxplot(aes(x = predicted_type, y = confidence_score)) +
  geom_jitter(aes(x = predicted_type, y = confidence_score, color = match), alpha = .3, height = 0) +
  facet_wrap(~type) +
  labs(title = model_name_best_f1_aktiva,
       subtitle = method__best_f1_aktiva) +
  coord_cartesian(ylim = c(0,1))
```

Figure \@ref(fig:confidence-match-plot-mistral-multi) shows the reported
confidence scores for the predictions for the best performing
model-strategy combination among the small models limited to
*n_examples* with n smaller five[^05_results-20],
Ministral-8B-Instruct-2410 with *3_rag_examples*. One can see there are
some wrong classifications for the minority classes as well. Especially,
the **Passiva** target type is often classified as *other*. This is
problematic for a smooth workflow (see discussion chapter?)

[^05_results-20]: The best performance results with
    *top_11_rag_examples* but the plot was less interesting and its F1
    score was not listed in Table
    \@ref(tab:table-metrics-llm-page-identification-multi-small-models).

```{r confidence-match-plot-mistral-multi, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing the reported confidence scores for all predictions of Ministral 8B grouped by the true target type. Errors have only been made within the majority class."}
df_filtered <- df_selected %>% filter(
  model == "Ministral-8B-Instruct-2410", 
  method == "3_rag_examples"
  ) %>% 
  arrange(desc(f1_score))
df_temp <- df_filtered[1,"predictions"][[1]][[1]] %>% as_tibble()
df_flipped_score <- df_temp %>% 
  mutate(
    confidence_score = if_else(predicted_type == "no", 1-confidence_score, confidence_score),
    is_aktiva = str_detect(type, "Aktiva")
  )
model_name_best_f1_aktiva <- df_filtered[1, "model"]
method__best_f1_aktiva <- df_filtered[1, "method"]

df_flipped_score %>% 
  ggplot() +
  geom_boxplot(aes(x = predicted_type, y = confidence_score)) +
  geom_jitter(aes(x = predicted_type, y = confidence_score, color = match), alpha = .3, height = 0) +
  facet_wrap(~type) +
  labs(title = model_name_best_f1_aktiva,
       subtitle = method__best_f1_aktiva) +
  coord_cartesian(ylim = c(0,1))
```

```{r def-micro-pr-function-multi, echo=echo_flag, warning=warning_flag, message=message_flag}
plot_pr_double_curve_multi <- function(df, selected_type, x_stuff = FALSE) {
  df_temp2 <- df %>% 
    filter(loop == 0) %>% select(-filepath) %>% unnest(predictions) %>% 
    select(type, predicted_type, confidence_score) %>% 
    mutate(
      type = if_else(type == selected_type, type, "no"),
      predicted_type = if_else(predicted_type == selected_type, predicted_type, "no")
      ) %>% 
    mutate(type = factor(type, levels = c(selected_type, "no"))) %>% 
    mutate(
      confidence_score = if_else(predicted_type == "no", 1-confidence_score, confidence_score)
    )

  
  # Precision-Recall Curve with ggplot2
  
  pr_df <- df_temp2 %>% pr_curve(type, confidence_score) %>%
    rename(threshold = .threshold) %>% filter(threshold<=1) %>% 
    mutate(f1 = 2 * precision * recall / (precision + recall))
  
  pr_auc <- round(df_temp2 %>% pr_auc(type, confidence_score) %>% .$.estimate, 3)
  best_F1_row <- pr_df %>% slice_max(n = 1, f1)  
  best_F1_row_high_recall <- pr_df %>% 
    filter(recall > 0.999, precision > 0.1) %>% slice_max(n = 1, f1)  
  best_F1 <- best_F1_row  %>% pull(f1)
  best_F1_high_recall <- best_F1_row_high_recall %>% pull(f1)
  best_threshold <- best_F1_row  %>% pull(threshold)
  best_threshold_high_recall <- best_F1_row_high_recall  %>% pull(threshold)
  best_precision_high_recall <- best_F1_row_high_recall  %>% pull(precision)
  
  g1 <- pr_df %>%
    ggplot(aes(x = recall, y = precision)) +
    geom_line(aes(color = threshold), size = 1.2) +
    scale_color_viridis_c(option = "plasma", limits = c(0, 1)) +
    labs(
      subtitle = paste0("Precision-Recall Curve for ", selected_type, " (AUC = ", pr_auc, ")"),
      x = "Recall",
      y = "Precision"
    ) +
    coord_cartesian(ylim = c(0,1)) +
    theme(
      legend.position = "bottom"
    )
  
  g2 <- pr_df %>%
    ggplot(aes(x = recall, y = precision, color = f1)) +
    geom_line(size = 1.2) +
    scale_color_viridis_c(option = "viridis", limits = c(0, 1)) +
    labs(
      # title = "Precision-Recall Curve colored by F1 score",
      x = "Recall",
      y = NULL,
      color = "F1 score"
    ) +
    coord_cartesian(ylim = c(0,1))+
    theme(
      legend.position = "bottom"
    ) 
  
  if (x_stuff == FALSE) {
    g1 <- g1 + guides(color = FALSE) +
      labs(x = element_blank())
    
    g2 <- g2 + guides(color = FALSE) +
      labs(x = element_blank())
  }
  
  combined_plot <- g1 + g2
  combined_plot + plot_annotation(caption = paste0(
    'Best F1 score of ', round(best_F1,3) , ' gets reached with threshold of value ', round(best_threshold,3), '\n',
    'Best F1 score with recall > 0.999 of ', round(best_F1_high_recall,3) , ' gets reached with threshold of value ', round(best_threshold_high_recall,3), ' (corresp. precision value: ', round(best_precision_high_recall,3) , ')'))
}
```

Figure \@ref(fig:micro-pr-curve-llm-multi-ministral) shows the
precision-recall-curve for Ministral-8B-Instruct-2410 with
*3_rag_examples* twice for each target type. On the left plots the line
color represents the threshold score one could use to decide when to
accept a response as it is. On the right plots the line color is showing
the F1 score that results with a chosen threshold.

The \acr{AUC} is highest for **GuV** again. But for the multi-class
classification **Passiva** shows the lowest \acr{AUC}, not **Aktiva** as
it was in the binary classification task. The precision-recall-curve for
**Aktiva** and \*\*Passiva\*+ show a "step" in the area of high recall.
This has a strong effect on the threshold one should choose, if one
wants to guarantee a high recall. The corresponding precision values of
0.2 and 0.13 mean that a human has to check five to eight pages in
average to get a correct classified page of type **Aktiva** and
**Passiva**.

The corresponding plot for the best performing model,
Llama-4-Scout-17B-16E-Instruct, can be found in Figure
\@ref(fig:micro-pr-curve-llm-multi-llama-scout). Here the
precision-recall-curve for **Passiva** and **GuV** is almost perfect.
Just the single prediction for **Aktiva** wit a lower confidence shows
an influence on the precision-recall-curve.

```{r micro-pr-curve-llm-multi-ministral, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.height=8, fig.cap="Showing the precsion-recall-curve for Ministral-8B-Instruct-2410."}
model_name <- "Ministral-8B-Instruct-2410"
method_name <- "3_rag_examples"

df_temp_mistral_multi <- df_multi %>% 
  filter(
    model == model_name,
    method == method_name
  ) # %>% unnest(metrics) # %>% calc_micro_f1(model, method)
p_aktiva <- df_temp_mistral_multi %>% plot_pr_double_curve_multi("Aktiva", x_stuff = FALSE)
p_passiva <- df_temp_mistral_multi %>% plot_pr_double_curve_multi("Passiva", x_stuff = FALSE)
p_guv <- df_temp_mistral_multi %>% plot_pr_double_curve_multi("GuV", x_stuff = TRUE)
# p_other <- df_temp_mistral_multi %>% plot_pr_double_curve_multi("other", x_stuff = TRUE)

wrap_elements(p_aktiva) / wrap_elements(p_passiva) / wrap_elements(p_guv) +
  plot_layout(heights = c(2, 2, 3)) +
  plot_annotation(title = str_c(model, " with ", method))
```

Figure \@ref(fig:table-identification-llm-confidence-multi) summarizes
the relation between reported confidence and correctness of the
classification for all target types and compares it among the best
performing model-strategy combinations for
Llama-4-Scout-17B-16E-Instruct, Ministral-8B-Instruct-2410 and Qwen3-8B.
It seems, as the reported confidence for correct and incorrect
classifications are separable in most cases for Mistral-8B. For Llama 4
Scout this seems not true for the target type **GuV**. For Qwen3-8B
there is almost no separation at all.

Figure \@ref(fig:table-identification-llm-confidence-lm-multi) shows,
that there is almost no area, where the empirical rate of wrong
classifications is zero[^05_results-21]. Only for Ministral-8B we find
intervals, where a human don't has to double check the classification
for target types **Aktiva** and **GuV**. These intervals include 90 % of
all predictions. If error rates of 1 % are accepted almost all
predictions by Llama Scout 4 and about 96 % of the predictions by
Ministral-8B are included in the corresponding intervals. For Qwen3-8B
we find no interval without an error rate below 1 %.

[^05_results-21]: The size of intervals has been narrowed down to 0.1 %
    and still there was no range without wrong classification for Llama
    4 Scout.

```{r table-identification-llm-confidence-multi, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparing the reported confidence scores for the multi-class page identification task for the Mistral and Qwen 3 with 8B parameters.", cache=TRUE, cache.extra = tools::md5sum("data_storage/data_storage/page_identification_llm.rds"), fig.height=3}
confidence_vs_truth_multi <- df_multi %>% 
  filter(model %in% c("Ministral-8B-Instruct-2410", "Llama-4-Scout-17B-16E-Instruct", "Qwen3-8B")) %>% 
  unnest(metrics) %>% 
  group_by(method, model, metric_type) %>% mutate(
    mean_f1 = mean(f1_score, na.rm=TRUE), .before = 1
  ) %>% group_by(model, metric_type) %>% 
  arrange(desc(mean_f1)) %>% 
  slice_max(mean_f1, n = 1, with_ties = FALSE) %>% 
  select(-filepath) %>% 
  unnest(predictions) %>% mutate(
    match = factor(match, levels = c(F, T)),
    # truth_NA = factor(truth_NA, levels = c(F, T))
  ) %>% filter(metric_type %in% c("Aktiva", "Passiva", "GuV"))

confidence_vs_truth_multi %>% ggplot() +
  geom_boxplot(
    aes(x = match, y = confidence_score, fill = metric_type), 
    position = position_dodge2(preserve = "single")) +
  scale_fill_discrete(drop = FALSE) +
  scale_x_discrete(drop = FALSE) +
  facet_grid(~ model)
```

```{r table-identification-llm-confidence-lm-multi, echo=echo_flag, warning=warning_flag, message=message_flag, fig.height=5, out.width="100%", fig.cap="Estimating the relative frequency to find a wrong classification over different confidence intervals for the multi-class classification task."}
confidence_vs_truth_multi %>% rename(confidence = confidence_score) %>% 
  mutate(
    conf_interval = cut(confidence, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE),
    conf_center = as.numeric(sub("\\((.+),(.+)\\]", "\\1", levels(conf_interval))[conf_interval]) + 0.025
  ) %>%
  group_by(conf_center, metric_type, model) %>%
  summarise(
    n_true = sum(match == TRUE, na.rm = TRUE),
    n_false = sum(match == FALSE, na.rm = TRUE),
    total = n_true + n_false,
    chance_false = if_else(total > 0, n_false / total * 100, NA_real_),
    chance_zero = chance_false == 0,
    chance_below_1 = chance_false < 1,
    chance_low = if_else(chance_zero, 0, if_else(chance_below_1, 1, 2)),
    chance_low = factor(chance_low, levels = c(0,1,2), labels = c("equls 0 %", "below 1 %", "more"))
  ) %>% group_by(metric_type, model) %>% mutate(
    perc = total/sum(total)*100
  ) %>% ungroup() %>% 
  mutate(
    chance_false_interval = cut(
      chance_false,
      breaks = c(0, 1, 2, 4, 8, 16, 32, 64, Inf),
      labels = c("[0,1)", "[1,2)", "[2,4)", "[4,8)", 
                 "[8,16)", "[16,32)", "[32,64)", "[64,Inf)"),
      right = FALSE,
      ordered_result = TRUE
    ),
  ) %>%
  ggplot() +
  geom_col(aes(x = conf_center, y = perc, color = chance_low, fill = chance_false_interval), alpha = 1) +
  geom_text(
    aes(x = conf_center, y = perc, label = round(perc, 0)), 
    position = position_stack(vjust = 1), vjust = -0.6, 
    size = 3, color = "black"
  ) +
  scale_color_manual(values = c("#00CC00", "orange", "#555555")) +
  scale_fill_manual(values = rev(c("#d53e4f", "#f46d43", "#fdae61", "#fee08b", "#e6f598", "#abdda4", "#66c2a5", "#3288bd"))) +
  labs(
    x = "Confidence Interval Center", 
    y = "Percentage of predictions", 
    color = "mistake rate") +
  coord_cartesian(
    ylim = c(0, 100), 
    xlim = c(0,1)
  ) +
  facet_grid(metric_type ~ model)

```

### Term frequency based classifier {#tf-classifier}

The fourth approach uses term frequencies for a key word list and the
number of floats to rank pages. The approach is inspired by
\acr{TF-IDF} - a technique commonly used for information retrieval. It
is similar to the baseline approach, because it uses a key word list and
regular expressions to count terms and floats. But it is more flexible
because the works in the key word list are not mandatory. This makes the
approach robust against issues in the text extracts for single key
words.

The key word list is generated removing the stop words from the law
about **Aktiva**, **Passiva** and **GuV**. The key words from the
\acr{regex} approach are added, e.g. *GuV* and *Gewinn- und
Verlustrechnung*. Since real life representations of those target types
never contain all entries, it is not possible to include most of those
words in a strict \acr{regex} search as presented in the first approach.

This approach sums the counts of each word from the key word list per
page in a first variable. In a second variable it counts the number of
floats on each page. These two variables are then divided by the number
of words found on the page. These densities are used to rank all pages
from a single document. This is done with a unique key word list for
each target type.

A random forest is trained to determine which density should be weighted
to what amount. Because of the imbalanced data set undersampling is used
when the training data set is created[^05_results-22]. A single random
forest is trained because the density of floats and specific words is
assumed to be similar. The actual type of the page is not taken into
account. The model just knows if the page is a page of any target type,
based on the term and float density. This trippels the data points of
the target class.

[^05_results-22]: The random forest build with undersampling performs
    much better as a classifier, that is trained using n oversamples
    train dataset.

This single random forest performs much better than random forests that
are trained using the dataset for each target type separatly. The
performance is tested on all data points not included in the
undersampled train dataset. Thus the test dataset is again highly
imbalanced.

The random forest performs a binary classification task. But instead of
the actual classifications, the predicted scores are used to rank the
pages. Instead of precision or recall the metric used for the evaluation
is top k recall. It is of interest which value of k is required to get a
recall of 100 %.

The code can be found at:
"benchmark_jobs/page_identification/term_frequency.ipynb"

```{python random-forest-2-predictors, include=TRUE, eval=FALSE, echo=echo_flag, warning=warning_flag, message=message_flag}
import pandas as pd
import numpy as np
from sklearn.ensemble import RandomForestClassifier
from sklearn.metrics import accuracy_score
from sklearn.model_selection import train_test_split

df_train_us = pd.read_csv("../benchmark_results/page_identification/term_frequency_table.csv")

# Drop rows without ground truth
# df_train_us = df_word_counts.merge(df_truth, on=["filepath", "type"], how="left")
df_train_us["is_truth"] = (df_train_us["page"] == df_train_us["page_truth"]).astype(int)
df_train_us = df_train_us.dropna(subset=["page_truth"])

# Undersample the majority class (is_truth == 0)
df_true = df_train_us[df_train_us["is_truth"] == 1]
df_false = df_train_us[df_train_us["is_truth"] == 0]
df_false_undersampled = df_false.sample(n=len(df_true), random_state=42)
df_train_us_balanced = pd.concat([df_true, df_false_undersampled]).sample(frac=1, random_state=42).reset_index(drop=True)
# df_train_us_balanced

# Features and target
X = df_train_us_balanced[["term_frequency", "float_frequency"]].values
y = df_train_us_balanced["is_truth"].values

# Train-test split (70% train, 30% test)
X_train, X_test, y_train, y_test, df_train_split, df_test_split = train_test_split(
    X, y, df_train_us_balanced, test_size=0.3, random_state=42, stratify=y
)

# Train Random Forest model
clf = RandomForestClassifier(n_estimators=100, random_state=42)
clf.fit(X_train, y_train)
score = clf.score(X_train, y_train)
# print(f"Training accuracy: {score:.2%}")
score = clf.score(X_test, y_test)
# print(f"Test accuracy: {score:.2%}")

# Predict and rerank: get predicted probabilities for each page
df_train_split["score"] = clf.predict_proba(X_train)[:, 1]
df_test_split["score"] = clf.predict_proba(X_test)[:, 1]

# Add all not-chosen negatives from df_false to test split
df_false_unused = df_false.loc[~df_false.index.isin(df_false_undersampled.index)]
df_false_unused = df_false_unused.copy()
df_false_unused["score"] = clf.predict_proba(df_false_unused[["term_frequency", "float_frequency"]].values)[:, 1]
df_false_unused["rank"] = np.nan  # Not ranked yet

# Concatenate with test split
df_test_split = pd.concat([df_test_split, df_false_unused], ignore_index=True)

# For each group (filepath, type), sort by score descending
df_train_split["rank"] = df_train_split.groupby(["filepath", "type"])["score"].rank(ascending=False, method="first")
df_test_split["rank"] = df_test_split.groupby(["filepath", "type"])["score"].rank(ascending=False, method="first")
```

-   top 1
-   top k

low precision llm linked to position of correct page? numeric frequency?

Figure \@ref(fig:forest-map-tf-2-predictors) shows how the test data
points are distributed in the two dimensional value map for the random
forest with two predictors. The target pages have a *float_frequncy*
between 0.2 and 0.5 and pages with a *term_frequency* value over 0.07
get classified as target. One target page shows a lower *term_frequency*
and thus does not get ranked correct. (recall, precision?)

```{python forest-map-tf-2-predictors, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum('data_storage/rf-tf-2preds.sav'), results='hold', fig.cap="Classification map showing which score a data point gets based on its term and float frequency and which type the data points in the test dataset actually have."}
import shap
import pickle
import pandas as pd

import matplotlib.pyplot as plt
from sklearn.inspection import DecisionBoundaryDisplay
from matplotlib.colors import ListedColormap

predictors = [
    "term_frequency", 
    "float_frequency", 
    "date_count", 
    "integer_count"
]

clf = pickle.load(open('data_storage/rf-tf-2preds.sav', 'rb'))
df_test_split = pd.read_csv("data_storage/term_frequency_results_2_predictors_test.csv")


X = df_test_split[["term_frequency", "float_frequency"]].values

cm = plt.cm.RdBu
cm_bright = ListedColormap(["#FF0000", "#0000FF"])
x_min, x_max = X[:, 0].min() - 0.05, X[:, 0].max() + 0.05
y_min, y_max = X[:, 1].min() - 0.05, X[:, 1].max() + 0.05

plt.figure(figsize=(8, 6))
ax = plt.gca()
catch_this_message =DecisionBoundaryDisplay.from_estimator(
    clf, X, cmap=cm, alpha=0.8, ax=ax, eps=0.05
)

df_test_split_sorted = df_test_split.sort_values('is_truth')
X_test = df_test_split_sorted[["term_frequency", "float_frequency"]].values
y_test = df_test_split_sorted["is_truth"].values
score = clf.score(X_test, y_test)

# ax.scatter(X_train[:, 0], X_train[:, 1], c=y_train, cmap=cm_bright, edgecolors="k", label="Train")
ax.scatter(X_test[:, 0], X_test[:, 1], c=y_test, cmap=cm_bright, alpha=0.6, edgecolors="k", label="Test")
catch_this_message =ax.set_xlim(x_min, x_max)
catch_this_message =ax.set_ylim(y_min, y_max)
ax.set_xlabel("term_frequency")
ax.set_ylabel("float_frequency")
ax.set_title(f"RandomForestClassifier (accuracy={score:.2f})")
plt.legend()
plt.show()
```

A second random forest is trained supplementing the two predictors *term
density* and *float density* with two additional predictors: *date
count* and *integer count*. Figure
\@ref(fig:top-n-recall-term-frequency) shows the top n recall for both
random forests. On the left side the top n recall on the imbalanced test
dataset is shown. On the right side the performance on the train
dataset.

Both random forests perform similar on the train dataset. The random
forest with four predictors reaches perfect recall faster for **Aktiva**
on the test dataset. Thus, with $n = 5$ 100 % recall is reached for the
random forest with four predictors. With the random forest with two
predictors it needs $n = 7$.

```{r top-n-recall-term-frequency, echo=echo_flag, warning=warning_flag, message=message_flag, fig.cap="Comparing the top n recall on training and test dataset among the random forest with two and four predictors."}
df_2_predictors_test <- read_csv("/home/simon/Documents/data_science/Thesis/benchmark_results/page_identification/term_frequency_results_2_predictors_test.csv") %>% 
  mutate(data_split = 'test', n_predictors = 2)
df_2_predictors_train <- read_csv("/home/simon/Documents/data_science/Thesis/benchmark_results/page_identification/term_frequency_results_2_predictors_train.csv") %>% 
  mutate(data_split = 'train', n_predictors = 2)
df_4_predictors_test <- read_csv("/home/simon/Documents/data_science/Thesis/benchmark_results/page_identification/term_frequency_results_4_predictors_test.csv") %>% 
  mutate(data_split = 'test', n_predictors = 4)
df_4_predictors_train <- read_csv("/home/simon/Documents/data_science/Thesis/benchmark_results/page_identification/term_frequency_results_4_predictors_train.csv") %>% 
  mutate(data_split = 'train', n_predictors = 4)

df_rf_results <- bind_rows(
  df_2_predictors_train, df_2_predictors_test,
  df_4_predictors_train, df_4_predictors_test
  )

max_rank = df_rf_results %>% filter(is_truth == 1) %>% pull(rank) %>% max()
results <- map_dfr(1:max_rank, function(i_rank) {
  df_rf_results %>% 
    filter(is_truth == 1) %>% 
    group_by(type, data_split, n_predictors) %>% 
    mutate(le = if_else(rank <= i_rank, 1, 0)) %>% 
    summarise(mean = mean(le), .groups = "drop") %>% 
    mutate(i_rank = i_rank)
})

results %>% ggplot() +
  geom_col(aes(x = i_rank, y = mean)) +
  facet_nested(type ~ data_split + n_predictors) +
  labs(
    x = "rank",
    y = "top n recall",
    # title = "Top n recall for different ranks, data splits and number of predictors"
  )

```

```{python random-forest-4-predictors, include=TRUE, eval=FALSE, echo=echo_flag, warning=warning_flag, message=message_flag}
import pandas as pd
import pickle
import numpy as np
from sklearn.ensemble import RandomForestClassifier
from sklearn.metrics import accuracy_score
from sklearn.model_selection import train_test_split

df_train_us = pd.read_csv("../benchmark_results/page_identification/term_frequency_table.csv")

# Drop rows without ground truth
# df_train_us = df_word_counts.merge(df_truth, on=["filepath", "type"], how="left")
df_train_us["is_truth"] = (df_train_us["page"] == df_train_us["page_truth"]).astype(int)
df_train_us = df_train_us.dropna(subset=["page_truth"])

# Undersample the majority class (is_truth == 0)
df_true = df_train_us[df_train_us["is_truth"] == 1]
df_false = df_train_us[df_train_us["is_truth"] == 0]
df_false_undersampled = df_false.sample(n=len(df_true), random_state=42)
df_train_us_balanced = pd.concat([df_true, df_false_undersampled]).sample(frac=1, random_state=42).reset_index(drop=True)
# df_train_us_balanced

predictors = [
    "term_frequency", 
    "float_frequency", 
    "date_count", 
    "integer_count"
]

# Features and target
X = df_train_us_balanced[predictors].values # only better with date and integer counts; otherwise worse
y = df_train_us_balanced["is_truth"].values

# Train-test split (70% train, 30% test)
X_train, X_test, y_train, y_test, df_train_split, df_test_split = train_test_split(
    X, y, df_train_us_balanced, test_size=0.3, random_state=42, stratify=y
)

# Train Random Forest model
clf = RandomForestClassifier(n_estimators=100, random_state=42)
catch_this_message =clf.fit(X_train, y_train)

# Predict and rerank: get predicted probabilities for each page
df_train_split["score"] = clf.predict_proba(X_train)[:, 1]
df_test_split["score"] = clf.predict_proba(X_test)[:, 1]

# Add all not-chosen negatives from df_false to test split
df_false_unused = df_false.loc[~df_false.index.isin(df_false_undersampled.index)]
df_false_unused = df_false_unused.copy()
df_false_unused["score"] = clf.predict_proba(df_false_unused[predictors].values)[:, 1]
df_false_unused["rank"] = np.nan  # Not ranked yet

# Concatenate with test split
df_test_split = pd.concat([df_test_split, df_false_unused], ignore_index=True)

# For each group (filepath, type), sort by score descending
df_train_split["rank"] = df_train_split.groupby(["filepath", "type"])["score"].rank(ascending=False, method="first")
df_test_split["rank"] = df_test_split.groupby(["filepath", "type"])["score"].rank(ascending=False, method="first")
```

Figure \@ref(fig:shap-tf-4-predictors) shows that the two additional
predictors *date_count* and *integer_count* have little importance. But
since it is computationally cheap to determine their value and the
effiency of a random forest classifier, there is little reason not to
use them.

```{python shap-tf-4-predictors, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum('data_storage/rf-tf-preds.sav'), fig.cap="Beeswarm plot of SHAP importance values for the four predictors of the second random forest classifier."}
import shap
import pickle
import pandas as pd

predictors = [
    "term_frequency", 
    "float_frequency", 
    "date_count", 
    "integer_count"
]

clf = pickle.load(open('data_storage/rf-tf-4preds.sav', 'rb'))
df_test_split = pd.read_csv("data_storage/term_frequency_results_4_predictors_test.csv")

# Use the same predictors and trained RandomForestClassifier (clf) as in cell 20
explainer = shap.TreeExplainer(clf)
shap_values = explainer.shap_values(df_test_split[predictors].values)

# Plot summary for class 1 (is_truth == 1)
shap.summary_plot(shap_values[:,:,1], df_test_split[predictors].values, feature_names=predictors)
```

Fianlly, figure \@ref(fig:micro-pr-curve-tf-4-predictors) shows the
precision-recall-curves for the term frequency approach for all three
target types. The \acr{AUC} for all types is below 0.5. The precision
and F1 score stay below 0.5 as well. A high recall can be maintained for
all types for threshold values up to at least 0.72.

```{r micro-pr-curve-tf-4-predictors, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing the precsion-recall-curve for the random forest with four predictors.", fig.height=8}
plot_pr_double_curve_tf <- function(df, selected_type, x_stuff = FALSE) {
  df_temp2 <- df %>% filter(
    n_predictors == 4, data_split == "test",
    type == selected_type
  ) %>% 
    rename(
      confidence_score = score
    ) %>% 
    mutate(
      type = factor(is_truth, levels = c(1, 0))
    )
  
  # Precision-Recall Curve with ggplot2
  
  pr_df <- df_temp2 %>% pr_curve(type, confidence_score) %>%
    rename(threshold = .threshold) %>% filter(threshold<=1) %>% 
    mutate(f1 = 2 * precision * recall / (precision + recall))
  
  pr_auc <- round(df_temp2 %>% pr_auc(type, confidence_score) %>% .$.estimate, 3)
  best_F1_row <- pr_df %>% slice_max(n = 1, f1)  
  best_F1_row_high_recall <- pr_df %>% 
    filter(recall > 0.999, precision > 0.1) %>% slice_max(n = 1, f1)  
  best_F1 <- best_F1_row  %>% pull(f1)
  best_F1_high_recall <- best_F1_row_high_recall %>% pull(f1)
  best_threshold <- best_F1_row  %>% pull(threshold)
  best_threshold_high_recall <- best_F1_row_high_recall  %>% pull(threshold)
  best_precision_high_recall <- best_F1_row_high_recall  %>% pull(precision)
  
  g1 <- pr_df %>%
    ggplot(aes(x = recall, y = precision)) +
    geom_line(aes(color = threshold), size = 1.2) +
    scale_color_viridis_c(option = "plasma", limits = c(0, 1)) +
    labs(
      subtitle = paste0("Precision-Recall Curve for ", selected_type, " (AUC = ", pr_auc, ")"),
      x = "Recall",
      y = "Precision"
    ) +
    coord_cartesian(ylim = c(0,1), xlim = c(0,1)) +
    theme(
      legend.position = "bottom"
    )
  
  g2 <- pr_df %>%
    ggplot(aes(x = recall, y = precision, color = f1)) +
    geom_line(size = 1.2) +
    scale_color_viridis_c(option = "viridis", limits = c(0, 1)) +
    labs(
      # title = "Precision-Recall Curve colored by F1 score",
      x = "Recall",
      y = NULL,
      color = "F1 score"
    ) +
    coord_cartesian(ylim = c(0,1), xlim = c(0,1))+
    theme(
      legend.position = "bottom"
    ) 
  
  if (x_stuff == FALSE) {
    g1 <- g1 + guides(color = FALSE) +
      labs(x = element_blank())
    
    g2 <- g2 + guides(color = FALSE) +
      labs(x = element_blank())
  }
  
  combined_plot <- g1 + g2
  combined_plot + plot_annotation(caption = paste0(
    'Best F1 score of ', round(best_F1,3) , ' gets reached with threshold of value ', round(best_threshold,3), '\n',
    'Best F1 score with recall > 0.999 of ', round(best_F1_high_recall,3) , ' gets reached with threshold of value ', round(best_threshold_high_recall,3), ' (corresp. precision value: ', round(best_precision_high_recall,3) , ')'))
}

df_temp <- df_rf_results

p_aktiva <- df_temp %>% plot_pr_double_curve_tf("Aktiva", x_stuff = FALSE)
p_passiva <- df_temp %>% plot_pr_double_curve_tf("Passiva", x_stuff = FALSE)
p_guv <- df_temp %>% plot_pr_double_curve_tf("GuV", x_stuff = TRUE)
# p_other <- df_temp_mistral_multi %>% plot_pr_double_curve_multi("other", x_stuff = TRUE)

wrap_elements(p_aktiva) / wrap_elements(p_passiva) / wrap_elements(p_guv) +
  plot_layout(heights = c(2, 2, 3)) +
  plot_annotation(title = str_c(model, " with ", method))
```

### Comparison {#comparison-page-identification}

```{r load-page-identification-comapison-data, echo=echo_flag, warning=warning_flag, message=message_flag}
df_page_identification_comparing_performance <- readODS::read_ods("data_storage/page_identification_summary.ods", sheet = 1)
df_page_identification_comparing_efficiency <- readODS::read_ods("data_storage/page_identification_summary.ods", sheet = 2)

df_df_page_identification_comparing_performance_formatted <- df_page_identification_comparing_performance %>% 
  mutate(
    across(
      c(precision, recall, F1, `top 1 recall`),
      ~ifelse(      
        . == max(., na.rm = TRUE),
        paste0("**", round(., 3), "**"),
        round(., 3)
        )
      )) %>% 
  mutate(`k for full recall` = ifelse(      
        `k for full recall` == min(`k for full recall`, na.rm = TRUE),
        paste0("**", round(`k for full recall`, 3), "**"),
        round(`k for full recall`, 3)
        ))
```

In this subsection we summarize the performnce and efficiency for all
four presented approaches and compare it with the the results a human
may achieve with manual labour.

##### Prediction performance

Table \@ref(tab:page-identification-performance-overview-all) shows the
best performance of achieved by the four presented approaches. The best
F1 score is reached by Llama 4 Scout for the target types **Akiva** and
**Passiva** in the multi-class classification approach. For **GuV** the
best F1 score (0.985) is found with Ministral-8B-Instruct in the binary
classification approach. Llama 4 Scout reaches a F1 score of 0.971 for
target type **GuV** and multi-class classification.

In the dataset preparation for the table extraction task (see section
\@ref(table-extraction-introduction) 107 **Aktiva** pages have been
selected. In this manual process we made two mistakes, accidently
selecting one **Passiva** and one **GuV** page. Thus the human baseline
to compete with is 0.981. Thus, Llama 4 Scout is more precise than us.

Furthermore, Llama 4 Scout reached a recall of 1.0 for all target types.
This means, the results can be used downstream, even though the
precision is not always perfect. The pages classified as target can be
double checked by a human, without missing any page.

The other approaches' performance is way worse. Only the term-frequency
approach's results could be used downstream, because we find a recall of
1.0. Table
\@ref(tab:page-identification-performance-overview-top-k-recall) shows
the results of the top k recall for the term-frequency and \acr{LLM}
approaches. The \acr{LLM}s always rate the correct **GuV** page highest.
With Llama Scout 4 we find all target pages within the first two ranked
pages. For the term-frequency approach a human sometimes has to check up
to five pages.

```{r page-identification-performance-overview-all, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
 df_df_page_identification_comparing_performance_formatted %>% 
   select(-c(`top 1 recall`, `k for full recall`)) %>% 
  render_table(
    alignment = "lllrrr",
    caption = "Comparing page identification perfromance among all four approaches.",
    ref = opts_current$get("label"))
```

```{r page-identification-performance-overview-top-k-recall, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
 df_df_page_identification_comparing_performance_formatted %>% 
   select(-c(precision, recall, F1)) %>% 
   filter(!is.na(`top 1 recall`)) %>% 
  render_table(
    alignment = "lllrr",
    caption = "Comparing the top k recall for the term-frequency and LLM approaches.",
    ref = opts_current$get("label"))
```

##### Energy usage and runtime

Table \@ref(tab:page-identification-efficiency-overview-all) shows the
runtime in seconds per document, estimated energy consumption in Joule
per document and costs in CENTS **per 1000 documents**. The runtime for
the \acr{LLM}s was normalized to seconds on a nvidia B200 and thus the
TDP of 700 W is used to calculate the energy consumption. For the other
approaches, running on my laptop (see section \@ref(local-machine)) a
TDP of 28 Watts is used. For manual work by a human additional 60 W are
added for the screen used. It is assumed that the \acr{LLM} is hosted
locally.

Table \@ref(tab:page-identification-efficiency-overview-all) shows, that
the regular expression approach is fastest and consumes least energy.
Nevertheless, since the results are not sufficient another approach has
to be chosen if the amount of manual labor should be reduced for the
human inn the loop.

Second place regarding all these criteria is the term-frequency
approach, which guarantees a perfect recall, while reducing the number
of pages to investigate to five per target type. This is similar to the
number of pages a human has to investigate to find the \acr{TOC} of the
document. And it is a reduction to 7.4 % of the average 67 pages the
documents in this dataset have. The costs are still negligible.

The \acr{LLM} approaches have the highest runtime and energy
consumption. This is the case, because they process every page with very
computational demanding algorithms. The fastest and least energy
consuming strategy is to use a small model as Ministral-8B-Instruct for
the multi-class approach. This is more effective than running three
binary classifications. An alternative approach could be to binary
predict if the page is of any target type and then perform a
classification, which type exactly the page is of. But the results of
the multi-class strategy are good enough as well. In both strategies the
k required for perfect recall is three, using the Ministral-8B-Instruct
model[^05_results-23].

[^05_results-23]: Potentially smaller fine tuned models can solve the
    task even more efficient.

For the \acr{TOC} approach \acr{LLM}s are used as well, but they process
far less of the documents pages. This can be achieved for the good
performing classification strategies by combining the term-frequency
approach with the \acr{LLM} approach.

The manual approach is the slowest. For the benchmark ten random
documents were processed by us. We used the \acr{TOC} and the search
function to find key words like **Aktiva** or **Bilanz**. Anyhow, its
almost as fast as the multi-classification using Llama 4 Scout. But it
requires eight times less energy. Comparing it to Ministral-8B-Instruct
it take three times longer but consumes less then half of the energy.

Not taken into account fo this comparison are factors like

-   costs to buy and maintain hardware (i.e. a GPU cluster).

-   higher costs per runtime if the \acr{LLM} compute is purchased from
    cloud providers. CLOUD: price if LLM is in the cloud \<-- print
    tokens used

-   payment and insurance to pay for a human (e.g . student coworker).

-   the training time and energy consumption for training either

    -   a \acr{LLM} (probably done by the \acr{LLM} provider).

    -   a human (growing up, getting educated).

-   the energy consumed with the food humans eat.

Since all approaches but the manual identification need the text
extract, this runtime and energy consumption are also not listed (but
low).

```{r page-identification-efficiency-overview-all, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
df_page_identification_comparing_efficiency %>% filter(!drop == 1) %>% 
  select(approach, strategy, `runtime per document in s`, `energy in J`, `costs in CENTS per 1000 documents`) %>%
  mutate(
    across(
      where(is.numeric),
      ~ifelse(      
        . == min(., na.rm = TRUE),
        paste0("**", round(., 3), "**"),
        round(., 3)
        )
      )) %>% 
  render_table(
    alignment = "llrrr",
    caption = "Comparing page identification efficiency among all four approaches.",
    ref = opts_current$get("label"))
```

## Table extraction {#table-extraction-introduction}

\ChapFrame

The second research question asks, how \acr{LLM}s can be used, to
effectively extract specific information from a financial report.

Which tasks have there been? Which models have been used for which task?
What data has been used?

### Baseline: Regex

```{r table-extraction-regex-data-loading, echo=FALSE, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum('data_storage/table_extraction_regex.rds')}
df_table_extraction_regex <- readRDS("data_storage/table_extraction_regex.rds")

real_table_extraction_regex_total_performance_mean <- df_table_extraction_regex %>% filter(table_type == "real_tables") %>% pull(percentage_correct_total) %>% mean()
real_table_extraction_regex_num_performance_mean <- df_table_extraction_regex %>% filter(table_type == "real_tables") %>% pull(percentage_correct_numeric) %>% mean(na.rm = TRUE)
real_table_extraction_regex_NA_F1_mean <- df_table_extraction_regex %>% filter(table_type == "real_tables") %>% pull(NA_F1) %>% mean()

synth_table_extraction_regex_total_performance_mean <- df_table_extraction_regex %>% filter(table_type == "synth_tables", extraction_backend == "pymupdf") %>% pull(percentage_correct_total) %>% mean()
synth_table_extraction_regex_num_performance_mean <- df_table_extraction_regex %>% filter(table_type == "synth_tables", extraction_backend == "pymupdf") %>% pull(percentage_correct_numeric) %>% mean(na.rm = TRUE)
synth_table_extraction_regex_NA_F1_mean <- df_table_extraction_regex %>% filter(table_type == "synth_tables", extraction_backend == "pymupdf") %>% pull(NA_F1) %>% mean(na.rm = TRUE)
```

The baseline for the table extraction task is set by an approach using
regular expressions on the text extract. The approach performs much
better[^05_results-24] on the synthetic dataset compared to the real
dataset (see Figure \@ref(fig:table-extraction-regex-performance)). Even
though, it does not perform perfectly and its performance is more
consistent on the text extracted with pymupdf compared to pdfium. Some
possible explanations are:

[^05_results-24]: A comparison of the numeric values over all methods
    can be found in section \@ref(comparing-table-extraction-methods).

-   a duplicated row name[^05_results-25]
-   numeric columns extracted separated from row names by extraction
    libraries
-   sums in the same row as the single values[^05_results-26]
-   with pdfium: missing white space[^05_results-27]
-   with pdfium: random line breaks[^05_results-28]

[^05_results-25]: The row *Geleistete Anzahlungen* can be found in two
    parts of the table and the simple approach just matches the numbers
    to the first found entry.

[^05_results-26]: In this case the \acr{regex} takes the sum as the
    value for the previous year.

[^05_results-27]: This can form unexpected numeric patterns or prevent
    the row names to be recognized.

[^05_results-28]: The approach takes care of line breaks between words,
    but not within. This leads to unrecognized row names as well.

You can find some examples for incorrect extracted texts in section
\@ref(regex-extraction-mistakes).

On the real dataset the approach shows a wider spread for the percentage
of correct extracted numeric values as well as a considerable number of
annual reports where the extraction did not work at all. Interestingly,
the used text extraction library has no noticeable influence on the real
dataset.

```{r table-extraction-regex-performance, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Performance overall and on numeric value extraction with regular expressions. Showing single scores for *percentage correct numeric* on real tables to explain wide boxes."}
df_table_extraction_regex %>%
  select(c(table_type, percentage_correct_numeric, percentage_correct_total, T_EUR, extraction_backend)) %>% 
  pivot_longer(cols = -c(table_type, T_EUR, extraction_backend)) %>% 
  ggplot() +
  geom_point(
    data= . %>% filter(table_type == "real_tables", name == "percentage_correct_numeric"),
    aes(x = table_type, y = value, alpha = extraction_backend), color = "#264DEB",
    shape = 4, position=position_jitterdodge(dodge.width=0.9, jitter.width = 0.2, jitter.height = 0.005)
  ) +
  geom_boxplot(aes(x = table_type, y = value, fill = extraction_backend), alpha = .3) +
  # geom_jitter(data= . %>% filter(table_type == "real_tables"), aes(x = table_type, y = value, color = T_EUR), alpha = .8, height = 0, shape = 4) +
  # facet_wrap(~name, ncol = 1) +
  scale_alpha_manual(values = c(1, 1), guide = "none") +
  scale_fill_manual(values = c("#94EB1F", "orange")) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_grid(~name)
```

The random line breaks result in some missed row names which is
reflected by the bigger spread for NA precision with pdfium on the
synthetic dataset (see Figure \@ref(fig:table-extraction-regex-NA)).
Nevertheless, the NA precision for the majority of the cases is perfect.
This is different with the real dataset. The NA precision is found to be
at only
`r df_table_extraction_regex %>% filter(table_type == "real_tables") %>% pull(NA_precision) %>% mean() %>% round(2)`.

```{r table-extraction-regex-NA, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Performance on classification for missing values with regular expressions"}
# NAs for F1 are valid if there is not thinning
df_table_extraction_regex %>% select(c(table_type, NA_precision, NA_recall, NA_F1, T_EUR, extraction_backend)) %>% 
  pivot_longer(cols = -c(table_type, T_EUR, extraction_backend)) %>% 
  ggplot() +
  geom_boxplot(aes(x = table_type, y = value, fill = extraction_backend), alpha = .3) +
  scale_alpha_manual(values = c(1, 1)) +
scale_fill_manual(values = c("#94EB1F", "orange")) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_grid(~name)
```

##### Hypotheses

The formulated hypotheses have been evaluated visually using the
dependence and beeswarm plots from the shapviz library based on the
\acr{SHAP} values calculated with a random forest.

###### Real dataset

```{r real-table-extraction-regex-shap-loading, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", cache=TRUE, cache.extra = tools::md5sum('data_storage/h2o/real_table_extraction_regex_h2o_results_sample_50000_shap_2000.rds')}
results <- readRDS("data_storage/h2o/real_table_extraction_regex_h2o_results_sample_50000_shap_2000.rds")

shap_num <- results$perc_numeric$shap_values$rf %>% convert_shap_x()
p1a <- shap_num %>% sv_importance(show_numbers = TRUE) +
  labs(title = "% numeric correct") +
  coord_cartesian(xlim = c(0, 0.225))
p1b <- shap_num %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.35, 0.35))

shap_f1 <- results$NA_F1$shap_values$rf %>% convert_shap_x()
p2a <- shap_f1 %>% sv_importance(show_numbers = TRUE) +
  labs(title = "F1 NA") +
  coord_cartesian(xlim = c(0, 0.225))
p2b <- shap_f1 %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.35, 0.35))

shap_binom <- results$binomial$shap_values$rf %>% convert_shap_x()
p3a <- shap_binom %>% sv_importance(show_numbers = TRUE) +
  labs(title = "binomial") +
  coord_cartesian(xlim = c(0, 0.225))
p3b <- shap_binom %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.35, 0.35))

real_table_extraction_regex_shap_plot <- (p1a | p1b) /
  (p2a | p2b) /
  (p3a | p3b)
```

There are multiple hypotheses that don't get supported by the visual
results (see Figure \@ref(fig:real-table-extraction-regex-shap-plot)).
The pretty surprising results are:

1.  The visual separation of columns or rows has an effect on the text
    processing.
2.  It seems to have a positive effect on F1 and numeric correctness
    rate if the Passiva table is on the same page, even though it has no
    influence on the single predictions.

But one has to keep in mind that the number of data points on the
aggregated values for the test set of the real dataset is only
`r shap_num$X %>% nrow()`. So these findings are not strongly supporting
any interpretation at all. Furthermore, the found effects are not very
large - most below 5 %. Only the hypothesis for a positive influence of
a missing of a value for the binomial prediction gets solid support with
a mean absolute \acr{SHAP} value of over
`r floor((shap_binom %>% sv_importance(kind = "no") %>% .['missing'])*100)`
%. To get reliable results more tables have to be included which would
require additional manual encoding.

```{r hypotheses-and-results-real-table-extraction-regex-table, class.chunk="big-table", echo=echo_flag, warning=warning_flag, message=message_flag}
htmltools::includeHTML(
  textConnection(
    xml2::read_html("../benchmark_results/table_extraction/hypotheses_and_results_real_table_extraction_regex.html") %>%
      xml2::xml_find_first("//table") %>%
      as.character()
  )
)
```

###### Synthetic dataset

```{r synth-table-extraction-regex-shap-loading, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", cache=TRUE, cache.extra = tools::md5sum('data_storage/h2o/synth_table_extraction_regex_h2o_results_sample_50000_shap_2000.rds')}
results <- readRDS("data_storage/h2o/synth_table_extraction_regex_h2o_results_sample_50000_shap_2000.rds")

shap_num <- results$perc_numeric$shap_values$rf %>% convert_shap_x()
p1a <- shap_num %>% sv_importance(show_numbers = TRUE) +
  labs(title = "% numeric correct") +
  coord_cartesian(xlim = c(0, 0.15))
p1b <- shap_num %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.5, 0.5))

p1c <- shap_num %>% sv_dependence("extraction_backend") +
  labs(title = "dependence plot for extraction_backend")
p1d <- shap_num %>% sv_dependence("header_span", color_var = "extraction_backend") +
  labs(title = "dependence plot for header_span")


shap_f1 <- results$NA_F1$shap_values$rf %>% convert_shap_x()
p2a <- shap_f1 %>% sv_importance(show_numbers = TRUE) +
  labs(title = "F1 NA") +
  coord_cartesian(xlim = c(0, 0.15))
p2b <- shap_f1 %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.5, 0.5))

shap_binom <- results$binomial$shap_values$rf %>% convert_shap_x()
p3a <- shap_binom %>% sv_importance(show_numbers = TRUE) +
  labs(title = "binomial") +
  coord_cartesian(xlim = c(0, 0.15))
p3b <- shap_binom %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.5, 0.5))

synth_table_extraction_regex_shap_plot <- (p1a | p1b) /
  (p2a | p2b) /
  (p3a | p3b)

synth_table_extraction_regex_shap_num_details_plot <- p1c | p1d
```

Interpreting the visual results for the \acr{SHAP} analysis on the
synthetic dataset brought some interesting insides into the question
under which condition the two PDF extraction libraries perform
differently. These results can be treated as reliable since the model
has been trained with 50_000 rows and the \acr{SHAP} values have been
calculated on 2_000 rows each.

Very interestingly the number of columns is having an opposite effect
for the two libraries (see Figure
\@ref(fig:synth-table-extraction-regex-shap-num-details-plot) A).
Besides that often only pdfium struggled with some of the table
characteristics while pymupdf is not influenced by them (for an example
with header_span see Figure
\@ref(fig:synth-table-extraction-regex-shap-num-details-plot) B).

It might be worth noting that the row for *Anteile an verbundenen
Unternehmen* was rated to have a clear negative effect on the chance to
extract the correct value.

Since there has no synthetic data created where also the Passiva table
is present the result found with the real dataset can't be investigated
further. Also the question if visual separation is having an effect was
not studied, even though, creating such additional synthetic data would
be very easy with the current generation process and could be done in
future work. It would be interesting if the visual separation is cause
for the maleous text extractions of pdfium as well.

```{r hypotheses-and-results-synth-table-extraction-regex-table, class.chunk="big-table", echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
if (knitr::is_latex_output()) {
  # Render as kable for PDF
  # Extract table from HTML and convert to data.frame
  html_table <- xml2::read_html("../benchmark_results/table_extraction/hypotheses_and_results_synth_table_extraction_regex.html") %>%
    xml2::xml_find_first("//table")
  df <- rvest::html_table(html_table, fill = TRUE)
  kable(df, format = "latex", booktabs = TRUE, longtable = TRUE)
} else {
  # Render as HTML for HTML output
  htmltools::includeHTML(
    textConnection(
      xml2::read_html("../benchmark_results/table_extraction/hypotheses_and_results_synth_table_extraction_regex.html") %>%
        xml2::xml_find_first("//table") %>%
        as.character()
    )
  )
}
```

```{r synth-table-extraction-regex-shap-num-details-plot, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.width=8, fig.height=4, fig.cap="Showing the influence of the extraxtion library on the numeric text extraction task with synthetic data"}
synth_table_extraction_regex_shap_num_details_plot + 
  plot_annotation(tag_levels = 'A')
```

### Extraction with LLMs

```{r loading-table-extraction-datasets, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum(c("data_storage/real_table_extraction_llm.rds", "data_storage/real_table_extraction_synth.rds", "data_storage/real_table_extraction_azure.rds"))}
df_real_table_extraction <- readRDS("data_storage/real_table_extraction_llm.rds") %>% 
  filter(!model %in% c("deepseek-ai_DeepSeek-R1-Distill-Qwen-32B", 'google_gemma-3n-E4B-it')) %>% 
  mutate(model = gsub("^[^_]+_", "", model))
df_real_table_extraction_synth <- readRDS("data_storage/real_table_extraction_synth.rds") %>% 
  mutate(model = gsub("^[^_]+_", "", model))
df_real_table_extraction_azure <- readRDS("data_storage/real_table_extraction_azure.rds") %>% 
  mutate(model = gsub("^[^_]+_", "", model))

model_by_size <- c(
  'gemma-3-4b-it', #'gemma-3n-E4B-it', 
  "gemma-3-12b-it", "gemma-3-27b-it", 
  "Llama-3.1-8B-Instruct", "Llama-3.1-70B-Instruct", "Llama-3.3-70B-Instruct",
  "Llama-4-Scout-17B-16E-Instruct", "Llama-4-Maverick-17B-128E-Instruct-FP8",
  "Mistral-8B-Instruct-2410", "Mistral-Small-3.1-24B-Instruct-2503",
  "Mistral-Large-Instruct-2411", "Qwen2.5-0.5B-Instruct",
  "Qwen2.5-1.5B-Instruct", "Qwen2.5-3B-Instruct", "Qwen2.5-7B-Instruct",
  "Qwen2.5-14B-Instruct", "Qwen2.5-32B-Instruct", "Qwen2.5-72B-Instruct",
  "Qwen3-0.6B", "Qwen3-1.7B", "Qwen3-4B",
  "Qwen3-8B", "Qwen3-14B", "Qwen3-30B-A3B-Instruct-2507", "Qwen3-32B",
  "Qwen3-235B-A22B-Instruct-2507-FP8", "Qwen3-235B-A22B-Instruct-2507",
  # "gpt-4.1-nano", "gpt-4.1-mini", "gpt-4.1",
  "Falcon3-10B-Instruct", "phi-4"
)

method_order <- c("top_n_rag_examples", "n_random_examples", "top_n_rag_examples_out_of_sample", "static_example", "zero_shot" )

norm_factors <- read_csv("../benchmark_jobs/page_identification/gpu_benchmark/runtime_factors_real_table_extraction.csv") %>% 
  mutate(
    model_name = model_name %>% str_replace("/", "_")
  )
norm_factors_few_examples <- norm_factors %>% filter((str_ends(filename, "binary.yaml") | str_ends(filename, "multi.yaml") | str_ends(filename, "vllm_batched.yaml")))

df_real_table_extraction <- df_real_table_extraction %>% left_join(
  norm_factors_few_examples %>% mutate(model_name = gsub("^[^_]+_", "", model_name)) %>% select(model_name, parameter_count, normalization_factor),
  by = c("model" = "model_name")
  )

df_overview <- bind_rows(df_real_table_extraction, df_real_table_extraction_azure) %>% 
  filter(out_of_company != TRUE | is.na(out_of_company), n_examples <= 3) %>% 
  filter(model %in% model_by_size) %>%
  mutate(
    model = factor(model, levels = model_by_size),
    method_family = factor(method_family, levels = method_order),
    n_examples = fct_rev(ordered(paste("n =", n_examples)))
    )

units_real_tables <- read_csv("../benchmark_truth/real_tables/table_characteristics.csv") %>% mutate(
  filepath = paste0('/pvc/benchmark_truth/real_tables/', company, '__', filename),
  T_EUR = (T_in_year + T_in_previous_year)>0,
  T_EUR_both = (T_in_year + T_in_previous_year)>1
) %>% select(filepath, T_EUR, T_EUR_both)

df_real_table_extraction_synth <- df_real_table_extraction_synth %>% left_join(units_real_tables)

# real

zero_shot_stars <- df_real_table_extraction %>% filter(method == "zero_shot") %>% group_by(model) %>% 
  reframe(mean_total = mean(percentage_correct_total, na.rm = TRUE), mean_num = mean(percentage_correct_numeric, na.rm = TRUE), mean_F1 = mean(NA_F1, na.rm = TRUE)) %>% filter(mean_total>real_table_extraction_regex_total_performance_mean, mean_num>real_table_extraction_regex_num_performance_mean, mean_F1>real_table_extraction_regex_NA_F1_mean)

static_example_stars <- df_real_table_extraction %>% filter(method == "static_example") %>% group_by(model) %>% 
  reframe(mean_total = mean(percentage_correct_total, na.rm = TRUE), mean_num = mean(percentage_correct_numeric, na.rm = TRUE), mean_F1 = mean(NA_F1, na.rm = TRUE)) %>% filter(mean_total>real_table_extraction_regex_total_performance_mean, mean_num>real_table_extraction_regex_num_performance_mean, mean_F1>real_table_extraction_regex_NA_F1_mean)

underperformer <- df_real_table_extraction %>% filter(!method %in% c('zero_shot', 'static_example')) %>% group_by(model, method) %>% 
  reframe(mean_total = mean(percentage_correct_total, na.rm = TRUE), mean_num = mean(percentage_correct_numeric, na.rm = TRUE), mean_F1 = mean(NA_F1, na.rm = TRUE)) %>% group_by(model) %>% filter(any(mean_total<real_table_extraction_regex_total_performance_mean, mean_num<real_table_extraction_regex_num_performance_mean, mean_F1<real_table_extraction_regex_NA_F1_mean)) %>% arrange(mean_total) %>% slice_head(n = 1)

super_underperformer <- df_real_table_extraction %>% filter(!method %in% c('zero_shot', 'static_example'), n_examples>1) %>% group_by(model, method) %>% 
  reframe(mean_total = mean(percentage_correct_total, na.rm = TRUE), mean_num = mean(percentage_correct_numeric, na.rm = TRUE), mean_F1 = mean(NA_F1, na.rm = TRUE)) %>% group_by(model) %>% filter(any(mean_total<real_table_extraction_regex_total_performance_mean, mean_num<real_table_extraction_regex_num_performance_mean, mean_F1<real_table_extraction_regex_NA_F1_mean)) %>% arrange(mean_total) %>% slice_head(n = 1)
```

-   confidence usable to head for user checks?
-   not handled new entries
-   five examples bring not much more, but a little
-   random forest / SHAP

#### Real tables only

For the table extraction task
`r length(unique(df_real_table_extraction$model))+2` open source models
have been benchmarked[^05_results-29]. The results are presented in
Figure \@ref(fig:table-extraction-llm-performance-total-overview),
\@ref(fig:table-extraction-llm-performance-numeric-overview) and
\@ref(fig:table-extraction-llm-f1-overview)).

[^05_results-29]: The models *deepseek-ai_DeepSeek-R1-Distill-Qwen-32B*
    and *google_gemma-3n-E4B-it* have been tested as well but don't get
    presented as they never performed anywhere beyond random guessing.

Most models need a context learning approach to beat the performance of
the regular expression approach at total and numeric correctness rate
and F1 score. Only `r nrow(zero_shot_stars)` models perform better
without any guidance[^05_results-30] (see Table
\@ref(tab:zero-and-synth-performing-models-tab)).
`r nrow(static_example_stars)` models achieved an performance better as
the regex baseline using the approach to learn with a fixed example from
the synthetic dataset.

[^05_results-30]: There is an external guidance through the provided
    xgrammar template but it is not communicated to the model in a
    promt.

In contrast: most of the models achieved a better performance than the
regex baseline when they were provided with one or more examples from
real *Aktiva* tables. Just `r nrow(super_underperformer)` don't achieve
a better value even with three or five realistic examples (see Table
\@ref(tab:super-underperforming-models-tab)). Here we find the smallest
models with less than 2B parameters which don't achieve a consistence
performance no matter how many examples they get. But we also find
models that start to perform bad if they get a too long context with too
many examples like the very recent and large model Llama 4 Maverick.

With one and three examples the performance within one model family is
positive correlated with the number of parameters the models have. Once
the 4B parameters are passed the improvements get less and less getting
closer to a perfect performance but never reaching it on all documents.
Table \@ref(tab:table-metrics-llm-real-table-extraction) shows the mean
performance for the best model-method approach for each model family.
Most of the top performing model-method combinations rely on the maximum
number of examples provided. Only the Llama-3 and Falcon3 model perform
best with three examples[^05_results-31].

[^05_results-31]: Phi4 also perfroms best with three examples. But this
    is the maximum it can process due to a limited context length.

```{r zero-and-synth-performing-models-tab, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
zero_shot_stars %>% full_join(static_example_stars, join_by(model), suffix = c("_zero_shot", "_static_example")) %>% select(model, starts_with("mean_total")) %>% 
mutate_if(
    is.numeric, 
    ~ifelse(
      . == max(., na.rm = TRUE),
      paste0("**", round(., 3), "**"),
      round(., 3)
    )
  ) %>% 
  render_table(
    alignment = "lrr",
    caption = "Comparing table extraction performance with real 'Aktiva' dataset for models that perform well without or with little context learning",
    ref = opts_current$get("label"), dom="t")
```

```{r super-underperforming-models-tab, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
super_underperformer %>% select(model, method, mean_total) %>% ungroup() %>% # mutate(model = str_replace_all(model, "_", " ")) %>% 
mutate_if(
    is.numeric, 
    ~ifelse(
      . == max(., na.rm = TRUE),
      paste0("**", round(., 3), "**"),
      round(., 3)
    )
  ) %>% 
  render_table(
    alignment = "llr",
    caption = "Comparing table extraction performance with real 'Aktiva' dataset for models that worse than the regex baselin with 3 or 5 examples for in-context learning",
    ref = opts_current$get("label"), dom="t")
```

```{r table-metrics-llm-real-table-extraction, echo=echo_flag, warning=warning_flag, message=message_flag, results="asis"}
df_real_table_extraction %>% group_by(model, method) %>% mutate(mean_total = mean(percentage_correct_total, na.rm = TRUE)) %>%
  arrange(desc(mean_total)) %>% group_by(model_family) %>% slice_head(n = 1) %>% 
  select(model_family, model, method_family, n_examples, mean_total) %>% arrange(desc(mean_total)) %>% 
  mutate(mean_total = round(mean_total, 3)) %>% 
  render_table(caption = "Comparing best mean table extraction performance with real 'Aktiva' dataset for each model family", ref = opts_current$get("label"), alignment = "lllrr", dom="t")
```

```{r table-metrics-llm-real-table-extraction-small, echo=echo_flag, warning=warning_flag, message=message_flag, results="asis"}
df_real_table_extraction %>% group_by(model, method) %>% filter(parameter_count<17) %>% mutate(mean_total = mean(percentage_correct_total, na.rm = TRUE)) %>%
  arrange(desc(mean_total)) %>% group_by(model_family) %>% slice_head(n = 1) %>% 
  select(model_family, model, method_family, n_examples, mean_total) %>% arrange(desc(mean_total)) %>% 
  mutate(mean_total = round(mean_total, 3)) %>% 
  render_table(caption = "Comparing best mean table extraction performance with real 'Aktiva' dataset for each model family for models with less than 17B parameters", ref = opts_current$get("label"), alignment = "lllrr", dom="t")
```

Based on a small sample of
`r df_overview %>% select(filepath) %>% filter(str_detect(filepath, "Statistik")) %>% unique() %>% nrow()`
documents by the *Amt für Statistik Berlin-Brandenburg* it seems that
there is support for the hypothesis, that providing Aktiva tables from
the same company in in-context learning, is improving the results. This
is especially noticeable for models with very few parameters and when
providing only a single example. This seems intuitive, since there the
potential for possibilities is much bigger. Figure
\@ref(fig:table-extraction-llm-performance-total-overview-out-of-company)
shows that on this limited sample

-   the improvement is bigger for Qwen 3 than for Qwen 2.5
-   Googles gemma 27b and GPT 4.1 mini could overcome an unnoticed issue
    with the extraction with just one example.
-   the effect of being overwhelmed by a too rich context with LLamas
    Maverick model could get reduced a bit.

To examine the question, if the reported confidence score of the
responses can be used, to flag the predicted values as potentially
wrong. Again, Figure \@ref(fig:table-extraction-llm-confidence) shows,
that Qwen 3 reports very high confidence values no matter if the results
are correct or not. With the Mistral model we find a wider range of
confidences given and for wrong results lower confidence is reported.

Figure \@ref(fig:table-extraction-llm-confidence-lm) shows, that the
chance to make an mistake by believing the prediction is rising with
lower confidence. The chance to make a mistake is higher for predictions
of numeric values than for believing a value is not present in the
table. The chance to make such a mistake is higher using the confidence
reported by Qwen 3.

```{r table-extraction-llm-confidence, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparing the reported confidence scores for the table extraction task on real dataset for the Mistral and Qwen 3 with 8B parameters.", cache=TRUE, cache.extra = tools::md5sum("data_storage/real_table_extraction_llm.rds")}
confidence_vs_truth <- df_real_table_extraction %>% 
  filter(model %in% c("Ministral-8B-Instruct-2410", "Qwen3-8B")) %>% 
  group_by(method, model) %>% mutate(
    mean_percentage_correct_total = mean(percentage_correct_total, na.rm=TRUE), .before = 1
    ) %>% group_by(model) %>% 
  arrange(desc(mean_percentage_correct_total)) %>% 
  slice_max(mean_percentage_correct_total, n = 1, with_ties = TRUE) %>% 
  mutate(predictions_processed = map(predictions, ~{
    .x %>% 
      select(-"_merge") %>% 
      mutate(
        match = (year_truth == year_result) | (is.na(year_truth) & is.na(year_result)),
        confidence = confidence_this_year,
        truth_NA = is.na(year_truth),
        predicted_NA = is.na(year_result),
        .before = 4
      ) %>% nest(
        tuple_year = c(match, confidence, truth_NA, predicted_NA)
      ) %>% 
      mutate(
        confidence = confidence_previous_year,
        match = (previous_year_truth == previous_year_result) | (is.na(previous_year_truth) & is.na(previous_year_result)),
        truth_NA = is.na(previous_year_truth),
        predicted_NA = is.na(previous_year_result),
        .before = 4
      ) %>% nest(
        tuple_previous_year = c(match, confidence, truth_NA, predicted_NA)
      ) %>% select(
        -c(year_truth, previous_year_truth, year_result, previous_year_result,
           confidence_this_year, confidence_previous_year)
      ) %>% 
      pivot_longer(-c("E1", "E2", "E3")) %>% 
      unnest(cols = value) %>% mutate(
        match = if_else(is.na(match), FALSE, match)
      )
  })) %>% 
  unnest(predictions_processed) %>% mutate(
    match = factor(match, levels = c(F, T)),
    truth_NA = factor(truth_NA, levels = c(F, T))
  )

confidence_vs_truth %>% ggplot() +
  geom_boxplot(
    aes(x = match, y = confidence, fill = truth_NA), 
    position = position_dodge2(preserve = "single")) +
  scale_fill_discrete(drop = FALSE) +
  scale_x_discrete(drop = FALSE) +
  facet_grid(~ model)
```

```{r table-extraction-llm-confidence-lm, echo=echo_flag, warning=warning_flag, message=message_flag, fig.height=3, out.width="100%", fig.cap="Estimating the relative frequency to find a wrong extraction result over different confidence intervals"}
confidence_vs_truth %>%
  mutate(
    conf_interval = cut(confidence, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE),
    conf_center = as.numeric(sub("\\((.+),(.+)\\]", "\\1", levels(conf_interval))[conf_interval]) + 0.005
  ) %>%
  group_by(conf_center, predicted_NA, model) %>%
  summarize(
    n_true = sum(match == TRUE, na.rm = TRUE),
    n_false = sum(match == FALSE, na.rm = TRUE),
    total = n_true + n_false,
    chance_false = if_else(total > 0, n_false / total * 100, NA_real_),
    .groups = "drop"
  ) %>%
  ggplot(aes(x = conf_center, y = chance_false, color = predicted_NA)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(x = "Confidence Interval Center", y = "Chance False (%)", color = "Predicted NA") +
  coord_cartesian(ylim = c(0, 100), xlim = c(0,1)) +
  facet_grid(~ model)
```

##### Hypotheses

The formulated hypotheses have been evaluated visually using the
dependence and beeswarm plots from the shapviz library based on the
\acr{SHAP} values calculated with a random forest.

```{r real-table-extraction-llm-shap-loading, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", cache=TRUE, cache.extra = tools::md5sum('data_storage/h2o/real_table_extraction_h2o_results_sample_50000_shap_2000.rds')}
results <- readRDS("data_storage/h2o/real_table_extraction_h2o_results_sample_50000_shap_2000.rds")

shap_num <- results$perc_numeric$shap_values$rf %>% convert_shap_x()
p1a <- shap_num %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "% numeric correct") +
  coord_cartesian(xlim = c(0, 0.12))
p1b <- shap_num %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))
p1c <- shap_num %>% sv_dependence("n_examples")
p1d <- shap_num %>% sv_dependence("T_in_year")

shap_f1 <- results$NA_F1$shap_values$rf %>% convert_shap_x()
p2a <- shap_f1 %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "F1 NA") +
  coord_cartesian(xlim = c(0, 0.12))
p2b <- shap_f1 %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

shap_binom <- results$binomial$shap_values$rf %>% convert_shap_x()
p3a <- shap_binom %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "binomial") +
  coord_cartesian(xlim = c(0, 0.12))
p3b <- shap_binom %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

shap_binom <- results$confidence$shap_values$rf %>% convert_shap_x()
p4a <- shap_binom %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "confidence") +
  coord_cartesian(xlim = c(0, 0.12))
p4b <- shap_binom %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

real_table_extraction_llm_shap_plot <- (p1a | p1b) /
  (p2a | p2b) /
  (p3a | p3b) /
  (p4a | p4b)

real_table_extraction_llm_shap_num_details_plot <- p1c | p1d
```

Even though the samples size of Aktiva tables did not increase, the
available training, test and \acr{SHAP} sample size is much larger,
because the experiment has been repeated with different models and
methods. Thus, the interpretations based on the visual evaluation (see
Figure \@ref(fig:real-table-extraction-llm-shap-plot))) are more
reliable for model and method specific predictors. Since there is one
Aktiva example for every company files were found for they might even be
generalizable for this population. But one has to keep in mind that
there have been more Aktiva tables for *Amt Stat BBB* which might nudge
the results a bit.

The results assign much more influence on model and method specific
attributes than on the table specific attributes. The importance of the
table attributes are as low as found with the regular expresion
approach. Only for the binomial prediction we find the predictor
*missing* to get assigned more importance than to all model and method
specific attributes. Same is true for the *label* that is having the
highest influence on the reported confidence. Nevertheless, in the case
of the binomial prediction there is half of the predictors *missing* and
*label* importance shifted to model and method specific predictors.

Again, multiple hypotheses don't get supported by the visual results.
The surprising results are:

1.  In general more examples are helpful except for Llamas Maverick
    model that performs poorly with five examples. But this effect is
    only noticeable with the aggregated metrics nor for the case wise
    binomial evaluation.
2.  The number of columns has a negative effect on the performance but
    no effect on the reported confidence.
3.  There was no negative effect found if the *Passiva* table is on the
    same page as the Aktiva table.
4.  Larger models start to report less confidence again. This is not
    unexpected for the Mistral model but was surprising for the largest
    Qwen 3 model. (Discussion: New Generation? Aktive paramaters count?
    Irrelevant because not well distinguishing?)
5.  It not only influences the the performance to extract the correct
    numeric value from a row where there are additional sums present but
    also the F1 score.

Two interesting details found while inspecting the dependence plots for
the metric *percentage_numeric_correct* are (see Figure
\@ref(fig:real-table-extraction-llm-shap-num-details-plot) A) that the
bad performance of LLamas Maverick with five examples is easily
spottable and that the negative effect of *T_in_year* might be caused by
an interaction with *vis_separated_rows* completely (see Figure
\@ref(fig:real-table-extraction-llm-shap-num-details-plot) B). To
investigate the second finding one would need tables where the uni is
present in the year column and having no visual separation of the rows
at the same time. Synthetic data potentiall could help to answer such
questions.

```{r hypotheses-and-results-real-table-extraction-llm-table, class.chunk="big-table", echo=echo_flag, warning=warning_flag, message=message_flag}
htmltools::includeHTML(
  textConnection(
    xml2::read_html("../benchmark_results/table_extraction/hypotheses_and_results_real_table_extraction_llm.html") %>%
      xml2::xml_find_first("//table") %>%
      as.character()
  )
)
```

```{r real-table-extraction-llm-shap-num-details-plot, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.width=8, fig.height=4, fig.cap="Showing the influence of many examples on Llama 4 Maverick (A) and interaction between *T in year* and *vis separated rows* (B)"}
real_table_extraction_llm_shap_num_details_plot + 
  plot_annotation(tag_levels = 'A')
```

##### GPT

```{r, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum(c("data_storage/real_table_extraction_llm.rds", "data_storage/real_table_extraction_azure.rds"))}
model_by_size_gpt <- c(
  "Qwen3-0.6B", "Qwen3-8B", "Qwen3-30B-A3B-Instruct-2507", "Qwen3-235B-A22B-Instruct-2507",
 "gpt-4.1-nano", "gpt-4.1-mini", "gpt-5-mini",
 "gpt-oss-20b", "gpt-oss-120b", "gpt-4.1"
)

df_overview_gpt <- bind_rows(df_real_table_extraction, df_real_table_extraction_azure) %>% 
  filter(out_of_company == TRUE | is.na(out_of_company), n_examples <= 3) %>% 
  filter(model %in% model_by_size_gpt) %>%
  mutate(
    model = factor(model, levels = model_by_size_gpt),
    method_family = factor(method_family, levels = method_order),
    n_examples = fct_rev(ordered(paste("n =", n_examples)))
    ) %>% mutate(
      n_predicted_values = NA_true_positive+NA_false_positive+NA_false_negative+NA_true_negative
    )

perc_wrong_prediction_count <- round((df_overview_gpt %>% filter(!str_detect(model, "Qwen"), n_predicted_values != 58) %>% nrow()/df_overview_gpt %>% filter(!str_detect(model, "Qwen")) %>% nrow()) * 100, 1)
```

Even though a lot of documents to process at \acr{RHvB} will not be
public and thus must not be processed on public cloud infrastructure,
the performance of models like OpenAI's GPT or Google's Gemini are
interesting benchmark references within this thesis and for comparing
these findings with other papers results. Therefore for this thesis the
public available versions of annual reports have been used instead of
the ones used internally or for public administration purposes. Those
public available reports often are visually more appealing and more
heterogeneous in their structure.

As a reference to compare the performance of OpenAI's models with the
results of four Qwen 3 models are shown as well (see Figure
\@ref(fig:table-extraction-llm-performance-total-gpt),
\@ref(fig:table-extraction-llm-performance-numeric-gpt) and
\@ref(fig:table-extraction-llm-f1-gpt)). Surprisingly gpt-5-mini is
almost performing as good as the top Qwen 3 model and gpt-4.1. But
besides gpt-4.1-nano and Qwen3-0.6B all models perform pretty well with
the random or top-rag example methods. The ranking for the best
model-method combination can be found in Table
\@ref(tab.table-extraction-llm-performance-total-gpt-ranking). Since
gpt-4.1 costs five times more than gpt-4.1-mini (see Table
\@ref{tab:costs-azure}) it seems reasonable to prefer the smaller model
for this specific task.

Costs for gpt-5-mini not shown in Azure yet. :(

The author was not able to get OpenAI's models to stick to the provided
\acr{json} schema strictly. Passing the \acr{ebnf} grammar did not work
at all. This means that with gpt-4.1-nano there have been
`r df_overview_gpt %>% filter(n_predicted_values == 0) %>% nrow()`
predictions that have been completely empty. Overall there have been
`r perc_wrong_prediction_count` % of the responses of OpenAI's models
that were compatible with the schema but had a wrong number of rows
predicted (see Figure
\@red(fig:table-extraction-llm-prediction-count-gpt)).

Using gpt-5-nano and gpt-5-chat for the table extraction task was not
working. With gpt-5-nano the answers were not respecting the provided
grammar. Running gpt-5-chat resulted in the error informing that a
*json_schema* can't be used with this model. With gpt-5-mini the very
approach worked flawless. Running gpt-oss-20b with the vllm offline
inference framework was possible and the new harmony output format could
be processed after minor code changes for most
approaches[^05_results-32]. With a gpt-oss-120b instance hosted on Azure
the guided decoding worked flawless.

[^05_results-32]: With the static example approach there have been 24
    files where the resonse could not get parsed into valid \acr{json}.
    With the other approaches there are one to four unparsable
    responses.

```{r table-extraction-llm-performance-total-gpt-ranking, echo=echo_flag, warning=warning_flag, message=message_flag}
df_overview_gpt %>% group_by(model, method) %>% summarise(mean_percentage_correct_total = mean(percentage_correct_total)) %>% arrange(desc(mean_percentage_correct_total)) %>% slice_head(n = 1) %>% ungroup() %>% arrange(desc(mean_percentage_correct_total)) %>% mutate(mean_percentage_correct_total = round(mean_percentage_correct_total, 2)) %>% rename("mean correct total" = mean_percentage_correct_total) %>% kbl()
```

```{r table-extraction-llm-performance-total-gpt, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", out.height="95%", out.extra='keepaspectratio', fig.height=10, fig.cap="Comparing the percentage of correct predictions overall for OpenAi's LLMs with some Qwen 3 models"}
df_overview_gpt %>% 
  ggplot() +
  geom_hline(yintercept = real_table_extraction_regex_total_performance_mean, linetype = "dashed") +
  geom_boxplot(aes(x = model, y = percentage_correct_total, fill = model_family)) +
  # geom_jitter(aes(x = model, y = percentage_correct_total)) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_nested(method_family + n_examples ~ .) +
  theme(
    axis.title.x = element_blank()
  )
```

```{r table-extraction-llm-performance-numeric-gpt, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", out.height="95%", out.extra='keepaspectratio', fig.height=10, fig.cap="Comparing the percentage of correct numeric predictions for OpenAi's LLMs with some Qwen 3 models"}
df_overview_gpt %>% 
  ggplot() +
  geom_hline(yintercept = real_table_extraction_regex_num_performance_mean, linetype = "dashed") +
  geom_boxplot(aes(x = model, y = percentage_correct_numeric, fill = model_family)) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_nested(method_family + n_examples ~ .) +
  theme(
    axis.title.x = element_blank()
  )
```

```{r table-extraction-llm-f1-gpt, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", out.height="90%", out.extra='keepaspectratio', fig.height=10, , fig.cap="Comparing the F1 score for predicting the missingness of a value for OpenAi's LLMs with some Qwen 3 models. The green crosses indicate results where a model has predicted only numeric values even though there have been missing values."}
df_overview_gpt %>% 
  mutate(NA_F1 = if_else(is.na(NA_F1), 0, NA_F1)) %>% 
  ggplot() +
  geom_hline(yintercept = real_table_extraction_regex_NA_F1_mean, linetype = "dashed") +
  geom_boxplot(aes(x = model, y = NA_F1, fill = model_family)) +
  geom_jitter(data = df_overview_gpt %>% filter(is.na(NA_F1)), aes(x = model, y = 0), height = 0, color = "green", alpha = .5, shape = 4) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_nested(method_family + n_examples ~ .) +
  theme(
    axis.title.x = element_blank()
  )
```

```{r table-extraction-llm-prediction-count-gpt, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Showing the number of predictions OpenAI's models made."}
df_overview_gpt %>% filter(n_predicted_values != 58) %>% 
  mutate(color_code = if_else(n_predicted_values==0, "zero", if_else(n_predicted_values > 58, "too many", "too little"))) %>% 
  ggplot() +
  geom_histogram(aes(x = n_predicted_values, fill = color_code)) + facet_grid(model~.)
```

```{r costs-azure, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum(c("data_storage/real_table_extraction_llm.rds", "data_storage/real_table_extraction_azure.rds"))}
costs_azure <- read_csv("../CostManagement_master-thesis_2025.csv")

token_prop <- df_real_table_extraction %>% group_by(model, method, n_examples) %>% summarize(
    request_tokens_total = sum(request_tokens[[1]])) %>% 
    group_by(method, n_examples) %>% 
    summarize(mean = mean(request_tokens_total, na.rm = TRUE)) %>% mutate(five_examples = n_examples == 5) %>% group_by(five_examples) %>% summarise(sum = sum(mean))

five_ex_tokens <- token_prop %>% filter(five_examples == TRUE) %>% pull(sum)
other_tokens <- token_prop %>% filter(five_examples == FALSE) %>% pull(sum)

costs_azure %>% mutate(
  Cost_all_tasks = Cost,
  Cost_all_tasks = if_else(Meter == "gpt 4.1 Inp glbl Tokens", Cost_all_tasks+Cost_all_tasks*five_ex_tokens/other_tokens, Cost_all_tasks),
  Cost_all_tasks = if_else(Meter == "gpt 4.1 Outp glbl Tokens", Cost_all_tasks+Cost_all_tasks*3/11, Cost_all_tasks)
  ) %>% mutate(
    Cost = round(Cost, 2),
    Cost_all_tasks = round(Cost_all_tasks, 2)
  ) %>% rename(
    Model = Meter
  ) %>% select(
    Model, Cost, Cost_all_tasks, Currency
  ) %>% kbl()
```

#### Synthetic tables only

```{r loading-synth-table-extraction-datasets, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum(c("data_storage/synth_table_extraction_llm.rds")), cache.lazy = FALSE}
df_synth_table_extraction <- readRDS("data_storage/synth_table_extraction_llm.rds") %>% 
  filter(!model %in% c("deepseek-ai_DeepSeek-R1-Distill-Qwen-32B", 'google_gemma-3n-E4B-it')) %>% 
  mutate(model = gsub("^[^_]+_", "", model)) %>% sample_frac(size = .1)

norm_factors <- read_csv("../benchmark_jobs/page_identification/gpu_benchmark/runtime_factors_synth_table_extraction.csv") %>% 
  mutate(
    model_name = model_name %>% str_replace("/", "_")
  )
norm_factors_few_examples <- norm_factors %>% filter((str_ends(filename, "binary.yaml") | str_ends(filename, "multi.yaml") | str_ends(filename, "vllm_batched.yaml")))

df_synth_table_extraction <- df_synth_table_extraction %>% left_join(
  norm_factors_few_examples %>% mutate(model_name = gsub("^[^_]+_", "", model_name)) %>% select(model_name, parameter_count, normalization_factor),
  by = c("model" = "model_name")
  )

# model_by_size <- c(
#   'gemma-3-4b-it', #'gemma-3n-E4B-it', 
#   "gemma-3-12b-it", "gemma-3-27b-it", 
#   "Llama-3.1-8B-Instruct", "Llama-3.1-70B-Instruct", "Llama-3.3-70B-Instruct",
#   "Llama-4-Scout-17B-16E-Instruct", "Llama-4-Maverick-17B-128E-Instruct-FP8",
#   "Mistral-8B-Instruct-2410", "Mistral-Small-3.1-24B-Instruct-2503",
#   "Mistral-Large-Instruct-2411", "Qwen2.5-0.5B-Instruct",
#   "Qwen2.5-1.5B-Instruct", "Qwen2.5-3B-Instruct", "Qwen2.5-7B-Instruct",
#   "Qwen2.5-14B-Instruct", "Qwen2.5-32B-Instruct", "Qwen2.5-72B-Instruct",
#   "Qwen3-0.6B", "Qwen3-1.7B", "Qwen3-4B",
#   "Qwen3-8B", "Qwen3-14B", "Qwen3-30B-A3B-Instruct-2507", "Qwen3-32B", "Qwen3-235B-A22B-Instruct-2507",
#   # "gpt-4.1-nano", "gpt-4.1-mini", "gpt-4.1",
#   "Falcon3-10B-Instruct", "phi-4"
# )
# 
# method_order <- c("top_n_rag_examples", "n_random_examples", "top_n_rag_examples_out_of_sample", "static_example", "zero_shot" )

df_overview_synth <- df_synth_table_extraction %>% 
  mutate(
    model = factor(model, levels = model_by_size),
    method_family = factor(method_family, levels = method_order),
    n_examples = fct_rev(ordered(paste("n =", n_examples)))
  )

# synth

zero_shot_stars_synth <- df_synth_table_extraction %>% filter(method == "zero_shot") %>% group_by(model) %>% 
  reframe(mean_total = mean(percentage_correct_total, na.rm = TRUE), mean_num = mean(percentage_correct_numeric, na.rm = TRUE), mean_F1 = mean(NA_F1, na.rm = TRUE)) %>% filter(mean_total>synth_table_extraction_regex_total_performance_mean, mean_num>synth_table_extraction_regex_num_performance_mean, mean_F1>synth_table_extraction_regex_NA_F1_mean)

static_example_stars_synth <- df_synth_table_extraction %>% filter(method == "static_example") %>% group_by(model) %>% 
  reframe(mean_total = mean(percentage_correct_total, na.rm = TRUE), mean_num = mean(percentage_correct_numeric, na.rm = TRUE), mean_F1 = mean(NA_F1, na.rm = TRUE)) %>% filter(mean_total>synth_table_extraction_regex_total_performance_mean, mean_num>synth_table_extraction_regex_num_performance_mean, mean_F1>synth_table_extraction_regex_NA_F1_mean)

underperformer_synth <- df_synth_table_extraction %>% filter(!method %in% c('zero_shot', 'static_example')) %>% group_by(model, method) %>% 
  reframe(mean_total = mean(percentage_correct_total, na.rm = TRUE), mean_num = mean(percentage_correct_numeric, na.rm = TRUE), mean_F1 = mean(NA_F1, na.rm = TRUE)) %>% filter(any(mean_total<synth_table_extraction_regex_total_performance_mean, mean_num<synth_table_extraction_regex_num_performance_mean, mean_F1<synth_table_extraction_regex_NA_F1_mean)) %>% arrange(mean_total) %>% slice_head(n = 1)

super_underperformer_synth <- df_synth_table_extraction %>% filter(!method %in% c('zero_shot', 'static_example'), n_examples>1) %>% group_by(model, method) %>% 
  reframe(mean_total = mean(percentage_correct_total, na.rm = TRUE), mean_num = mean(percentage_correct_numeric, na.rm = TRUE), mean_F1 = mean(NA_F1, na.rm = TRUE)) %>% group_by(model) %>% filter(any(mean_total<synth_table_extraction_regex_total_performance_mean, mean_num<synth_table_extraction_regex_num_performance_mean, mean_F1<synth_table_extraction_regex_NA_F1_mean)) %>% arrange(mean_total) %>% slice_head(n = 1)

df_synth_top_performance <- df_synth_table_extraction %>% group_by(model, method) %>% mutate(mean_total = mean(percentage_correct_total, na.rm = TRUE)) %>%
  arrange(desc(mean_total)) %>% group_by(model_family) %>% slice_head(n = 1) %>% 
  select(model_family, model, method_family, n_examples, mean_total) %>% arrange(desc(mean_total)) %>% 
  mutate(mean_total = round(mean_total, 3))

df_synth_top_performance_small <- df_synth_table_extraction %>% group_by(model, method) %>% filter(parameter_count < 17) %>% mutate(mean_total = mean(percentage_correct_total, na.rm = TRUE)) %>%
  arrange(desc(mean_total)) %>% group_by(model_family) %>% slice_head(n = 1) %>% 
  select(model_family, model, method_family, n_examples, mean_total) %>% arrange(desc(mean_total)) %>% 
  mutate(mean_total = round(mean_total, 3))

n_better_as_regex <- df_synth_table_extraction %>% group_by(model, method) %>% summarise(mean_total = mean(percentage_correct_total, na.rm = TRUE)) %>%
  arrange(desc(mean_total))  %>% 
  mutate(
    better_than_regex =  mean_total>synth_table_extraction_regex_total_performance_mean
  ) %>% group_by(better_than_regex) %>% summarise(n())

n_better_as_regex_families <- df_synth_top_performance %>% mutate(
    better_than_regex =  mean_total>synth_table_extraction_regex_total_performance_mean
) %>% group_by(better_than_regex) %>% summarise(n())
```

Table \@ref(tab:table-metrics-llm-real-table-extraction-synth-context)
shows that for `r deframe(n_better_as_regex_families)['TRUE']` from
`r nrow(n_better_as_regex_families)` model families there is at least
one model-method combination that performed better than the regex
baseline. For the synthetic table extraction task the baseline is
`r synth_table_extraction_regex_total_performance_mean`.

Only `r deframe(n_better_as_regex)['TRUE']` from
`r nrow(n_better_as_regex)` model-method combinations performed better
than this baseline. There has been no model that performed better than
this baseline with the zero or static example approach.

span argument was not implemented correct in html tables and md :/

already just using 10 % of documents generated; and then 10 % of that
sum of all experiment results (factor 14) with random forest?

\@ref(fig:synth-table-extraction-llm-performance-total-overview)

```{r table-metrics-llm-real-table-extraction-synth-context, echo=echo_flag, warning=warning_flag, message=message_flag, results="asis"}
df_synth_top_performance %>% 
  mutate(
    mean_total = if_else(
      mean_total>synth_table_extraction_regex_total_performance_mean,
      paste0("**", mean_total, "**"),
      as.character(mean_total)
      )
  ) %>% 
  render_table(caption = "Comparing best mean table extraction performance with synthetic 'Aktiva' dataset for each model family", ref = opts_current$get("label"), alignment = "lllrr", dom="t")
```

```{r table-metrics-llm-real-table-extraction-synth-context-small, echo=echo_flag, warning=warning_flag, message=message_flag, results="asis"}
df_synth_top_performance_small %>% 
  mutate(
    mean_total = if_else(
      mean_total>synth_table_extraction_regex_total_performance_mean,
      paste0("**", mean_total, "**"),
      as.character(mean_total)
      )
  ) %>% 
  render_table(caption = "Comparing best mean table extraction performance with synthetic 'Aktiva' dataset for each model family for models with less than 17B parameters", ref = opts_current$get("label"), alignment = "lllrr", dom="t")
```

###### Hypotheses

HTML and Markdown better but expected interaction effects mostly not
found - except: - columns help pdf - thinning least bad for pdf - pdf
worst with numbers that have currency units (short numbers, maybe no
1000er delimiter) - enumeration positive for pdf (and interaction with
log10 mult)

line breaks are no problem

zero shot gets confused by text around

Markdown might be even better than HTML

respecting units was bad - except for: Top n rag finds examples with
same currency units (shorter numbers more important than currnency in
header?)

log10 multiplier has many interaction effects

LLama 4 Maverick again problem with five examples

Positive column count effect (different for real data)

#### Extract from real tables with synthetic content

```{r preparing-synth-context-learning-data, echo=echo_flag, warning=warning_flag, message=message_flag, cache=TRUE, cache.extra = tools::md5sum(c("data_storage/real_table_extraction_llm.rds", "data_storage/real_table_extraction_synth.rds"))}
best_models_with_examples <- df_real_table_extraction_synth %>% filter(str_detect(method_family, "n_")) %>% group_by(model, method) %>% summarise(mean_synth = mean(percentage_correct_total)) %>% arrange(desc(mean_synth)) %>% slice_head(n = 1) %>% ungroup() %>% arrange(desc(mean_synth)) %>% left_join(
  df_real_table_extraction %>% group_by(model, method) %>% summarise(mean_real = mean(percentage_correct_total))) %>% 
  left_join(
    df_real_table_extraction %>% group_by(model, method) %>% summarise(mean_zero_shot = mean(percentage_correct_total)) %>% filter(method == "zero_shot") %>% select(-method)
  ) %>% 
  mutate_if(is.numeric, ~round(., 3))

best_mistral <- df_real_table_extraction_synth %>% filter(model_family == "mistralai") %>% group_by(model, method) %>% summarise(mean_synth = mean(percentage_correct_total)) %>% arrange(desc(mean_synth)) %>% slice_head(n = 1) %>% pull(mean_synth)

df_real_table_extraction_synth_respect_units <- df_real_table_extraction_synth %>% 
  mutate(n_col_T_EUR = T_EUR_both + T_EUR) %>% 
  mutate(
    method_family = factor(method_family, levels = method_order),
    n_examples = fct_rev(ordered(paste("n =", n_examples))),
    respect_units = !ignore_units
  )

confidence_vs_truth <- df_real_table_extraction_synth %>% 
  filter(method_family %in% c("top_n_rag_examples", "n_random_examples")) %>% 
  filter(model %in% c("Ministral-8B-Instruct-2410", "Qwen3-8B")) %>% 
  group_by(method, model) %>% mutate(
    mean_percentage_correct_total = mean(percentage_correct_total, na.rm=TRUE), .before = 1,
    respect_units = !ignore_units
  ) %>% group_by(respect_units, model) %>% 
  arrange(desc(mean_percentage_correct_total)) %>% 
  slice_max(mean_percentage_correct_total, n = 1, with_ties = TRUE) %>% 
  mutate(predictions_processed = map(predictions, ~{
    .x %>% 
      select(-"_merge") %>% 
      mutate(
        match = (year_truth == year_result) | (is.na(year_truth) & is.na(year_result)),
        confidence = confidence_this_year,
        truth_NA = is.na(year_truth),
        predicted_NA = is.na(year_result),
        .before = 4
      ) %>% nest(
        tuple_year = c(match, confidence, truth_NA, predicted_NA)
      ) %>% 
      mutate(
        confidence = confidence_previous_year,
        match = (previous_year_truth == previous_year_result) | (is.na(previous_year_truth) & is.na(previous_year_result)),
        truth_NA = is.na(previous_year_truth),
        predicted_NA = is.na(previous_year_result),
        .before = 4
      ) %>% nest(
        tuple_previous_year = c(match, confidence, truth_NA, predicted_NA)
      ) %>% select(
        -c(year_truth, previous_year_truth, year_result, previous_year_result,
           confidence_this_year, confidence_previous_year)
      ) %>% 
      pivot_longer(-c("E1", "E2", "E3")) %>% 
      unnest(cols = value) %>% mutate(
        match = if_else(is.na(match), FALSE, match)
      )
  })) %>% 
  unnest(predictions_processed) %>% mutate(
    match = factor(match, levels = c(F, T)),
    truth_NA = factor(truth_NA, levels = c(F, T))
  )

confidence_vs_truth_linear <- confidence_vs_truth %>%
  mutate(
    conf_interval = cut(confidence, breaks = seq(0, 1, by = 0.05), include.lowest = TRUE),
    conf_center = as.numeric(sub("\\((.+),(.+)\\]", "\\1", levels(conf_interval))[conf_interval]) + 0.005
  ) %>%
  group_by(conf_center, predicted_NA, respect_units, model) %>%
  summarize(
    n_true = sum(match == TRUE, na.rm = TRUE),
    n_false = sum(match == FALSE, na.rm = TRUE),
    total = n_true + n_false,
    chance_false = if_else(total > 0, n_false / total * 100, NA_real_),
    .groups = "drop"
  )
```

Table
\@ref(tab:comparing-extraction-performance-real-and-synth-context-learning)
shows that using real examples for in-context- learning is better than
using the created synthetic data. Nevertheless, it is improving the
overall performance for the table extraction task by almost 20 % for all
models but Google's gemma-3-12b-it. However, the performance difference
with and without in-context learning is the smallest for the
gemma-3-12b-it model as well. The spread of the performance is bigger
using synthetic in-context learning data for all models but Llama 3 8B
Instruct.

Any pattern?

```{r comparing-extraction-performance-real-and-synth-context-learning, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
best_models_with_examples %>% mutate_if(
    is.numeric, 
    ~ifelse(
      . == max(., na.rm = TRUE),
      paste0("**", ., "**"),
      .
    )
  ) %>% 
  render_table(
    alignment = "llrrr",
    caption = "Comparing extraction performance for real Aktiva extraction task with synthetic and real examples for in-context learning with a zero shot approach",
    ref = opts_current$get("label"), dom="t")
```

```{r compare-real-table-extraction-by-context-type, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.height=10, fig.cap="Comparing table extraction perfromance for real Aktiva extraction task with synthetic and real examples for in-contect learning"}
bind_rows(
  df_real_table_extraction_synth %>% mutate(context = "synth"), 
  df_real_table_extraction %>% mutate(context = "real")
  ) %>% 
  filter(model %in% best_models_with_examples$model) %>% 
  filter(method_family %in% c("top_n_rag_examples", "n_random_examples")) %>% 
  mutate(
    # model = factor(model, levels = model_by_size),
    method_family = factor(method_family, levels = method_order),
    n_examples = fct_rev(ordered(paste("n =", n_examples)))
  ) %>% 
  ggplot() +
  geom_boxplot(aes(fill=context, y = percentage_correct_total), alpha = .5) +
  # scale_fill_manual(values = c("blue", "orange")) +
  scale_x_discrete(guide = guide_axis(angle = 30)) +
  facet_nested(method_family+n_examples~model)
```

Table
\@ref(tab:comparing-extraction-performance-synth-context-learning-respecting-units-table)
shows, that synthetic data can be used for in-context learning in a task
where the currency units given for a table or specific
columns[^05_results-33]. Except for Llama 3.1 all models achieved much
better numeric extraction results for tables where currencies are given
for all columns. The model also achieved better for cases where there
was only a single column with units, even though there have been no
examples with units only for one column in the synthetic data. On the
other hand, being prompted to respect currency units decreased the
performance for tables where no units are given for all models but Llama
4 Schout. This decrease was highest for Qwen 3 and also higher than 10 %
for Mistral Large.

[^05_results-33]: Synthetic data is used here because the
    characterization, which real *Aktiva* table has units in which
    column, was created too late.

\@ref(fig:compare-effect-of-respecting-units-on-overall-performance),
\@ref(fig:compare-effect-of-respecting-units-on-numeric-performance) and
\@ref(fig:compare-effect-of-respecting-units-on-NA-F1)

Thus, synthetic data can be used to solve new tasks and substitute
missing data for rare classes.

Confidence with both tasks (respect or ignore units) the same.

```{r comparing-extraction-performance-synth-context-learning-respecting-units-table, echo=echo_flag, warning=warning_flag, message=message_flag, results='asis'}
df_real_table_extraction_synth_respect_units %>% filter(method_family %in% c("top_n_rag_examples", "n_random_examples")) %>% group_by(model, method, method_family, respect_units, n_col_T_EUR) %>% reframe(mean_num = mean(percentage_correct_numeric)) %>% pivot_wider(names_from = respect_units, values_from = mean_num, names_prefix = "units_") %>% mutate(delta = units_FALSE - units_TRUE) %>% select(-c(units_FALSE, units_TRUE)) %>% pivot_wider(names_from = n_col_T_EUR, values_from = delta, names_prefix = "n_cols_T_EUR_") %>% group_by(model) %>% summarise_if(is.numeric, mean) %>% mutate(across(is.numeric, ~round(., 2))) %>% mutate_if(
    is.numeric, 
    ~ifelse(
      . == min(., na.rm = TRUE),
      paste0("**", ., "**"),
      .
    )
  ) %>% 
  render_table(
    alignment = "lrrr",
    caption = "Comparing extraction performance for real Aktiva extraction task dependent on the promt addition to respect currency units",
    ref = opts_current$get("label"), dom="t")
```

```{r comparing-extraction-performance-synth-context-learning-respecting-units-histogram, echo=echo_flag, warning=warning_flag, message=message_flag}
# df_real_table_extraction_synth_respect_units %>% group_by(model, method, respect_units, n_col_T_EUR) %>% reframe(mean_num = mean(percentage_correct_numeric)) %>% pivot_wider(names_from = respect_units, values_from = mean_num, names_prefix = "units_") %>% mutate(delta = units_FALSE - units_TRUE) %>% select(-c(units_FALSE, units_TRUE)) %>% ggplot() + geom_histogram(aes(x = delta)) + facet_grid(model~n_col_T_EUR)
```

```{r table-extraction-llm-confidence-synth-context, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Comparing confidence for reported values in the table extraction perfromance for real Aktiva extraction task with synthetic examples for in-context learning for two models."}
# confidence_vs_truth %>% ggplot() +
#   geom_boxplot(
#     aes(x = match, y = confidence, fill = truth_NA), 
#     position = position_dodge2(preserve = "single")) +
#   scale_fill_discrete(drop = FALSE) +
#   scale_x_discrete(drop = FALSE) +
#   facet_nested(~model+respect_units)
```

```{r table-extraction-llm-confidence-synth-context-table, results='asis'}
confidence_vs_truth %>% group_by(model, method, method_family, respect_units, predicted_NA) %>% reframe(mean_conf = mean(confidence)) %>% pivot_wider(names_from = respect_units, values_from = mean_conf, names_prefix = "mean_conf_units_") %>% # mutate(delta = units_FALSE - units_TRUE) %>% select(-c(units_FALSE, units_TRUE)) %>% pivot_wider(names_from = n_col_T_EUR, values_from = delta, names_prefix = "n_cols_T_EUR_") %>% group_by(model) %>% summarise_if(is.numeric, mean) %>% 
  mutate(across(is.numeric, ~round(., 2))) %>% 
  render_table(
    alignment = "lllrr",
    caption = "Comparing extraction confidence for real Aktiva extraction task dependent on the promt addition to respect currency units. No difference in confidence appearent.",
    ref = opts_current$get("label"), dom="t")
```

Risk for false NAs less with synth data for Mistral but greater for
numeric values (both).

```{r table-extraction-llm-confidence-lm-synth-context, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.cap="Estimating the relative frequency to find a wrong extraction result over different confidence intervals for predictions based on synthetic examples for in-context learning.", fig.height=3}
confidence_vs_truth_linear %>%
  filter(!respect_units) %>% 
  ggplot(aes(x = conf_center, y = chance_false, color = predicted_NA)) +
  geom_point(alpha = 0.7) +
  geom_smooth(method = "lm", se = TRUE) +
  labs(x = "Confidence Interval Center", y = "Chance False (%)", color = "Predicted NA") +
  coord_cartesian(ylim = c(0, 100), xlim = c(0,1)) +
  facet_nested(~model#+paste("respect units: ", respect_units)
               )
```

##### Hypotheses

### Comparison {#comparing-table-extraction-methods}

Most models performe better on synth tables once they have enough
in-context examples. (Needing more für random examples thanwith
top-n-rad approach). Especially Llama 3 models show wider performance
spread even with three examples
\@ref(fig:comparing-table-extraction-performance-among-real-and-synth-aktiva-data)
