---
editor_options: 
  markdown: 
    wrap: none
---

# Appendix D - Feature effect analysis
\phantomsection
\ChapFrame[Feature effect analysis][bhtred]

## Regular expressions

::: paragraph-start
##### Hypotheses

The formulated hypotheses have been evaluated visually using the
dependence and beeswarm plots from the *shapviz* library based on the
\acr{SHAP} values calculated for a random forest.
:::

```{r real-table-extraction-regex-shap-loading, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", cache=TRUE, cache.extra = tools::md5sum('data_storage/h2o/real_table_extraction_regex_h2o_results_sample_50000_shap_2000.rds')}
results <- readRDS("data_storage/h2o/real_table_extraction_regex_h2o_results_sample_50000_shap_2000.rds")

shap_num <- results$perc_numeric$shap_values$rf %>% convert_shap_x()
p1a <- shap_num %>% sv_importance(show_numbers = TRUE) +
  labs(title = "% numeric correct") +
  coord_cartesian(xlim = c(0, 0.225))
p1b <- shap_num %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.35, 0.35))

shap_f1 <- results$NA_F1$shap_values$rf %>% convert_shap_x()
p2a <- shap_f1 %>% sv_importance(show_numbers = TRUE) +
  labs(title = "F1 NA") +
  coord_cartesian(xlim = c(0, 0.225))
p2b <- shap_f1 %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.35, 0.35))
shap_f1_sum_same_line <- shap_f1 %>% sv_dependence("sum_same_line", "T_in_previous_year")

shap_binom <- results$binomial$shap_values$rf %>% convert_shap_x()
p3a <- shap_binom %>% sv_importance(show_numbers = TRUE) +
  labs(title = "binomial") +
  coord_cartesian(xlim = c(0, 0.225))
p3b <- shap_binom %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.35, 0.35))

real_table_extraction_regex_shap_plot <- (p1a | p1b) /
  (p2a | p2b) /
  (p3a | p3b)
```

```{r synth-table-extraction-regex-shap-loading, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", cache=TRUE, cache.extra = tools::md5sum('data_storage/h2o/synth_table_extraction_regex_h2o_results_sample_50000_shap_2000.rds')}
results <- readRDS("data_storage/h2o/synth_table_extraction_regex_h2o_results_sample_50000_shap_2000.rds")

shap_num <- results$perc_numeric$shap_values$rf %>% convert_shap_x()
p1a <- shap_num %>% sv_importance(show_numbers = TRUE) +
  labs(title = "% numeric correct") +
  coord_cartesian(xlim = c(0, 0.15))
p1b <- shap_num %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.5, 0.5))

p1c <- shap_num %>% sv_dependence("extraction_backend") +
  labs(title = "dependence plot for extraction_backend")
p1d <- shap_num %>% sv_dependence("header_span", color_var = "extraction_backend") +
  labs(title = "dependence plot for header_span")

shap_f1 <- results$NA_F1$shap_values$rf %>% convert_shap_x()
p2a <- shap_f1 %>% sv_importance(show_numbers = TRUE) +
  labs(title = "F1 NA") +
  coord_cartesian(xlim = c(0, 0.15))
p2b <- shap_f1 %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.5, 0.5))

shap_f1_extraction_backend <- shap_f1 %>% sv_dependence(colnames(.$X), "extraction_backend")

shap_binom <- results$binomial$shap_values$rf %>% convert_shap_x()
p3a <- shap_binom %>% sv_importance(show_numbers = TRUE) +
  labs(title = "binomial") +
  coord_cartesian(xlim = c(0, 0.15))
p3b <- shap_binom %>% sv_importance(kind = "beeswarm") +
  coord_cartesian(xlim = c(-0.5, 0.5))

synth_table_extraction_regex_shap_plot <- (p1a | p1b) /
  (p2a | p2b) /
  (p3a | p3b)

synth_table_extraction_regex_shap_num_details_plot <- p1d | p1c
```

::: paragraph-start
###### Real dataset

The formulated hypotheses have been evaluated visually using the
dependence and beeswarm plots from the shapviz library based on the
\acr{SHAP} values calculated with a random forest.
:::

Table
\@ref(tab:hypotheses-and-results-real-table-extraction-regex-table)
shows in the first column the predictors included in the random forests.
Subsequent groups of two columns show the hypotheses and the found
effects of those predictors for two aggregated measures (F1 score and
percentage of correct numeric predictions) and one value based measure
(binomial correctness rating).

Predictors that are marked with an asterisk only have five or less
representatives. Thus, those results are not reliable. Bold set
hypotheses show the predictors, that showed the highest mean \acr{SHAP}
values. For all measures but the binomial this means the effect is at
least 0.025. For the binomial measure the effect of a predictor with
bold hypothesis is at least 0.05. Results with red text highlight
hypotheses that are not supported by the visual evaluation.

Table
\@ref(tab:hypotheses-and-results-real-table-extraction-regex-table)
shows many red results, meaning the corresponding hypothesis is getting
no support. Since most of these findings show only minor effect strength
we don't interpret them as strongly challenging those hypotheses. Only
three findings regarding the F1 score show a strong effect and findings
that do not align with our hypotheses. First, it seems, that finding a
sum in the same row, has a negative effect of finding any valid number
there. Second, it has negative effect, if the previous year column is
given as *T€*. Third, it has negative effect, if the columns are
visually separated.

see Figure \@ref(fig:real-table-extraction-regex-shap-plot)

```{r hypotheses-and-results-real-table-extraction-regex-table, echo=echo_flag, warning=warning_flag, message=message_flag, class.chunk="big-table"}
create_hypotheses_table("../benchmark_results/table_extraction/hypotheses_and_results_real_table_extraction_regex.html", caption = "Comparing the formulated hypotheses and the found results for the table extraction on real Aktiva tables with the regular expression approach.")
```

::: paragraph-start
###### Synthetic dataset

Table
\@ref(tab:hypotheses-and-results-synth-table-extraction-regex-table)
shows, many red results as well, meaning the corresponding hypothesis is
getting no support. We find more predictors with a strong effect
compared to the real **Aktiva** table extraction task. The results are
based on 24_576 extracted tables and the \acr{SHAP} values have been
calculated on 2_000 examples each.
:::

Contrary to our assumption, does the *extraction_backend* have a strong
effect on all measures. We find, that *pdfium* is struggling with some
of the table characteristics while *pymupdf* is not influenced by them.
Figure \@ref(fig:synth-table-extraction-regex-shap-num-details-plot) A
shows this exemplary for the characteristic *header_span*. An example
for a erroneous text extraction with *pdfium* can be found in section
\@ref(#regex-extraction-mistakes). Actually, all results that are marked
with an asterisk are showing this effect if *pdfium* is used as
extraction backend. This can be inspected in Figure
\@ref(fig:synth-table-extraction-regex-dependence-plot-backend).

Furthermore, does the number of columns is have a positive effect
overall. Figure
\@ref(fig:synth-table-extraction-regex-shap-num-details-plot) B shows,
that this effect has inverse direction for the two libraries.

It might be worth noting, that the row for *Anteile an verbundenen
Unternehmen* was rated to have a clear negative effect on the chance to
extract the correct value.

The question, if visual separation of columns is having an effect, as
found for the real data, is not studied here, because in the synthetic
tables all columns are visually separated. But this could be
investigated in future work. It is possible, that the visual separation
is causing the faulty text extractions of *pdfium*.

```{r hypotheses-and-results-synth-table-extraction-regex-table, echo=echo_flag, warning=warning_flag, message=message_flag, class.chunk="big-table"}
create_hypotheses_table("../benchmark_results/table_extraction/hypotheses_and_results_synth_table_extraction_regex.html", caption = "Comparing the formulated hypotheses and the found results for the table extraction on synthetic Aktiva tables with the regular expression approach.")
```

```{r synth-table-extraction-regex-shap-num-details-plot, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", fig.width=8, fig.height=4, dev=std_dev, fig.cap="Showing the influence of the extraxtion library on the numeric text extraction task with synthetic data for the percentage of correct numeric predictions."}
synth_table_extraction_regex_shap_num_details_plot + 
  plot_annotation(tag_levels = 'A')
```

## Real tables

::: paragraph-start
##### Hypotheses

The formulated hypotheses have been evaluated visually using the
dependence and beeswarm plots from the shapviz library based on the
\acr{SHAP} values calculated with a random forest.
:::

```{r real-table-extraction-llm-shap-loading, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", cache=TRUE, cache.extra = tools::md5sum('data_storage/h2o/real_table_extraction_h2o_results_sample_50000_shap_2000.rds')}
results <- readRDS("data_storage/h2o/real_table_extraction_h2o_results_sample_50000_shap_2000_NA_recoded_extended.rds")

shap_num <- results$perc_numeric$shap_values$rf %>% convert_shap_x()
p1a <- shap_num %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "% numeric correct") +
  coord_cartesian(xlim = c(0, 0.12))
p1b <- shap_num %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))
p1c <- shap_num %>% sv_dependence("T_in_year", "vis_seperated_rows")
p1d <- shap_num %>% sv_dependence("vis_seperated_rows", "T_in_year")

shap_f1 <- results$NA_F1$shap_values$rf %>% convert_shap_x()
p2a <- shap_f1 %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "F1 NA") +
  coord_cartesian(xlim = c(0, 0.12))
p2b <- shap_f1 %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

shap_binom <- results$binomial$shap_values$rf %>% convert_shap_x()
p3a <- shap_binom %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "binomial") +
  coord_cartesian(xlim = c(0, 0.12))
p3b <- shap_binom %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

shap_binom <- results$confidence$shap_values$rf %>% convert_shap_x()
p4a <- shap_binom %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "confidence") +
  coord_cartesian(xlim = c(0, 0.12))
p4b <- shap_binom %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

real_table_extraction_llm_shap_plot <- (p1a | p1b) /
  (p2a | p2b) /
  (p3a | p3b) /
  (p4a | p4b)

real_table_extraction_llm_shap_num_details_plot <- p1c | p1d
```

Table \@ref(tab:hypotheses-and-results-real-table-extraction-llm-table)
shows in the first column the predictors included in the random forests.
Subsequent groups of two columns show the hypotheses and the found
effects of those predictors for two aggregated measures (F1 score and
percentage of correct numeric predictions) and two value based measures
(binomial correctness rating and reported confidence for the
prediction).

Predictors that are marked with an asterix only have five or less
representatives. Thus, those results are not reliable. Bold set
hypotheses show the predictors, that showed the highest mean \acr{SHAP}
values. For all measures but the binomial this means the effect is at
least 0.025. For the binomial measure the effect of a predictor with
bold hypothesis is at least 0.05. Results with red text highlight
hypotheses that are not supported by the visual evaluation.

For most measures the model and method related predictors
(*model_family*, *parameter_count*, *method_family* and *n_examples*)
show the strongest effects. Worth mentioning is, that *method_family*
and *n_examples* show no strong effect on the reported confidence score.
From the table related characteristics most strong effects show the
hypothesized direction. Not predicted was the negative effect of label
length on the reported confidence score. The visual separation of
columns and rows shows small effects. We find no support for a negative
effect of the fact that the **Passiva** table is on the same page as the
**Aktiva** table.

see Figure \@ref(fig:real-table-extraction-llm-shap-plot)

```{r hypotheses-and-results-real-table-extraction-llm-table, echo=echo_flag, warning=warning_flag, message=message_flag, class.chunk="big-table"}
create_hypotheses_table("../benchmark_results/table_extraction/hypotheses_and_results_real_table_extraction_llm.html", caption = "Comparing the formulated hypotheses and the found results for the table extraction on real Aktiva tables the LLM approach.")
```

## Hybrid approach

::: paragraph-start
##### Hypotheses

Table
\@ref(tab:hypotheses-and-results-real-table-extraction-llm-synth-context-table)
shows only one unsupported hypothesis for a predictor with a strong
effect: *method_family*. Figure
\@ref(fig:static-example-best-dependence-plot) shows the dependence plot
for the this predictor. The prompting strategy *static_example* shows
the highest \acr{SHAP} values. This is surprising, because the the
*static example* is equivalent to providing a single random example.
:::

```{r real-table-extraction-llm-synth-context-shap-loading, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", cache=TRUE, cache.extra = tools::md5sum('data_storage/h2o/real_table_extraction_h2o_results_sample_50000_shap_2000.rds')}
results <- readRDS("data_storage/h2o/real_table_extraction_synth_context_h2o_results_sample_50000_shap_2000_NA_recoded_extended.rds")

shap_num <- results$perc_numeric$shap_values$rf %>% convert_shap_x()
p1a <- shap_num %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "% numeric correct") +
  coord_cartesian(xlim = c(0, 0.12))
p1b <- shap_num %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))
p1c <- shap_num %>% sv_dependence("method_family")
# p1d <- shap_num %>% sv_dependence("vis_seperated_rows", "T_in_year")

shap_f1 <- results$NA_F1$shap_values$rf %>% convert_shap_x()
p2a <- shap_f1 %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "F1 NA") +
  coord_cartesian(xlim = c(0, 0.12))
p2b <- shap_f1 %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

shap_binom <- results$binomial$shap_values$rf %>% convert_shap_x()
p3a <- shap_binom %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "binomial") +
  coord_cartesian(xlim = c(0, 0.12))
p3b <- shap_binom %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

shap_binom <- results$confidence$shap_values$rf %>% convert_shap_x()
p4a <- shap_binom %>% sv_importance(show_numbers = TRUE, max_display = 25) +
  labs(title = "confidence") +
  coord_cartesian(xlim = c(0, 0.12))
p4b <- shap_binom %>% sv_importance(kind = "beeswarm", max_display = 25) +
  coord_cartesian(xlim = c(-0.5, 0.35))

real_table_extraction_llm_synth_context_shap_plot <- (p1a | p1b) /
  (p2a | p2b) /
  (p3a | p3b) /
  (p4a | p4b)

real_table_extraction_llm_synth_context_shap_num_details_plot <- p1c # | p1d
```

```{r hypotheses-and-results-real-table-extraction-llm-synth-context-table, echo=echo_flag, warning=warning_flag, message=message_flag, class.chunk="big-table"}
create_hypotheses_table("../benchmark_results/table_extraction/hypotheses_and_results_real_table_extraction_synth_context_llm.html", caption = "Comparing the formulated hypotheses and the found results for the table extraction on real Aktiva tables with the hybrid LLM approach.")
```

```{r static-example-best-dependence-plot, echo=echo_flag, warning=warning_flag, message=message_flag, out.width="100%", dev=std_dev, fig.cap="Estimating the relative frequency to find a wrong extraction result over different confidence intervals for predictions based on synthetic examples for in-context learning."}
real_table_extraction_llm_synth_context_shap_num_details_plot
```
